### 2023-12-04 main.tex
Comments Removed

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{mathtools,amsmath,amssymb}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
%\usepackage{microtype}
\usepackage[pagebackref=false]{hyperref}
\usepackage{bbm}
\usepackage{bm}
\usepackage{appendix}
\usepackage{comment}
\usepackage{mathtools}
\usepackage{array}
\numberwithin{equation}{section}
\numberwithin{equation}{subsection}

% \usepackage[margin=2.5cm]{geometry}

%%%%%  TODO NOTES ETC  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage[left=1.7cm,right=4.3cm,bottom=4cm]{geometry}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{todocolor}{HTML}{D7E1E5}
\usepackage[textsize=tiny, 
            backgroundcolor=todocolor, 
            bordercolor=todocolor, 
            linecolor=todocolor, 
            textwidth=3.9cm]{todonotes}
\newcommand{\todoWARN}[1]{\todo{\colorbox{yellow}{\parbox{3.4cm}{#1}}}}
\newcommand{\todoERR}[1]{\todo{\colorbox{red}{\parbox{3.4cm}{#1}}}}
\usepackage{comment}
\usepackage{rotating}
\usepackage[left]{showlabels}
\renewcommand{\showlabelfont}{\tiny}
\renewcommand{\showlabelsetlabel}[1]
{\begin{turn}{65}\showlabelfont \llap{#1}\end{turn}}
\newcommand{\wiptitle}[1]{
    \par\noindent\hspace*{-0.5cm}\texttt{\scriptsize\color{gray} \lbrack #1\rbrack} \par\noindent\ignorespaces}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage[textsize=tiny]{todonotes}

% \usepackage[notref,notcite]{showkeys}
\usepackage{xspace}
%\usepackage{bibentry}
%\usepackage{easybmat}

\usepackage{graphicx}

\usepackage{subcaption}

%\usepackage[numbers,sort&compress]{natbib}
%\usepackage{hypernat}

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
% MACROS
% % 
% \newcommand{\cristian}[1]{{\bf (* {\color{ForestGreen}V:{} \small #1}*)}}
% \newcommand{\jorge}[1]{{\bf (* {\color{BlueViolet}G:{} \small #1}*)}}
% \newcommand{\rouven}[1]{{ (* {\color{red}{} \small #1}*)}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\frenchspacing

\newcommand{\changelocaltocdepth}[1]{%
  \addtocontents{toc}{\protect\setcounter{tocdepth}{#1}}%
  \setcounter{tocdepth}{#1}%
}
\newcommand{\id}{I}
\newcommand{\Xt}{\widetilde{X}}
\newcommand{\co}{\;,}
\newcommand{\dt}{\;.}

\newcommand{\len}{L}
\newcommand{\spe}{N}

\newcommand{\com}[1]{{ (* {\color{red}\small #1}*)}}

\setcounter{tocdepth}{2}

%%%%%  oscillators
\newcommand{\osc}[1]{\mathbf{#1}}
\newcommand{\oscgreea}[1]{\boldsymbol{#1}}
\newcommand{\dagg}[1]{\bar{#1}}
\newcommand{\ep}{\epsilon}
\newcommand{\ob}{\osc{b}}
\newcommand{\oc}{\osc{c}}
\newcommand{\od}{\osc{d}}
\newcommand{\oab}{\dagg{\osc{a}}}
\newcommand{\obb}{\dagg{\osc{b}}}
\newcommand{\ocb}{\dagg{\osc{c}}}
\newcommand{\odb}{\dagg{\osc{d}}}
\newcommand{\ox}{\oscgreek{\xi}}
\newcommand{\oxb}{\dagg{\oscgreek{\xi}}}
\newcommand{\och}{\oscgreek{\chi}}
\newcommand{\ochb}{\dagg{\oscgreek{\chi}}}
\newcommand{\NN}{\mathbf{N}}
\newcommand{\ccc}{\mathbf{C}}
\newcommand{\qop}{Q}
\newcommand{\sop}{Y}
\newcommand{\q}{\mathbf{\epsilon}}
\newcommand{\ID}{I}
\newcommand{\per}{P}

\newcommand{\EE}{\mathrm{E}}

\newmuskip\pFqmuskip

\newcommand*\pFq[6][8]{%
  \begingroup % only local assignments
  \pFqmuskip=#1mu\relax
  \mathchardef\normalcomma=\mathcode`,
  % make the comma math active
  \mathcode`\,=\string"8000
  % and define it to be \pFqcomma
  \begingroup\lccode`\~=`\,
  \lowercase{\endgroup\let~}\pFqcomma
  % typeset the formula
  {}_{#2}\phi_{#3}{\left[\genfrac..{0pt}{}{#4}{#5};#6\right]}%
  \endgroup
}
\newcommand{\pFqcomma}{{\normalcomma}\mskip\pFqmuskip}

\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Li}{Li}
\DeclareMathOperator{\diag}{diag}

\newcommand{\KL}{\mathcal{K}}
\newcommand{\KR}{\hat{\mathcal{K}}}
\newcommand{\oa}{\mathbf{a}}
\newcommand{\dd}{\mathcal{D}_{u(x)}}
\newcommand{\oad}{\mathbf{a}^{+}}

\newcommand{\phil}{\beta_-}
\newcommand{\phir}{\beta_+}
\newcommand{\ma}{n}

\newcommand{\oaq}{\mathbf{a}_\q}

\newcommand{\oadq}{\mathbf{\bar{a}}_\q}

\newcommand{\N}{\mathbf{N}}
\newcommand{\twoj}{\nu}
\allowdisplaybreaks

\begin{document}

\begingroup
% \centering
\begin{center}
 \begingroup\LARGE
% \bf Exact solution of the multispecies stirring process \\
%  with open boundaries

\bf Duality for the multispecies stirring process 
 with open boundaries
\par\endgroup
 \vspace{3.5em}
 \begingroup\large \bf
 {Francesco Casini$^{(a,b)}$, Rouven Frassek$^{(a)}$, Cristian Giardin\`a$^{(a)}$}
 \par\endgroup
\vspace{2em}

\begingroup\sffamily

 %Università degli Studi di Modena e Reggio Emilia
$(a)$ University of Modena and Reggio Emilia, \\
Department of Physics, Computer Science and Mathematics,\\
Via G. Campi 213/b, 41125 Modena, Italy\\ 
 \vspace{.5cm}
$(b)$ University of Parma, \\
Department of Mathematical, Physical and and Computer Sciences\\
Parco Area delle Scienze, 7/A, 43124 Parma, Italy\\
 %
\par\endgroup
\vspace{2em}

% Version: \today
\end{center}
%

\thispagestyle{empty}

\begin{abstract}
\noindent
We study the stirring process with $N-1$ species
on a generic graph $G=(V,\mathcal{E})$ introduced with reservoirs.
The multispecies stirring process generalizes the symmetric exclusion process, 
which is recovered in the case $N=2$.
We prove the existence of a dual process defined on
an extended graph $\widetilde{G}=(\widetilde{V},\widetilde{\mathcal{E})}$ which 
includes additional extra-sites $\widetilde{V}\setminus V$ where
dual particles get absorbed in the long-time limit.
We thus obtain a characterization of the non-equilibrium
steady state of the boundary-driven system in terms of the absorption
probabilities of dual particles.  The process is integrable
for the case of the one-dimensional chain with two reservoirs
at the boundaries and with maximally one particle per site. Based on the results of \cite{vanicat2017exact},
we compute the absorption probabilities by relying
on the underlying ${gl}(N)$ symmetry and the
matrix product ansatz. Thus one gets a closed-formula
for (long-ranged) correlations and for the 
non-equilibrium stationary measure.
Extensions beyond this integrable set-up
are also discussed.
\end{abstract}

%\vfill

\newpage
\tableofcontents

 \newpage
\section{Introduction} 
Boundary-driven particle systems are paradigmatic models for non-equilibrium statistical mechanics. In a boundary-driven system, the model is put in contact with reservoirs at its {\color{red}``boundaries'' } imposing different particle densities, and a current is generated through the system. In the long time limit, a non-equilibrium steady state sets in, with a stationary value of the current.
 The aim of this paper is to develop the theory of stochastic duality and  study the stationary measure of certain symmetric boundary-driven {\em multispecies} particle systems with exclusion.

Our strategy to solve for the non-equilibrium steady state is of a probabilistic nature and relies on the use of {\em duality} \cite{schutzSandow}. Starting from the pioneering work of Kipnis, Marchiorio and Presutti \cite{KMP}, this approach
has been substantially developed in recent years to study stochastic processes in the boundary-driven set-up \cite{giardina2009duality,carinci2013duality}.
The boundary-driven process is mapped to a dual process with absorbing extra sites and the problem of describing the stationary state of the original system
is simplified to the problem of computing the absorption probabilities of the dual particles. In general, however,  there is little hope to compute the absorption probabilities exactly.

{\color{blue}On one hand, this absorbing duality allows to formally write the non-equilibrium steady state in a different (simpler) form. On the other hand, absorption probabilities are hard to find since they fulfil multilinear algebraic equation with proper boundary conditions. The situation simplifies when the graph reduces to a chain with two reservoirs on the boundary sites. In this context, the integrable structure of the process can be combined with duality to compute non-equilibrium steady state. Recently, probabilistic techniques have been developped to compute absorption probabilities for the simple symmetric exclusion process \cite{floreani2023non}.}
{\color{red} Say that for a  general graph with boundaries attached to any site, this is a difficult task. }
% However,  for the case of integrable systems they can be computed exactly in closed-form.

{\color{red} Explain here that we consider the chain with only two reservoirs at the end of the chain}

For the single species case,  the solution of the exclusion process with open boundaries \cite{derrida1993exact,1993JSP....72..277S} has been of crucial importance in our understanding of the structural properties of non-equilibrium state, such as long-range correlations \cite{spohn1983long} and non-local density large deviation functions \cite{derrida2007non,derrida1998exact} and current large deviations \cite{mallick2022exact,bodineau2005current}. One may argue that the results available for the exclusion process are rooted in the algebraic symmetries of the corresponding integrable spin $s=1/2$ Heisenberg chain \cite{schutzManyBody,SSEPReviewRagoucy}.

The model studied in this paper is the multispecies analogue of the exclusion process on a generic graph as introduced in \cite{zhou2021orthogonal} with additional boundary reservoirs as proposed in \cite{vanicat2017exact}. {\color{blue}This particle system is defined on a generic connected and undirected graph. At each site $\nu\in \mathbb{N}$ particles of $N$ types (species) are hosted. There is an exclusion constraint that makes the occupancy of each site dependent on $N-1$ species only. Two sites do interact when there is an edge between them and this interaction consists in an swap of two different types of particles between these sites. Furthermore, reservoirs are attached at each site of the graph, making the system interact with external environments with fixed average particle densities. When the graph is a finite chain with two reservoirs attached to the first and the last sites respectively and when $\nu=1$, the process becomes integrable (see \cite{vanicat2017exact}).}
{\color{red}Explain the model here.... $N$-species (or $N-1$ plus empty?). What is the graph? say that it becomes integrable for the line }

We refer to the model as the {\em  multispecies stirring process with open boundaries}.
%

%The  paper is organized as follows.
\subsection*{Organization of the paper}
In Section~\ref{sec1} we define the open multispecies stirring
process on a general graph and provide its Lie algebraic description in terms of $gl(N)$ generators. In Section~\ref{sectionDuality} we prove an absorption duality (Theorem~\ref{thm-duality})
and show that the problem of characterizing the non-equilibrium
steady state can be reduced to the computation of the absorption probabilities similar to the singlespecies case on the line [SPOHN?] and for a general graph \cite{ frassek2020duality}. 
% The duality is proved for a system whose spatial structure is described by a general graph.
In Section~\ref{sectionIntegrabiliy}  we consider  the case of the one-dimensional chain with at most one 
particle per site, which is driven out-of-equilbrium by two reservoirs on the left and right boundaries. In this case we obtain a closed-form expression for the stationary correlations (Theorem~\ref{thm-correlations}). Here the computation
% of the absorption probabilities 
relies on the extension of the  Matrix Product Ansatz \cite{derrida1993exact} to multiple species \cite{vanicat2017exact} and the $gl(n)$ invariance of the process in the bulk. Finally in Section~\ref{sec5} we discuss a few extensions 
and generalizations of the model, including instantaneous thermalization limits and reaction-diffusions. {\color{blue} In particular, a thermalized boundary-driven version of the multispecies stirring process is introduced, showing that it has an absorbing duality. Moreover, a further interaction mechanism of reaction type is added in order to have closed evolution equation for average particle density (inspired by \cite{casini2022uphill,schutzReaction}). Absorbing duality is proved. The duality function for the termalized and the reaction diffusion process turns out to be the same of the multispecies stirring process.} {\color{red} mention the results you have in appendix}.

% In closing this introduction we mention that we are currently working on the boundary-driven multispecies 
% model with non-compact Lie algebras, both integrable and non-integrable, which will be presented in a 
% future paper. 
% \todo{I prefer to remove this sentence}

% \cite{schutz2017kardar, vanicat2017exact, crampe2016integrable, finn2018matrix, floreani2022switching, casini2022uphill, casini2023density, redig2022ergodic, redig2023equilibrium, chatterjee2023multi,  franceschini2022orthogonal,floreani2023non,carinci2023open}

%  See also \cite{crampe2016integrable, finn2018matrix} for the asymmetric setup.

% For recent examples in the framework of singlespecies systems, see \cite{frassek2020eigenstates, frassek2020duality,frassek2022exact}.
% {\color{red}
% Here we generalize the duality approach to the set-up of multispecies particle systems.}

% {\color{blue} I think this is to general...  Do all the references study the steady state? Also the main point is about duality no?}

% {\color{red} Missing the part about non integrable model and duality.}

{\bf Acknowledgments.} We acknowledge support from Istituto Nazionale di Alta Matematica (GNFM),  the INFN grant Gauge and String Theory (GAST),  the PRIN project CUP-E53D23002220006,  the FAR UNIMORE project CUP-E93C23002040005 and  the INdAM–GNFM project CUP-E53C22001930001.

% Prin-project {\color{blue} xxxx}, FAR-project {\color{blue} xxxx} UniMoRe.
% 
% INFN Progetto Giovani

%
%{\color{blue}
%In the past years duality has been developed to study stochastic processes, especially in the boundary driven case, i.e. when the system is put in contact with boundaries that generate a non-equilibrium current. The literature is huge. We can address , for instance, to \cite{giardina2009duality,carinci2013duality,frassek2020duality}. One application of duality is the attempt to characterize the non equilibrium steady state of boundary driven process via the so called absorption probabilities. Indeed, there are many cases where the dual process has absorbing boundaries. This means that the the chain is eventually empty in the long time scale. Moreover, there are cases in which the steady state of a stochastic process can be computed explicitly, thanks to the integrability property of the system (quantum inverse scattering method, matrix product ansatz). In this direction, some examples can be found in \cite{derrida1993exact,frassek2019non,frassek2020eigenstates}. In the past, the focus has been on processes where one can distinguish, in each site of a certain graph, a (or many) particles or an empty state. However, this particles are indistinguishable, in the sense that you can pile up in a site just one type of particles. Some natural questions arise: what happens if we imagine to assign different colours to the particles? Is the duality property still holding? How far can we push our analysis for the characterization of the non equilibrium steady state?\\
%The way one can define a multi-type (or multispecies) interacting particle system is not unique. In the literature there are many. Two of them are the following. On one hand, the \textit{multi-layer models}, in which the process is constructed on a many layer graph where each graph correspond to a type and perform a kind of dynamic (independent jump, exclusion/inclusion walkers) and the layers communicate each other with some "switching probabilities". The state space of the Markov process is the Cartesian product of the state spaces of each layer, i.e. of the two species. For example see \cite{floreani2022switching,redig2022ergodic,redig2023equilibrium}. On the other hand, the \textit{single layer models}. Here, there is just one layer, where particles of each type can pile up, in a way that the exclusion dynamic concerns many species together. In other words, now the species can interact "directly", not just by switching layer. Here the state space contains all the species and it is not product anymore. Some examples of such a dynamic can be found in \cite{casini2022uphill,vanicat2017exact,zhou2021orthogonal,casini2023density}. \\
%In this work we concentrate our effort in this last "single layer" case with a repulsive process, i.e. a dynamic in which a maximum amount of particles can be present in each site at each time. We will call this amount, the "maximal occupancy" and it will indicated with $\twoj$ where $\twoj\in\mathbb{N}$. Particle of any species can exchange their position in the graph, under this exclusion constraint. The system will be put out of equilibrium and some absorbing duality result are investigated. Moreover, we will use this duality, to study the out of equilibrium steady state. In particular, in case of $\twoj=1$ we retrieve the model studied in \cite{vanicat2017exact}. 
%{\color{red}
%The case of $\twoj=1$ distinguishes itself from other values of $\twoj$ as in this case the model can be mapped to an integrable higher rank ${gl}(N)$ Heisenberg XXX spin chain with open (non-diagonal) boundaries. Furthermore the matrix product ansatz \cite{derrida1993exact}  is applicable, see \cite{Crampe:2014aoa,vanicat2017exact}.
%}
%We will exploit the duality and the underlying symmetries of the model, to write explicitly all the correlations of this \textit{hard-core exclusion} case. The main tool that we will use is the Lie algebraic description of the process, that allows to find the duality as a proper symmetry of the generator. \\
% For a complete review of the quantum inverse scattering method see \cite{Sklyanin:1988yz} and \cite{faddeev1996algebraic}. For application of the QISM to interacting particle systems we refer to \cite{frassek2020eigenstates,frassek2019non,frassek2022exact,vanicat2017exact}. \\
% A new technique to compute absorption probabilities has been proposed in \cite{floreani2023non}.
% \\
% \\
% \textbf{Some interesting references and extensions?}:
% Multispecies version of non-compact chains. Extension of \cite{frassek2022exact},\cite{frassek2022integrable},\cite{frassek2020non},\cite{frassek2021duality}. \\
% Use exact solution to compute large deviation rate function extending to the multispecies case results proved in \cite{derrida1998exact}, \cite{mallick2022exact}. \\
% Extend the result to a-symmetric situation. See \cite{belitsky2018self}, \cite{belitsky2015quantum}, \cite{crampe2016integrable}, \cite{finn2018matrix}. \\
% Extend the QISM technique to models which include reaction trying to apply the method exposed in \cite{de2021yang} for stirring process with a proper reaction term. Extend the duality to stirring models with different mechanism of transition \cite{chatterjee2023multi}. 
%}
%
%
%
%
\section{The stirring process with open boundaries}
\label{sec1}
\subsection{Informal description of the process}
The process studied in this paper is the {\em multispecies stirring process}
with open boundaries. 
In words, the dynamics is described as follows. Each site
of a connected graph can host up to $\nu\in \mathbb{N}$ particles.
The particles have a {\em type} (sometimes called {\em species} or {\em color})
which can take $N$ values.
Then, the dynamics has two parts: on each edge of the graph, 
any two types are swapped at rate $1$; additionally, on each site $x$, 
a particles of a given type (if present) is replaced with a
particle of type $a$ at rate $\alpha_a^x >0$.
The swap dynamics taking place on the edges is of Kawasaki-type 
with $N$ conservation laws
(the total number of particles of each type). 
The site-dynamics is instead of Glauber-type. 
In the long-time limit, a so-called non-equilibrium
steady state sets in.

%\noindent
%\begin{remark}
%\label{hole}
In the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types. 
% It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
% (those of types $\{1,\ldots,N-1\}$), we retrieve the 
% boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
In the case $N=2$, we retrieve the  boundary-driven version of the symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
% \todo{I would prefer to say this is for N=2}
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. However, we shall see in Section~\ref{sectionDuality}, that
the dual process breaks the symmetry.
In particular, in the  long-time limit the dual process has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.   
%\end{remark}
\todo{It is misleading to have to different interpretations. Like what is the number of particles per site? Another confusion is for the reservoirs/open boundaries... Is it the same on general graph?}
{\color{blue} First question: Do you mean that we should specify it like "the maximal occupancy at each site is..."?\\
Second question: I think yes. The concept of boundary is general, not only for the chain}

\subsection{The process generator}\label{subsectionGeneratorStr}
We now give the mathematical description of the multispecies stirring process.
We consider a connected graph $G=(V,\mathcal{E})$ with vertex set $V$ and edge set $\mathcal{E}$.
At each vertex site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $a$-th component $n_{a}^{x}$ denotes the number of particles of species $a\in \{1,\ldots,N\}$. On each vertex we allow a total  maximal occupancy $\nu\in \mathbb{N}$. Then, the configuration space of the process on the graph $G$ is 
\begin{equation}\label{stateSpace}
    \Omega:=\bigotimes_{x\in V} \Omega_{x}
\end{equation}
where
\begin{equation}
\Omega_{x}:=\left\{n^x=(n_{1}^{x},\ldots,n_{N}^{x})\in\mathbb{N}_0^{N}\;:\; \sum_{a=1}^{N}n_{a}^{x}=\twoj\right\}\,.
\end{equation}
We denote a particle configuration on the graph as $\bm{n}\in \Omega$, where $\bm{n}=(n_{a}^{x})_{x\in V,\,a\in\{1,\ldots,N\}}$.
The infinitesimal generator of the process reads
\begin{equation}\label{Generator}
    \mathcal{L}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\mathcal{L}_{x}
\end{equation}
where  $ \omega_{x,y}\geq 0$ are so-called conductances and $\Gamma_{x}\geq 0$ are the couplings to reservoirs. 
The generator $\mathcal{L}_{x,y}$ is called the \textit{edge generator}, while $\mathcal{L}_{x}$ is called the \textit{site generator}. These linear operators act on functions $f:\Omega\to \mathbb{R}$ as follows
\begin{equation}\label{edgeGenerator}
\mathcal{L}_{x,y}f(\bm{n})=\sum_{a,b=1}^{N}n_{a}^{x}n_{b}^{y}\left[f(\bm{n}-\bm{\delta}^{x}_{a}+\bm{\delta}_{b}^{x}+\bm{\delta}_{a}^{y}-\bm{\delta}_{b}^{y})-f(\bm{n})\right]\co
\end{equation}
\begin{equation}\label{siteGenerator}
    \mathcal{L}_{x}f(\bm{n})=\sum_{a,b=1}^{N}\alpha_{a}^{x}n_{b}^{x}\left[f(\bm{n}+\bm{\delta}_{a}^{x}-\bm{\delta}_{b}^{x})-f(\bm{n})\right]\co
\end{equation}
where 
\begin{equation}
(\bm{\delta}_{a}^{x})^{y}_{b}=\begin{cases}
1\qquad &\text{if}\quad y=x,\;b=a\;,\\
0\qquad &\text{otherwise}\dt
\end{cases}
\end{equation}
Thus the dynamics consists in an exchange of particles between connected vertices at Poissonian times. 
This means that on the edge $(x,y)\in \mathcal{E}$  a particle of type $a$ at site $x$ is exchanged with a particle of type $b$ at site $y$ at rate $\omega_{x,y}n_{a}^{x}n_{b}^{y}$. 
Moreover, each vertex  exchanges particles with the external environment (reservoirs). 
Namely, on each site $x\in V$  a particle of type $b$ is replaced with a particle of type $a$ at rate $\Gamma_{x}\alpha_{a}^{x}n_{b}^{x}$. 
%
%For the sake of notation we denote $|\alpha^{x}|=\sum_{a=1}^{N}\alpha_{a}^{x}$. 
%See Figure~\ref{fig:1} and  Figure~\ref{fig:2} for a visualisation of the dynamics. 
%\begin{figure}[h]
%    \centering
%    \includegraphics[scale=0.45]{BulkStirring.eps}
%    \caption{The edge dynamics. Maximal occupation is $\nu=6$. A particle of type $2$ is removed in $x$ and placed in $y$. A particle of type $3$ is removed in $y$ and placed in $x$. This transition occurs at rate $n_{2}^{x}n_{3}^{y}=2\cdot 3$.}
%    \label{fig:1}
%\end{figure}
%\begin{figure}[h]
%    \centering
%    \includegraphics[scale=0.4]{BoundaryStirring.eps}
%    \caption{The site dynamics. Here the maxiaml occupation is $\nu=6$. The reservoir, represented by a square "box" containing all types of particles, remove a type $2$ from site $x$ and replace it with a type $3$. This transition occurs at rate $\alpha_{3}^{x}n_{2}^{x}$.}
%    \label{fig:2}
%\end{figure}
\subsection{Reversible measures in the equilibrium set-up}
For a particular choice of the reservoir parameters one has an $N$-parameter family of reversible measures. More precisely
when the parameters are the same on each site, i.e.
\begin{equation}\label{reversibilityCondition}
\alpha_{a}^{x}=\alpha_{a}\qquad \forall x\in V\,,
\end{equation}
then the process described by the generator \eqref{Generator} is reversible with respect to the 
homogeneous product measure 
\begin{equation}
\label{reversibleMeasure}
\mu_{\text{rev}}=\bigotimes_{x\in V}\mu_{\text{rev}}^{x}
\end{equation}
with marginals $\mu_{\text{rev}}^{x}$ given by  the multinomial distribution
\begin{equation}
 \mu^{x}_{\text{rev}}\sim \text{Multinomial}\left(\twoj,\rho_{1},\ldots,\rho_{N}\right)\,.
\end{equation}
Here
 $$
\rho_{a}=\frac{\alpha_{a}}{|\alpha|}\co
$$
is the density of species $a$ and we used the notation $|\alpha|=\sum_{a=1}^{N}\alpha_{a}$. Explicitly, 
\begin{equation}
\mu_{\text{rev}}^{x}(n^{x})=\frac{\nu!}{\prod_{a=1}^{N}n_{a}^{x}!}\prod_{a=1}^{N}\rho_{a}^{n_{a}^{x}}\,.
\end{equation}
This can be proved  by checking that detailed balance is satisfied. 
If condition \eqref{reversibilityCondition} is not met, then in general reversibility is lost: indeed, in this situation, each reservoir at site $x$ has  its own set of densities vector $\rho^{x}=(\rho_{1}^{x},\ldots,\rho_{N}^{x})$ with components
\begin{equation}
	\label{rhox}
	\rho_{a}^x=\frac{\alpha_{a}^x}{|\alpha^x|}
\end{equation}
and  $|\alpha^x|=\sum_{a=1}^{N}\alpha_{a}^x$. {\color{blue} As a consequence, particles are injected and removed in the graph with different rates. When, for any $a\in \{1,\ldots,N\}$, there are at least two reservoirs at two different sites $x,y$ such that $\rho_{a}^{x}\neq  \rho_{a}^{y}$, a  non-zero current sets in, breaking the reversibility. Informally, one can say that the reservoirs try to impose their different densities at the boundaries of the the graph, putting the system out-of-equilibrium.} %{\color{red}wants to fix}
\todo{?}
%Currents will arise when at least two reservoirs have different densities. 
\begin{remark}
\label{rem-rev}
		{\color{blue}One can check that conditions \eqref{reversibilityCondition} are not the most general that implies reversibility. Indeed, by choosing the reservoir parameters as }
		\begin{equation}
		\label{cond-gen}
			\alpha_{k}^{x}\left(|\alpha^{y}|-\alpha_{k}^{y}\right)=\alpha_{k}^{y}\left(|\alpha^{x}|-\alpha_{k}^{x}\right)\qquad\qquad \forall x,y\in V,\quad \forall k\in \{1,\ldots,N\}
		\end{equation}
		where
		\begin{equation}
			|\alpha^{x}|=\sum_{a=1}^{N}\alpha_{a}^{x}\dt
		\end{equation}
		reversibility holds. 
		%This is slightly more general than \eqref{reversibilityCondition}.
		 In the case where the graph is a chain with nearest neighbors interactions and two reservoirs are attached to the two boundary sites
		then condition \eqref{cond-gen} is the multispecies counterpart of $\alpha\delta=\beta\gamma$ of the boundary driven partially excluded process introduced in \cite{schutzSandow} (see Section 3.2 of \cite{carinci2013duality}).
		\end{remark}
% 		\todo{?}
\subsection{Lie algebraic description of the process}

In this paper we will often use the fact that the Markov generator of the multispecies stirring process can be described in terms a Lie algebra ${gl}(N)$.
In this section we provide details about this.

Consider the Lie algebra ${gl}(N)$ with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{cb}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}\,.
\end{equation}
The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with  $\lambda_i\geq \lambda_{i+1}$,  $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu\in\mathbb{N}$.  
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \;.
    %\qquad\text{where}\qquad \twoj\in\mathbb{N}\,.
\end{equation} 
The dimension $M_\twoj$ of this symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition, namely
\begin{equation}
	M_\twoj= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}\dt
\end{equation} 
The generators of the symmetric representations will be denoted by $E_{ab}$.
A basis of the vector space $\mathbb{C}^{M_\twoj}$ are the column vectors denoted by
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle,\quad \text{with}\quad n_{i}\in\mathbb{N}_{0}\quad \text{such that}\quad \sum_{i=1}^{N}n_{i}=\nu\,.
\end{equation}
%where $\sum_{a=1}^{N}n_{i}=\nu$. 
The basis vectors satisfy the orthogonality relation
\begin{equation}\label{ortho}
   \langle m|n \rangle =\langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{a=1}^{N}\delta_{m_{a},n_{a}}\co
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vector obtained by transposing $|m\rangle=|m_{1},\ldots,m_{N}\rangle$ and $\delta_{m_{a},n_{a}}$ is the Kronecker delta.

The explicit action of the algebra generators on the basis  vectors $|n\rangle$ is the following:
\begin{equation}\label{actionE}
	\begin{cases}
		E_{ab}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle =n_{b}|n_{1},\ldots,n_{a}+1,\ldots,n_{b}-1,\ldots,n_{N}\rangle\qquad\qquad a\neq b\\[0.1cm]
		E_{aa}|n_{1},\ldots,n_{a},\ldots,n_{N}\rangle = n_{a} |n_{1},\ldots,n_{a},\ldots,n_{N}\rangle\dt
	\end{cases}
\end{equation}  
The matrices defined in this way satisfy the commutation relations \eqref{eq:comgl} and yield highest Dynkin weight \eqref{eq:dynkin}. \\
%{\color{red} Is this remark relevant for us?}
%\newline 
%\textbf{Remark}: it is possible to define these matrices in terms of creation and annihilation operators: see Appendix~\ref{appA}.
\\ \\
%\textbf{Remark}: The symmetric representations $\lambda=(\twoj,0,\ldots,0)$ are dual to the representations $\lambda=(\twoj,\ldots,\twoj,0)$ that can be obtained via 
%\begin{equation}
%   \bar E_{ab}=\nu\delta_{ab}-E_{N-b+1,N-a+1}\dt
%\end{equation}
%{\color{red} This needs to be checked and give alternative form of Hamiltonian?}\\
As mentioned above, the process with generator \eqref{Generator} can be described in terms of ${gl}(N)$ Lie algebra generators. The state space \eqref{stateSpace} is given by the $|V|$-fold tensor product of the vector space with basis elements $|n^x\rangle$ at a given site. Namely,
a vector $|{\bm{n}}\rangle \in \Omega$ can be written as
%We now define the equivalent of \eqref{stateSpace} in the vector notation
% \begin{equation}
%	\Omega':=\left\{|n\rangle=|n_{1},\ldots,n_{N}\rangle \;:\;n\in\mathbb{N}_0^N,\;\;|n|=\twoj\right\}^{\otimes|V|}
%	\end{equation}
%where we denote 
\begin{equation}
|{\bm{n}}\rangle=\left(\,\bigotimes_{x\in V}	|n_{1}^{x},\ldots,n_{N}^{x}\rangle\right)
\end{equation}
%{\color{blue}
%	\begin{equation*}
%		|{\bm{n}}\rangle=\left(\,\bigotimes_{x\in V}	|n_{1},\ldots,n_{N}\rangle_{x}\right)
%	\end{equation*}}
%where we require that $n_{a}^{x}\in \mathbb{N}_{0}$ and $\forall x\in V$ we have 
with $\sum_{a=1}^{N}n_{a}^{x}=\nu$ for any $x\in V$. For a fixed $x\in V$ we write $|n^{x}\rangle=|n_{1}^{x},\ldots,n_{N}^{x}\rangle$. The following orthogonality relation is a consequence of the single site relation \eqref{ortho}
\begin{equation}
    \langle {\bm{n}}|{\bm{m}}\rangle =\prod_{x\in V}\prod_{a=1}^N\delta_{n^x_a,m^{x}_a}\,.
\end{equation}
We introduce the Hamiltonian operator
\begin{equation}\label{OriginalHamiltonian}
	\begin{split}
		H=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{H}_{x,y}+\sum_{x\in V}\Gamma_{x}H_{x}
	\end{split}
\end{equation}
where the edge Hamiltonian $\mathcal{H}_{x,y}$ that describes the interaction between two connected sites is
\begin{equation}\label{edgeHamiltonian}
\mathcal{H}_{x,y}=\sum_{a,b=1}^{N}\Big(E_{ab}^{x} E_{b a}^{y}-E_{bb}^{x} E_{aa}^{y}\Big)
 \end{equation}
  and where the site Hamiltonian $H_{x}$  is
 \begin{equation}\label{siteHamiltonian}
H_{x}=\sum_{a,b=1}^{N}\alpha_{a}^{x}\left(E_{ab}^{x}-E_{bb}^{x}\right)
\end{equation}
Here $E_{ab}^{x}$ denotes the generator $E_{ab}$ in \eqref{actionE} acting non-trivially on site $x$ (and as the identity on the other sites). 
The Hamiltonian in \eqref{OriginalHamiltonian} is stochastic and is linked to the Markov generator  by
\begin{equation}\label{Hamiltonian-Generator}
H=\mathcal{L}^{T} 
\end{equation}
where $T$ denotes transposition, see e.g.  \cite{belitsky2015self}.
The action of the generator on functions $f$ can then  be expressed as 
\begin{equation}
    \mathcal{L}f( {\bm{n}})=\langle f|H| {\bm{n}}\rangle
\end{equation}
where 
\begin{equation}
    \langle f|=\sum_{ {{\bm{m}}\in \Omega}}f( {\bm{m}})\langle  {\bm{m}}|\dt
\end{equation}
We can write the edge Hamiltonian \eqref{edgeHamiltonian} as a function of the coproduct of the quadratic Casimir of ${gl}(N)$
\begin{equation}\label{secondCasimir}
    C=\sum_{a,b=1}^{N}E_{ab}E_{ba}\co
\end{equation}
that acts diagonally as $C|{n}\rangle=\twoj(\twoj+N)|{n}\rangle$ on any state $|n\rangle$ and belogs to the center of ${gl}(N)$ (i.e. it commutes with all the algebra elements).  
More precisely,  considering the standard coproduct 
\begin{equation}
\begin{split}
\Delta:{gl}(N)\to {gl}(N)\otimes {gl}(N)
\end{split}
\end{equation}
with
\begin{equation}
E_{ab}\mapsto E_{ab}\otimes \mathbbm{1}+\mathbbm{1}\otimes E_{ab}\co
\end{equation} 
we have 
\begin{equation}
\Delta(C)=\sum_{a,b=1}^{N}\Delta(E_{ab})\Delta(E_{ba})=2\sum_{a,b=1}^{N}E_{ab}\otimes E_{ba}+C\otimes \mathbbm{1}+\mathbbm{1}\otimes C\,.
\end{equation}
Then, one can check that 
\begin{equation}\label{hamiltonianCasimir}
	\mathcal{H}_{x,y}=\frac{1}{2}\Delta_{x,y}(C)-\twoj(2\twoj+N)
\end{equation}
where $\Delta_{x,y}(C)$ denotes a copy of $\Delta(C)$ acting on  the sites of edge $(x,y)\in \mathcal{E}$ and acting trivially on the other sites of the graph.

\subsection{Integrable process on a line segment}\label{subsection-description-process-LINE}
In this section we specialize the multispecies stirring process to the geometry of the one-dimensional chain with sites $\{1,\ldots,L\}$ where two reservoirs are attached to the boundary sites $1$ and $L$.  The reservoirs, exchanging particles with the external environment, put the chain out of equilibrium.  The  Hamiltonian that we consider here is obtained from \eqref{OriginalHamiltonian} assuming that the conductances are
\begin{equation}
	\omega_{x,y}=\begin{cases}
		1 \quad \text{if}\quad |x-y|=1\\
		0\quad \text{otherwise}
	\end{cases}
\end{equation}
and the coupling to reservoirs are
\begin{equation}
	\Gamma_{x}=\begin{cases}
		1\quad \text{if} \quad x\in \{1,L\}\\
		0\quad \text{otherwise}
	\end{cases}\dt
\end{equation}
The case $\nu=1$, i.e. one particle at most for each site, is integrable and has been considered previously in \cite{vanicat2017exact}. 
For  $\nu=1$ we denote the ${gl}(N)$  generators   as  $(e_{ab})_{a,b\in\{1,\ldots,N\}}$ obeying $(e_{ab})_{cd}=\delta_{ac}\delta_{bd}$.
The Hamiltonian can then be written as
\begin{equation}\label{hamiltonian}
	H=H_{\text{left}}+H_{\text{bulk}}+H_{\text{right}}
\end{equation}
where
\begin{equation}
	H_{\text{bulk}}=\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}\dt
\end{equation}
Here $\mathcal{H}_{x,x+1}$  denotes  the two-site Hamiltonian 
\begin{equation}\label{H-corsivo}
	\begin{split}
		\mathcal{H}=P-\id
	\end{split}
\end{equation}
with the permutation matrix
\begin{equation}
	P=\sum_{a,b=1}^Ne_{ab}\otimes e_{ba}\,,
\end{equation} 
acting non-trivially on the vertices of the edge $x,x+1$.
 In this context it will be useful to introduce the following notation for the occupation variables of the process. Each configuration is denoted by
\begin{equation}\label{Tau-Notation}
	|\bm{\tau}\rangle=|\tau_{1},\ldots,\tau_{L}\rangle
\end{equation} 
with $\tau_{x}\in \{1,\ldots,N\}$, $\forall x\in \{1,\ldots,L\}$. Since the maximal occupancy at each site is $\nu=1$,\todo{but is it at max 1? I do not understand what we call the stirring process....}  the configuration $\bm{n}$ introduced in Section~\ref{subsectionGeneratorStr} and $\bm{\tau}$ are related by 
\begin{equation}\label{notation-change-relation}
	n_{a}^{x}=\delta_{\tau_{x},a}\dt
\end{equation}
The state space of the process is now 
\begin{equation}
	\Omega^{'}=\left\{(\tau_{1},\ldots,\tau_{L})\;:\; \tau_{x}\in\{1,\ldots,N\}\right\}\dt
\end{equation}
The action of the Hamiltonian density $\mathcal{H}$ on the tensor product of the configuration of two sites follows immediately from \eqref{H-corsivo} and reads
\begin{equation}
	\begin{split}
		\mathcal{H}| \tau\rangle\otimes   |\tau'\rangle&=|\tau'\rangle \otimes |\tau\rangle-|\tau\rangle \otimes|\tau'\rangle\dt
	\end{split}
\end{equation}
The boundary terms of the Hamiltonian \eqref{hamiltonian}  are given by 
\begin{equation}
	H_{\text{left}}=\begin{pmatrix}
		\alpha_{1}-1&\alpha_{1}&\alpha_{1}&\ldots&\ldots&\alpha_{1}\\
		\alpha_{2}&\alpha_{2}-1&\alpha_{2}&\ldots&\ldots&\alpha_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\alpha_{N-1}&\alpha_{N-1}&\ldots&\ldots&\alpha_{N-1}-1&\alpha_{N-1}\\
		\alpha_{N}&\alpha_{N}&\ldots&\ldots&\alpha_{N}&\alpha_{N}-1
	\end{pmatrix}
\end{equation}
and 
\begin{equation}
	H_{\text{right}}=\begin{pmatrix}
		\beta_{1}-1&\beta_{1}&\beta_{1}&\ldots&\ldots&\beta_{1}\\
		\beta_{2}&\beta_{2}-1&\beta_{2}&\ldots&\ldots&\beta_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\beta_{N-1}&\beta_{N-1}&\ldots&\ldots&\beta_{N-1}-1&\beta_{N-1}\\
		\beta_{N}&\beta_{N}&\ldots&\ldots&\beta_{N}&\beta_{N}-1
	\end{pmatrix}
\end{equation}
where, without loss of generality,  we assume that the parameters satisfy
\begin{equation}\label{ratesConditions}
	\sum_{a=1}^{N}\alpha_{a}=1,\qquad\sum_{a=1}^{N}\beta_{a}=1\;.
\end{equation} 
%We define also 
%\begin{equation}\label{lambdaConditions}
%	(\alpha_{a}-\beta_{a})=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=1}^{N}(\alpha_{a}-\beta_{a})=0\,.
%\end{equation}
Under these assumption the Hamiltonian~\ref{hamiltonian} is integrable. This has been schown in \cite{vanicat2017exact} within  the Quantum Inverse Scattering Method.  We will consider this process more extensively in Section~\ref{sectionIntegrabiliy}.

\section{Duality}\label{sectionDuality}
In this section we prove the first main result of this paper. We show that the multispecies stirring process with open boundaries
can be studied by means of a dual process. After recalling the definition of duality between two
Markov processes in Section~\ref{def-duality}, we formulate our duality result  in Theorem~\ref{thm-duality}. A proof of the theorem is given in Section~\ref{proof-th-duality}.
The duality result yields {\color{blue} to a dual process that} %several simplifications,  the main one being that the
%dual process
 has absorbing sites\todo{which else?}. Therefore, as shown in Corollary~\ref{Corolollary-ABS_Corr-abstract}, the study of correlations in the non-equilibrium steady state of the multispecies stirring process with open boundaries
is reduced to the study of the absorption probabilities of dual particles.

\subsection{Definition}
\label{def-duality}
Consider two Markov processes, $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
\begin{equation}\label{duality-expectation}
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
\end{equation}
where $\mathbb{E}_{\eta}$ denotes the expectation with respect to the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect to the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators, provided some minimal technical conditions are satisfied \cite{jansen2014notion}. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)\dt
\end{equation}
In the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the process is self-dual.
\begin{remark}
When the state spaces of the dual processes is countable, the generators of the processes and the duality function can be represented as matrices with elements $\mathcal{L}(\eta,\eta')$, $\widetilde{\mathcal{L}}(\xi,\xi')$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta'\in\Omega$ and $\xi,\xi'\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
\begin{equation}
    \sum_{\eta'\in\,\Omega}\mathcal{L}(\eta,\eta')D(\eta',\xi)=\sum_{\xi'\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi')D(\eta,\xi')\co
\end{equation}
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{\,T}
\end{equation}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines}  intertwines between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}^{\,T}$. Working in terms of the Hamiltonian operators the duality relation reads 
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}\dt
\end{equation}
\end{remark}
\subsection{Duality for the open multispecies stirring process}\label{statementDualitySubsection}
In this section we formulate duality for the multispecies stirring process $(\bm{n}(t))_{t\geq 0}$ with open boundaries, defined by the generator \eqref{Generator}.
The dual  process $(\bm{\xi}(t))_{t\geq 0}$ is defined on the \textit{enlarged graph} $\widetilde{G}=(\widetilde{V},\widetilde{\mathcal{E}})$ where 
\begin{equation}
	\widetilde{V}:=V\cup \left\{u(x)\,:\, x\in V\right\}\qquad \widetilde{\mathcal{E}}:=\mathcal{E}\cup \left\{(x,u(x))\,:\, x\in V\right\}\dt
\end{equation}
This means that to each site $x\in V$ we associate an ``extra-site'' via a bijection $u:V\to V$. We denote as $u(x)$ the extra-site associated to $x\in V$ . The configuration space of the dual process is the enlarged state space
\begin{equation}\label{dualStateSpace}
    \widetilde{\Omega}= \bigotimes_{x\in V} \widetilde{\Omega}_{x}\ = \bigotimes_{x\in V} (\Omega_{x}\times \mathbb{N}_{0}^{N-1})\dt
\end{equation}
Note that on the extra-site we allow an unbounded number of particles. Thus
dual particles will accumulate in these extra-sites in the course of time. 
We write the configurations $\bm{\xi} \in \widetilde\Omega$  as
\begin{equation}
    \bm{\xi}=\left(\left(\xi_{1}^{x},\ldots,\xi_{N}^{x}\right)\otimes (\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)} )\right)_{x\in V}
\end{equation}
\todo{OK? but this has no index $x$?}
where the component $\xi_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
and the component $\xi_{a}^{u(x)}$  gives the number of dual particles of type $a\in \{1,\ldots,N-1\}$ at 
the extra-site $u(x)$ connected to $x\in V$. {\color{blue} We observe that the configuration variable at the extra site $u(x)$ does not have to satisfy any exclusion constraint, i.e. an unbounded number of particles can be hosted at extra-site $u(x)$. Therefore, the occupation variable of species $N$ (the one dependent on the species $1,\ldots,N-1$) is not present}. We state the following duality result.
\begin{theorem}[Absorbing duality]\label{thm-duality}
	The multispecies stirring process $(\bm{n}(t))_{t\geq 0}$ defined on the state space $\Omega$ with generator $\mathcal{L}$ defined in \eqref{Generator} is dual to the process $(\bm{\xi}(t))_{t\geq 0}$ defined on the enlarged state space $\widetilde{\Omega}$ with generator
	 \begin{equation}\label{DualGenerator}
		\widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
	\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
	\widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=1}^{N-1}\xi_{a}^{x}\left(f(\bm{\xi}-\bm{\delta}_{a}^{x}+\bm{\delta}_{N}^{x}+\bm{\delta}_{a}^{u(x)})-f(\bm{\xi})\right)\dt
\end{equation}
The duality function is given by 
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\,\right)
\end{equation}
where we recall the definition of the reservoir densities (cf. \eqref{rhox})
\begin{equation}
	\rho_{a}^{x}=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}\dt
\end{equation}
\end{theorem}
The dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual Markov generator gives rise to  the multispecies stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{1,\ldots,N-1\}$ at site $x$ with a particle of type $N$ and creates a particle of the same type $a$ at the extra-site $u(x)$. This last transition is performed with rate $|\alpha^{x}|\xi_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{1,\ldots,N-1\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\begin{remark} In the reversible situation, i.e. when $\forall x\in V$ we have $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{\text{rev}} = \bigotimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu_{\text{rev}}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=1}^{N-1}\left(\rho_{a}\right)^{\sum_{x\in V}\xi_{a}^{x}+\sum_{x\in V}\xi_{a}^{u(x)}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}\dt
\end{equation}
\end{remark}
%
%
\subsection{Proof of Theorem~\ref{thm-duality}}
\label{proof-th-duality}
To prove duality between the process $(\bm{n}(t))_{t\geq 0}$ and the process  $(\bm{\xi}(t))_{t\geq 0}$ we  show that \eqref{DualityRelation} is fulfilled.  To show this, we will use the Hamiltonians and their Lie algebraic description. Indeed, in this formalism the proof
reduces to finding symmetries of the generator (for the bulk duality) and group like transformations (for the ``boundary duality'').
To a configuration $\bm{\xi}$ of the configuration space  \eqref{dualStateSpace} of a dual process we associate the vector
\begin{equation}
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes |\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle\right)\dt
\end{equation}
{\color{blue} Here $|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle$ is the vector associated to the configuration of in the extra site $u(x)$.  For all $q_{1},\ldots,q_{N-1}\in \mathbb{N}_{0}$, we assume that it satisfy an orthogonality relation
\begin{equation}\label{ortho-extraSite}
	\langle q_{1},\ldots,q_{N-1}|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle=\prod_{a=1}^{N-1}\delta_{\xi_{a}^{u(x)},q_{a}}
\end{equation} }
The Hamiltonian of the dual process reads
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\mathcal{H}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}
where $\mathcal{H}_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=|\alpha^{x}|\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{u(x)}\,E_{Na}^{x}-E_{aa}^{x}\right)\dt
\end{equation}
Here we introduced the pair of bosonic operators $\mathbf{a},\,\mathbf{a}^{+}$ satisfying $[\mathbf{a},\mathbf{a}^{+}]=1$ and acting as
\begin{equation}
	\mathbf{a}^{+}|q\rangle=|q+1\rangle\qquad \mathbf{a}|q\rangle=q|q-1\rangle
\end{equation}
and 
\begin{equation}\label{bosonic_2}
	\langle q|\mathbf{a}^{+}=\langle q-1|\qquad \langle q|\mathbf{a}=(q+1) \langle q+1|
\end{equation}
on a generic vector $\langle q|$ with $q\in \mathbb{N}_{0}$, so that in \eqref{siteDualHamiltonian} 
$(\mathbf{a}^{+})_{a}^{u(x)}$ denotes a copy of $\mathbf{a}^{+}$ acting on the extra-site $u(x)$ and on the species $a\in\{1,\ldots,N-1\}$. \\

We will show below that the Hamiltonians \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} are dual in the sense of \eqref{DualityRelation}. From an algebraic point of view, the duality matrix $D$ \eqref{dualityElements} is described as 
\begin{equation}\label{dualityMatrix}
    D=\prod_{x\in V}d_{x}\otimes \dd \;.
\end{equation}
Here
\begin{equation}\label{bulkElementDualityMatrix}
d_{x}=R_{x}\exp{(E^{x})}
\end{equation}
with 
\begin{equation}\label{Revmatrix}
    R_{x}=\sum_{n^{x}\in\Omega_{x}}\frac{\prod_{a=1}^{N}n_{a}^{x}!}{\nu!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle\langle n_{1}^{x},\ldots,n_{N}^{x}|
\end{equation}
and
\begin{equation}\label{EquationEx}
E^{x}=\sum_{a=1}^{N-1}E_{aN}^{x}\,.
\end{equation}
Furthermore
\begin{equation}\label{dualityMatrix2}
\dd=\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|\dt
\end{equation}
We observe that the matrix $R_{x}$ is diagonal. Its elements are related to the inverse of the weights of the reversible measure \eqref{reversibleMeasure}. In particular, to obtain these elements, we have considered the weights of \eqref{reversibleMeasure} when all the parameters $\rho_{a}=\frac{1}{N}$. Then, the constant $\left(\frac{1}{N}\right)^{\nu}$ has been neglected, since it does not change the duality relation. This $R_{x}$   is associated to reversibility, see \cite{giardina2009duality}. \\

Since \eqref{dualityMatrix} is product over sites, proving \eqref{DualityRelation} is equivalent to showing that 
\begin{equation}\label{edgeDualRealtion}
    \mathcal{H}_{x,y}^{T}D=D\mathcal{H}_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\end{equation}
and 
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
\end{equation}
We perform the proof of duality in three steps: first we will show that matrix \eqref{dualityMatrix} has elements \eqref{dualityElements}; second we will prove the bulk duality \eqref{edgeDualRealtion}; finally we will show the boundary duality \eqref{siteDualRelation}. 
As a preliminary result we note that 
\begin{equation}\label{transpositionPropertyR}
(E_{ba}^{x})^{T}=R_{x}E_{ab}^{x}R_{x}^{-1}\qquad \forall x\in V\co
\end{equation}
that follows immediately from the definition of $E_{ab}$. 
\paragraph{Elements of the duality matrix.}
Consider the matrix $D$ defined in \eqref{dualityMatrix}.
We aim to show that its matrix elements coincide with duality function \eqref{dualityElements}, i.e.
\begin{equation}\label{proofDualityElements}
\langle \bm{n}|D|\bm{\xi}\rangle=D(\bm{n},\bm{\xi})\qquad   \forall \bm{n}\in \Omega,\quad \bm{\xi}\in \widetilde{\Omega}\,.
\end{equation}
Fix an arbitrary site $x\in V$, then we have that 
\begin{align}
	 &\langle n^{x}_{1},\ldots,n_{N}^{x}|\left(d_{x}\otimes \dd\right)\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle\right)\nonumber\\=&\langle n_{1}^{x},\ldots,n_{N}^{x}| (\exp{(E_{N1}^{x}+\ldots+E_{N\,N-1}^{x}}))^{T}R_{x}\otimes\sum_{q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{q_{a}^{u(x)}}\langle q_{1}^{u(x)},\ldots,q_{N}^{u(x)}|\nonumber
	 \\&|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle \otimes |\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle
\end{align}
\todo{something wrong...} {\color{blue} [I have added the vector at site $u(x)$]}
where we used \eqref{transpositionPropertyR}. \\
On one hand, on the extra-site $u(x)$ we have 
\begin{align}
\sum_{q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{q_{a}^{u(x)}}\right)\langle q_{1}^{u(x)},\ldots,q_{N-1}^{u(x)}|\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}\rangle=\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}
\end{align}
\todo{what happens to $q_N$?} {\color{blue} 
where we used the orthogonality relation \eqref{ortho-extraSite}}. 
On the other hand, on the site $x$, we have 
\begin{align}
&\langle n_{1}^{x},\ldots,n_{N}^{x}|(\exp{(E_{N1}^{x}+\ldots+E_{N\,N-1}^{x})})^{T}R_{x}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber\\&= \langle  n_{1}^{x},\ldots,n_{N}^{x}|\left(\sum_{k_{1}=0}^{\infty}\frac{\left\{\left(E_{N1}^{x}\right)^{T}\right\}^{k_{1}}}{k_{1}!}\ldots\sum_{k_{N-1}=0}^{\infty}\frac{\left\{\left(E_{N\,N-1}^{x}\right)^{T}\right\}^{k_{N-1}}}{k_{N-1}!}\sum_{s\in\Omega_{x}}\frac{s_{1}^{x}!\cdots s_{N}^{x}!}{\nu!}|s_{1}^{x},\ldots,s_{N}^{x}\rangle\langle s_{1}^{x},\ldots,s_{N}^{x}|\right)\nonumber
\\&
\hspace{2.cm}\qquad\times |\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber
%\\&=
%\sum_{k_{1}=0}^{n_{1}^{x}}\ldots\sum_{k_{N-1}=0}^{n_{N-1}^{x}}\langle n_{1}^{x}-k_{1},\ldots,n_{N-1}^{x}-k_{N-1},n_{N}^{x}+k_{1}+\ldots+k_{N-1}|
%\\&
%\hspace{2.cm}\prod_{a=1}^{N-1}\left((n_{a}^{x}(n_{a}^{x}-1)\ldots (n_{a}^{x}-k_{a}+1)\right)\frac{1}{k_{1}!,\ldots,k_{N-1}!}\frac{\xi_{1}^{x}!\ldots \xi_{N}^{x}!}{\nu!}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle
\\&=
\sum_{k_{1}=0}^{n_{1}^{x}}\ldots\sum_{k_{N-1}=0}^{n_{N-1}^{x}}\langle n_{1}^{x}-k_{1},\ldots,n_{N-1}^{x}-k_{N-1},\ldots,n_{N}^{x}+k_{1}+\ldots+k_{N-1}|\frac{n_{1}^{x}!\cdots n_{N-1}^{x}!}{(n_{1}^{x}-k_{1})!\cdots(n_{N-1}^{x}-k_{N-1})!}\nonumber
\\& 
\hspace{2.cm}\qquad\times \frac{1}{k_{1}!\cdots k_{N-1}!}\frac{\xi_{1}^{x}!\cdots \xi_{N}^{x}!}{\nu!}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\nonumber
\\&=
\frac{(\nu-\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}
\end{align}
where we used the definitions of the action of $E_{ab}$  and the orthogonality relations \eqref{ortho}, together with the fact that $\xi_{N}^{x}=\nu-\sum_{a=1}^{N-1}\xi_{a}^{x}$. 
Finally, by taking the product over $x\in V$ \eqref{proofDualityElements} is proved.
\begin{flushright}
    $\square$
\end{flushright}

\paragraph{Proof of bulk duality \eqref{edgeDualRealtion}.}To show this relation we need two 'ingredients'. First the existence of a similarity transformation between the Hamiltonian $\mathcal{H}_{x,y}$  and its transposed. As we will show, this similarity transformation is $R_{x}R_{y}$.  Second, the possibility of finding a symmetry for the edge Hamiltonian. Exploiting \eqref{hamiltonianCasimir}, we can take any symmetry of the Casimir and apply the co-product. As we will see, it will be convenient to choose $\sum_{a=1}^{N-1}E_{aN}$.

We first look for the similarity between $\mathcal{H}_{x,y}$ and its transposed. Using \eqref{transpositionPropertyR} we obtain  
\begin{equation}\label{transpositionPropertyH}
    \mathcal{H}_{x,y}^{T}=\left(R_{x}R_{y}\right)\mathcal{H}_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the similarity transformation between $\mathcal{H}_{x,y}$ and its transposed. \\
We now look for a symmetry of $\mathcal{H}_{x,y}$, i.e. a matrix $S_{x,y}$ of the same dimension such that it satisfies 
\begin{equation}
	\mathcal{H}_{x,y}S_{x,y}=S_{x,y}\mathcal{H}_{x,y}
	%\quad \Leftrightarrow \quad [\mathcal{H}_{x,y},S_{x,y}]=0
	\dt
\end{equation}
Using \eqref{hamiltonianCasimir}, we observe that $\mathcal{H}_{x,y}$ is proportional to the coproduct of the second Casimir, up to a diagonal term. Therefore, it is enough to look for a symmetry of $\Delta (C)$. Using the bilinearity of the coproduct operator and the fact that $C$ belongs to the centre of the Lie algebra, it is easy to show that for any matrix $E\in {gl}(N)$ the following holds:
\begin{equation}
 \text{if } \quad	[C,E]=0\qquad  \text{ then }\qquad \left[\Delta (C),\Delta(E)) \right]=0\dt
\end{equation}
We chose 
\begin{equation}
	E=\sum_{a=1}^{N-1}E_{aN}
\end{equation}
and, for a fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
	S_{x,y}=\exp{(\Delta_{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta_{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. As a consequence, this operator $S_{x,y}$ satisfies
\begin{equation}\label{symmetryH}
	\left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. it is a symmetry of $\mathcal{H}_{x,y}$. Exploiting these considerations we may write

\begin{equation}
    \begin{split}
        \mathcal{H}_{x,y}^{T}D&=(R_{x}R_{y})\mathcal{H}_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{u(y)}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{u(y)}\right)\mathcal{H}_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=
        D\mathcal{H}_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $a\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $u(x)$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in {gl}(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the following we need the application of this with $B=\sum_{a=1}^{N-1}E_{aN}$ defined in \eqref{EquationEx} and with $A$ equal to some of the generator $E_{cd}$ of the Lie algebra ${gl}(N)$ with $c,d\in\{1,\ldots,N\}$. Therefore we compute 
\begin{align}
	\left[B,A\right]=\sum_{a=1}^{N-1}\left[E_{aN},E_{cd}\right]=\delta_{cN}\sum_{a=1}^{N-1}E_{ad}-(1-\delta_{dN})E_{cN}
\end{align}
and 
\begin{align}
	\left[B\left[B,A\right]\right]=&\left[\sum_{a=1}^{N-1}E_{aN},\delta_{cN}\sum_{b=1}^{N-1}E_{bd}-(1-\delta_{dN})E_{cN}\right]\nonumber\\=&
	\sum_{a=1}^{N-1}\left(\delta_{cN}\sum_{b=1}^{N-1}\left[E_{aN},E_{bd}\right]-(1-\delta_{dN})\left[E_{aN},E_{cN}\right]\right)\nonumber\\=&
	\sum_{a=1}^{N-1}\left(-\delta_{cN}\delta_{ad}\sum_{b=1}^{N-1}E_{bN}-(1-\delta_{dN})\delta_{cN}E_{aN}\right)\nonumber\\=&
	-\left(\delta_{cN}(1-\delta_{dN})+(1-\delta_{dN})\delta_{cN}\right)\sum_{b=1}^{N-1}E_{bN}\nonumber
	\\=&
	-2\delta_{cN}(1-\delta_{dN})\sum_{b=1}^{N-1}E_{bN}\dt
\end{align}
From the third commutator on we always obtain zero, i.e. $[B,[B,[B,A]]]=0$. All in all we have that 
 \begin{align}\label{HT-BA}
 	\exp{\left(-\sum_{a=1}^{N-1}E_{aN}\right)}E_{cd}\exp{\left(\sum_{a=1}^{N-1}E_{aN}\right)}=E_{cd}-\delta_{cN}\sum_{a=1}^{N-1}E_{ad}+(1-\delta_{dN})E_{cN}-\delta_{cN}(1-\delta_{dN})\sum_{b=1}^{N-1}E_{bN}\dt
 \end{align}

Using \eqref{transpositionPropertyR} we write the transpose of site Hamiltonian \eqref{siteHamiltonian} 
\begin{equation}
    \begin{split}
H_{x}^{T}=\sum_{a,b=1}^{N}\alpha_{a}^{x}\left(E_{ab}^{x}-E_{bb}^{x}\right)^{T}=R_{x}\sum_{a,b=1}^{N}\alpha_{a}^{x}\left(E_{b a}^{x}-E_{bb}^{x}\right)R_{x}^{-1}\dt
    \end{split}
\end{equation}
We multiply both sides by $R_{x}\exp{(E^{x})}$
\begin{equation}\label{intermediateTransposeSite}
    H_{x}^{T}R_{x}\exp{(E^{x})}=R_{x}\sum_{a,b =1}^{N}\alpha_{a}^{x}\left(E_{b a}^{x}-E_{bb}^{x}\right)\exp{(E^{x})}\dt
\end{equation}
%By inserting the identity $I=\exp{(E^{x})}\exp{(-E^{x})}$ and 
By using \eqref{HT-BA} and the fact that 
%the first Casimir operator $C_{1}=
$\sum_{a,b=1}^{N}E_{ab}$ is central for the algebra we have

\begin{equation}\label{HadTransfBoundary}
\exp{(-E^{x})}\sum_{a=1}^{N}\sum_{b=1}^{N}\alpha_{a}^{x}\left(E_{b a}^{x}-E_{bb}^{x}\right)\exp{(E^{x})}=	|\alpha^{x}|\sum_{a=1}^{N-1}\left(\rho_{a}^{x}E_{Na}^{x}-E_{aa}^{x}\right)\dt
\end{equation}
Thus, we rewrite \eqref{intermediateTransposeSite} as
\begin{equation}\label{siteHadamardI}
H_{x}^{T}R_{x}\exp{(E^{x})}=R_{x}\exp{(E^{x})}|\alpha^{x}|\sum_{a=1}^{N-1}\left(\rho_{a}^{x}E_{Na}^{x}-E_{aa}^{x}\right)\dt
\end{equation}
Taking the tensor product of both sides of \eqref{siteHadamardI}% with 
%\begin{equation}
%\langle \rho|_{u(x)}=\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}%\right)^{\xi_{a}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|_{u(x)}
%\end{equation}
%{\color{blue}Think about replacing $\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|$ with $\langle \xi_{1},\ldots,\xi_{N}|_{u(x)}$}\\
we obtain 
\begin{equation}\label{siteHadamardII}
    \begin{split}
    	&H_{x}^{T}d_{x}\otimes\left(\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|\right)\\=&
 d_{x}\,
|\alpha^{x}|\sum_{a=1}^{N-1}\left(\rho_{a}^{x}E_{Na}^{x}-E_{aa}^{x}\right)\otimes \left(\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|\right)
    \end{split}
\end{equation}
\todo{somehow this now has extra index for the space and everywhere else not.... please fix} {\color{blue} We have defined the bosonic operator with the extra-site, see under equation \eqref{bosonic_2}}
Recalling the action of the bosonic creation operator acting at site $u(x)$ and on the species $a\in \{1,\ldots,N-1\}$ we have that 
\begin{equation}\label{bosonicKX}
    \langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)}+1,\ldots,\xi_{N-1}^{u(x)}|(\mathbf{a}^{+})^{u(x)}_{a}=  \langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|\dt
\end{equation}
Using the above equation we rewrite the right hand side term of \eqref{siteHadamardII} with Lie generator $E_{Na}^{x}$ as 
\begin{equation}\label{extra-site-trick}
    \begin{split}
&\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{b=1}^{N-1}\left(\rho_{b}^{x}\right)^{\xi_{b}^{u(x)}}\right)\rho_{a}^{x}\langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|
\\=&
\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{b=1\,:\,b\neq a}^{N-1}\left(\rho_{b}^{x}\right)^{\xi_{b}^{u(x)}}\right)\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}+1}\langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)}+1,\ldots,\xi_{N-1}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}
\\=&
%\sum_{\xi_{1}^{u(x)},\ldots,\xi_{a-1},\xi_{a+1},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\sum_{\xi_{a}^{u(x)}=1}^{\infty}\left(\prod_{b=1}^{N-1}\left(\rho_{b}^{x}\right)^{\xi_{b}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}\\=&
\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{b=1}^{N-1}\left(\rho_{b}^{x}\right)^{\xi_{b}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}
    \end{split}
\end{equation}
where, in the up to last equality, we performed a change of summation variable and we used the fact that  
\begin{equation}
	\begin{split}
&\sum_{\xi_{a}^{u(x)}=0}^{\infty}(\rho_{a}^{x})^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}
%=&\sum_{\xi_{a}^{u(x)}=1}^{\infty}(\rho_{a}^{x})^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)},\ldots,\xi_{N}^-1{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}+\langle \xi_{1}^{u(x)},\ldots,0,\ldots,\xi_{N-1}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}\\
=\sum_{\xi_{a}^{u(x)}=1}^{\infty}(\rho_{a}^{x})^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{a}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}\dt
\end{split}
\end{equation}
%since
%\begin{equation}
%\langle \xi_{1}^{u(x)},\ldots,0,\ldots,\xi_{N-1}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}=0\dt
%\end{equation}
%This is a consequence of 
%\begin{equation}
%\langle \xi_{1}^{u(x)},\ldots,0,\ldots,\xi_{N-1}^{u(x)}|($\mathbf{a}$^{+})_{a}^{u(x)}=0
%\end{equation}
%that follows from \eqref{bosonicKX}.
Therefore, inserting \eqref{extra-site-trick} in \eqref{siteHadamardII} we obtain 
\begin{equation}
    \begin{split}
H_{x}^{T}d_{x}\otimes \dd&=
     H_{x}^{T}d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|
\\&=
\left\{d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|\right\}\,|\alpha^{x}|\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{u(x)}E_{Na}^{x}-E_{aa}^{x}\right)   
%\\&=
%\left\{d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}|\right\}\widetilde{H}_{x}
\\&=
\left(d_{x}\otimes \dd\right)\widetilde{H}_{x}\dt
    \end{split}
\end{equation}
Since the duality matrix \eqref{dualityMatrix} is product over sites, the above equality implies \eqref{siteDualRelation}. 
\begin{flushright}
$\square$
\end{flushright}

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Section~\ref{subsection-description-process-LINE} with the hard-core constrain, i.e. $\nu=1$. 
We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. 
We also recall the notation for the occupation variable which is called $\bm{\tau}=(\tau_1,\ldots,\tau_L)$ with $\tau_x\in\{1,\ldots,N\}$.
The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of ${gl}(N)$. The duality result stated in Section~\ref{statementDualitySubsection} is adapted to the present situation as follows. 
The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{ab}^T= e_{ba}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{\text{right}}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{\text{left}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{\text{right}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Section~\ref{sectionIntegrabiliy}.\\

%In this notation, the elements of the duality matrix reads
%\begin{equation}\label{ElementsDualityMatrixChain_1}
%	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
%\\
Using duality, one can compute the  $m$-point correlations between non-empty particles in terms of the absorption probabilities of $m$ dual particles. Therefore, to determine the non-equilibrium steady state correlations of the integrable multispecies stirring process, it is enough to compute these absorption probabilities. In the following we denote by $\mu$ the non-equilibrium steady state distribution and we call $\bm{Y} = (Y_1,\ldots,Y_L)$ the random vector with law $\mu$. 
Furthermore we write $\mathbb{E}[\cdot]$ for the expectation with respect to $\mu$.%For arbitrary $m\in \mathbb{N}$ such that $m<L$, we fix $a_{k}\in\{1,\ldots,N-1\}$ and $x_{k}\in\{1,\ldots,L\}$ such that $x_{k}<x_{k+1}$ for every $k\in\{1,\ldots,m\}$. 
\begin{corollary}[Correlations via duality]\label{Corolollary-ABS_Corr-abstract}
Let $m\in \{1,\ldots,L\}$ and consider $m$ sites $1 \le x_{1} < x_2 < \ldots < x_m \le L $ and
$m$ species $a_{k}\in\{1,\ldots,N-1\}$  with $k=1,2,\ldots, m$. Then, the $m$-point correlations in the non-equilibrium steady state read
\begin{equation}\label{NESS-expectation-ABS}
\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=\sum_{t_{1}=0}^{1}\ldots \sum_{t_{m}=0}^{1}\left[\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\right]\mathcal{P}_{x_1,\ldots,x_m}(t_{1},\ldots,t_{m})
\end{equation}
where
\begin{equation}\label{absProbabilitiesIntegrable}
	\mathcal{P}_{x_1,\ldots,x_m}(t_{1},\ldots,t_{m}):=\mathbb{P}\left(\bm{\xi}(\infty)=\sum_{k=1}^{m}\left(t_{k}\delta^{0}_{a_k}+(1-t_{k})\delta^{L+1}_{a_k}\right)\Big| \bm{\xi}(0)=\bm{\xi}\right)
\end{equation}
where the initial dual configuration is $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$, meaning that 
\begin{equation}
\label{xsi-init}
	\begin{cases}
		\xi_{a}^{x}=1\qquad \text{if}\;x=x_{k}\;\text{and}\; a=a_{k}\\
		\xi_{a}^{x}=0\qquad \text{otherwise}
	\end{cases}\dt
\end{equation}
In \eqref{absProbabilitiesIntegrable} the variable $t_k$ is $1$ (resp. $0$) when the dual particle of species $a_k$ initially positioned at site $x_{k}$ is absorbed at at the extra site $\{0\}$ (resp. $\{L+1\}$).
\end{corollary}
\textbf{Proof of Corollary~\ref{Corolollary-ABS_Corr-abstract}}:
For any $\bm{\tau}\in \Omega^{'}$ the duality matrix defined in \eqref{ElementsDualityMatrixChain_1} evaluated on the dual configuration $\bm{\xi}\in\widetilde{\Omega}$ given in \eqref{xsi-init} read
\begin{equation}
	D(\bm{\tau},\bm{\xi})%=\prod_{k=1}^{m}\mathbbm{1}_{\{\delta_{\tau_{x_{k}},a_{k}}=1\}}
	=\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\dt
\end{equation}
Therefore, by ergodicity and duality  we have 
\begin{equation}\label{CorrelationByABS-prob}
	\begin{split}
\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]&=\lim_{t\to \infty}\mathbb{E}_{\bm{\tau}}\left[D(\bm{\tau}(t),\bm{{\xi}})\right]=\lim_{t\to \infty}\mathbb{E}_{\bm{\xi}}\left[D(\bm{\tau},\bm{{\xi}}(t))\right]
	\\&=\sum_{t_{1}=0}^{1}\ldots \sum_{t_{m}=0}^{1}\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\mathcal{P}_{x_{1},\ldots,x_{m}}(t_{1},\ldots,t_{m})\dt
\end{split}
\end{equation}
\begin{flushright}
	$\square$
\end{flushright} 
    \begin{corollary}[Non-equilibrium steady state]\label{Corollary-NESS-ABS}
For any configuration $\bm{\tau}\in \Omega^{'}$  the probability mass function of the non-equilibrium steady states is given by 
\begin{equation}\label{implicit-mass-function}
	\mu(\bm{\tau})=\sum_{b=0}^{L}(-1)^{b}\sum_{1\leq q_{1}<\ldots<q_{b}\leq L}\left(\prod_{k=1}^{b}\delta_{\tau_{q_{k}},N}\right)\sum_{a_{q_{1}},\ldots,a_{q_{b}}=1}^{N-1}{\color{red}
	\mathbb{E}\left[\left( \prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_{x}=\tau_{x}\}}\right) \left(\prod_{k=1}^{b}\mathbbm{1}_{\{Y_{q_{k}}=a_{q_{k}}\}}\right)\right]}\,.
\end{equation}
{\color{blue}where the expectation is taken with respect to the non-equilibrium steady state distribution, then it is given in \eqref{exactCorrelations}.}
\todo{Explain this expectation...}
\end{corollary}
\textbf{Proof of Corollary~\ref{Corollary-NESS-ABS}:}
Given $\bm{\tau}\in \Omega^{'}$,  we have 
\begin{align}
	\mu\left(\bm{\tau}\right)=&\mathbb{E}\left[\prod_{x=1}^{L}\mathbbm{1}_{\{Y_{x}=\tau_{x}\}}\right]
	=\mathbb{E}\left[\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_{x}=\tau_{x}\}}\right)\left(\prod_{x=1\,:\,\tau_{x}=N}^{L}\mathbbm{1}_{\{Y_{x}=\tau_x\}}\right)\right]\nonumber\\
	=&\mathbb{E}\left[\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_{x}=\tau_{x}\}}\right)\left(\prod_{x=1\,:\,\tau_{x}=N}^{L}\left(1-\sum_{a=1}^{N-1}\mathbbm{1}_{\{Y_{x}=a\}}\right)\right)\right]
	%\\=&\sum_{\bm{\sigma}\in \Omega^{'}}\prod_{x=1}^{L}\mathbbm{1}_{\{Y_x=\tau_{x}\}}\langle \bm{\sigma}|\Psi\rangle\nonumber\\
	%	=&\sum_{\bm{\sigma}\in \Omega^{'}}\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_x=\tau_{x}\}}\right)\left(\prod_{x=1\,:\,\tau_{x}=N}^{L}\mathbbm{1}_{\{Y_x=N\}}\right)\langle\bm{\sigma}|\Psi\rangle\nonumber\\
	%	=&\sum_{\bm{\sigma}\in \Omega^{'}}\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_x=\tau_{x}\}}\right)\prod_{x=1\,:\,\tau_{x}=N}^{L}\left(1-\sum_{i_{x}=1}^{N-1}\mathbbm{1}_{\{Y_x=i_{x}\}}\right)\langle\bm{\sigma}|\Psi\rangle
	%=&\sum_{\bm{\sigma}\in \Omega^{'}}\langle \bm{\sigma}|D|\bm{\xi}\rangle\langle \bm{\sigma}|\Psi\rangle
\end{align}
% {\color{blue}Is it not an identity? Take for example 3 sites: $\tau_{1}=\tau_{2}=1$ and $\tau_{3}=N$. Then we have that 
% $$
% \prod_{x=1}^{3}\mathbbm{1}_{\{Y_{x}=\tau_{x}\}}=\mathbbm{1}_{\{Y_{1}=\tau_{1}\}}\mathbbm{1}_{\{Y_{2}=\tau_{2}\}}\mathbbm{1}_{\{Y_{3}=\tau_{3}\}}=\left(\mathbbm{1}_{\{Y_{1}=1\}}\mathbbm{1}_{\{Y_{2}=1\}}\right)\left(\mathbbm{1}_{\{Y_{3}=N\}}\right)=\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{3}\mathbbm{1}_{\{Y_{x}=\tau_{x}\}}\right)\left(\prod_{x=1\,:\,\tau_{x}= N}^{3}\mathbbm{1}_{\{Y_{x}=\tau_{x}\}}\right)
% $$} 
\todo{removed subindex $a_x$ here}
where in the last equality we exploited the fact that 
\begin{equation}\label{prova}			\mathbbm{1}_{\{Y_{x}=N\}}=1-\sum_{a=1}^{N-1}\mathbbm{1}_{\{Y_{x}=a\}}\dt
\end{equation}
% {\color{blue}
% Here I mean 
% \begin{equation}
% 	\mathbbm{1}_{\{Y_{x}=a\}}=\begin{cases}
% 		1\qquad \text{if}\qquad  Y_{x}=a\\
% 		0\qquad \text{if}\qquad Y_{x}\neq a
% 	\end{cases}
% \end{equation}
% Therefore, \begin{equation}
% 	\mathbbm{1}_{\{Y_{x}=N\}}=\begin{cases}
% 		1\qquad \text{if}\qquad  Y_{x}=N\\
% 		0\qquad \text{if}\qquad Y_{x}\neq N
% 	\end{cases}
% \end{equation}
% and 
% \begin{equation}
% 1-\sum_{a_{x}=1}^{N-1}\mathbbm{1}_{\{Y_{x}=a_{x}\}}=\begin{cases}
% 	1\qquad \text{if}\qquad  Y_{x}=N\\
% 	0\qquad \text{if}\qquad Y_{x}\neq N
% \end{cases}
% \end{equation}
% therefore \eqref{prova} follows.
% }
The configuration $\bm{\tau}$ has $\ell=\sum_{x=1}^{L}(1-\delta_{\tau_{x},N})$ occupied sites and then the remaining $L-\ell$ sites are empty. % We introduce an index $b\in \{0,\ldots,L-\ell\}$. 
Therefore, we may write
\begin{align}\label{sum}
&\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_x=\tau_{x}\}}\right)\left(\prod_{x=1\,:\,\tau_{x}=N}^{L}\left(1-\sum_{a_{x}=1}^{N-1}\mathbbm{1}_{\{Y_x=a_{x}\}}\right)\right)\nonumber\\
=&
\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_x=\tau_{x}\}}+\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_x=\tau_{x}\}}\right)(-1)\left\{\sum_{q_{1}=1}^{L}\delta_{\tau_{q_{1}},N}\sum_{a_{q_{1}}=1}^{N-1}\mathbbm{1}_{\{Y_{q_{1}}=a_{q_{1}}\}}\right\}+\nonumber\\
+&\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_x=\tau_{x}\}}\right)(-1)^{2}\left\{\sum_{1\leq q_{1}<q_{2}\leq L}\delta_{\tau_{q_{1}},N}\delta_{\tau_{q_{2}},N}\sum_{a_{q_{1}},a_{q_{2}}=1}^{N-1}\mathbbm{1}_{\{Y_{q_{1}}=a_{q_{1}}\}}\mathbbm{1}_{\{Y_{q_{2}}=a_{q_{2}}\}}\right\}+\ldots+\nonumber\\
+&\left(\prod_{x=1\,:\,\tau_{x}\neq N}^{L}\mathbbm{1}_{\{Y_x=\tau_{x}\}}\right)(-1)^{L-\ell}\sum_{1\leq q_{1}<\ldots<q_{L-\ell}\leq L}\left(\prod_{k=1}^{L-\ell}\delta_{\tau_{q_{k}},N}\right)\sum_{a_{q_{1}},\ldots,a_{q_{L-\ell}}=1}^{N-1}\prod_{k=1}^{L-\ell}\mathbbm{1}_{\{Y_{q_{k}}=a_{q_{k}}\}}\dt
\end{align} 
\todo{indices are strange. Say what happens here}
Therefore, using the linearity of the expectation and introducing the index $b\in \{0,\ldots,L\}$ {\color{blue} we rewrite the sum in equation \eqref{sum} as} \eqref{implicit-mass-function}.{\color{blue} Indeed, when $b>L-\ell$ we have vanishing terms in the sum \eqref{implicit-mass-function}.}
\begin{flushright}
	$\square$
\end{flushright}
\begin{remark}\label{Remark-extension-graph-nu}
	The results in Corollary~\ref{Corolollary-ABS_Corr-abstract} and in Corollary~\ref{Corollary-NESS-ABS}  can be generalized to arbitrary graphs $G=(V,\mathcal{E})$ and for higher maximal occupancies $\nu >1$.
	However the computation of absorption probabilities of the dual process is possible only in the set-up of the integrable chain with $\nu =1$ and two reservoirs at the boundaries.
	Thus we chose to present them in such a context. For the chain with higher maximal occupancies $\nu >1$ see Section~\ref{Subsection-ss-nonI}.
\end{remark}

\section{Correlations and steady state  for the integrable chain}\label{sectionIntegrabiliy}
In this section we combine duality, the $gl(N)$ invariance and the matrix product ansatz (MPA) to charaterize the non-equilibrium steady state for the integrable version of the multispecies stirring process, defined in Section~\ref{integrableChain-duality}. 
As it will turn out, the combination of the $gl(N)$ invariance  and the MPA allows to determine a closed formula for the absorption probabilities and thus (using Corollary~\ref{Corolollary-ABS_Corr-abstract}) for the correlations
and (using Corollary~\ref{Corollary-NESS-ABS}) for the probability mass-function of the non-equilibrium steady state.
We start by recalling the formulation of the MPA  and then we explain our strategy.

\subsection{Matrix product ansatz for the non-equilibrium steady state}
The matrix product ansatz for the multispecies stirring process has been formulated in \cite{vanicat2017exact}, here we briefly recall the main steps.
 Denoting by $|\Psi(t)\rangle$ the column vector that encodes the probability distribution of the chain with Hamiltonian \eqref{hamiltonian} at time $t\geq 0$, its evolution equation is given by 
 the master equation
\begin{equation}
    \frac{d|{\Psi}(t)\rangle}{dt}=H|{\Psi}(t)\rangle\dt
\end{equation}
This Markov chain is irreducible and positive recurrent, therefore there exists a unique stationary measure, that will be reached when time goes to infinity, regardless of the initial configuration. 
We denote by $|\Psi\rangle$ the column vector that gives the stationary distribution (non-equilibrium steady state). This vector is the right eigenvector with vanishing eigenvalue of $H$, i.e. it solves 
\begin{equation}\label{definition-SteadyStateH}
	H|\Psi\rangle =0\dt
\end{equation}
The MPA states the following
\begin{equation}
	|\Psi\rangle=\frac{1}{Z_{L}}\langle W|\begin{pmatrix}
		X_{1}\\
		\vdots\\
		X_{N}
	\end{pmatrix}\otimes \ldots\otimes \begin{pmatrix}
		X_{1}\\
		\vdots\\
		X_{N}
	\end{pmatrix}|V\rangle
\end{equation}
with the normalization 
\begin{equation}
	Z_{L}=\langle W|(X_{1}+\ldots +X_{N})^{L}|V\rangle\dt
\end{equation}
Here the operators $(X_{a})_{a\in \{1,\ldots,N\}}$ fulfil the commutators
\begin{equation}\label{bulk}
	\left[X_{a},X_{b}\right]=(\alpha_{a}-\beta_{a})X_{b}-(\alpha_{b}-\beta_{b})X_{a}%\qquad\forall a,b=1,\ldots,N
\end{equation}
and their action on the boundary vectors are
\begin{equation}\label{leftBoundary}
	\langle W|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=(\alpha_{a}-\beta_{a})\langle W|%\qquad\forall a=1,\ldots,N\co
\end{equation}
\begin{equation}\label{rightBoundary}
	\left(\beta_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)|V\rangle=-(\alpha_{a}-\beta_{a})|V\rangle%\qquad\forall a=1,\ldots,N\dt
\end{equation}
Without loss of generality, we further assume that $\langle W|V\rangle=1$. 
\begin{remark} Only $N-1$ of the $N$ equations \eqref{leftBoundary}  are independent. This can be seen by summing them over the species $a\in \{1,\ldots,N\}$, i.e.
	\begin{equation}
		\sum_{a=1}^{N}	\langle W|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=0,%\sum_{a=1}^{N}(\alpha_{a}-\beta_{a})\langle W|=0
	\end{equation}
	and using  \eqref{ratesConditions}.
 Similarly for the right boundary  \eqref{rightBoundary} there are only  $N-1$ independent equations. \end{remark} 
The MPA gives an abstract form of the non-equilibrium steady state in terms of an algebra of operators. However, the computation of  correlations are in general involved.

\subsection{Strategy}
 In the next Section~\ref{subsection-exact} we will show that the DEHP albegra in \eqref{bulk}-\eqref{rightBoundary} can be simplified substantially using the ${gl}(N)$ invariance of the bulk. In fact, such simplification can always be achieved if 
  there exists an absorbing dual process as established in Section~\ref{sectionDuality}

 The idea is summarized in the following scheme. 
 We define a sequence of similarity transformation such that 
 \begin{equation}
 	H\;\xleftrightarrow{S_{1}}\;H^{'}\;\xleftrightarrow{S_{2}}\;H^{''}
 \end{equation} 
 where $H^{'}= S_{1} H S_{1}^{-1} $ has both boundary terms in a triangular form, and $H^{''} = S_{2}H' S_{2}^{-1}$ has the left boundary in a triangular form and the right boundary is diagonal. Using these transformations the commutators \eqref{bulk} and the action on the boundary vectors \eqref{leftBoundary}, \eqref{rightBoundary} simplify significantly. More precisely for the components of $\tilde X=S_2S_1 X$ we obtain the bulk relations
\begin{equation}\label{commutationsBula2}
	\left[\Xt_{a},\Xt_{N}\right]=(\alpha_{a}-\beta_{a})\Xt_{N}\dt %\forall a\in \{1,\ldots,N-1\}\dt
\end{equation}
for $a=1,\ldots,N-1$. 
At the boundaries 
\begin{equation}\label{commLEFT2}
	\langle W|\left((\alpha_{a}-\beta_{a})\Xt_{N}-\Xt_{a}\right)=(\alpha_{a}-\beta_{a})\langle W|%\qquad\forall a=1,\ldots,N-1
	\co \qquad 
	\Xt_{a} |V\rangle= (\alpha_{a}-\beta_{a})|V\rangle%\qquad\forall a=1,\ldots,N-1
	\,.
\end{equation} 
where again we have  $a=1,\ldots,N-1$.   As a consequence, the ground state $|\Psi^{''}\rangle$ of $H^{''}$ can be written exactly in closed-form. Finally, reversing the transformations we retrieve $|\Psi\rangle$, i.e. the vector whose components are the probabilities of a certain configuration of the process in the non-equilibrium steady state.
 The first transformation $S_{1}$ is closely related to the duality matrix and we will see that $H^{'}$ is $\widetilde{H}^{T}$, up to the extra-site term described by the bosonic creation operator in \eqref{siteDualHamiltonian}. Therefore, $H^{'}$ is not stochastic, however it turns out that the components of its eigenvector with zero eigenvalue $|\Psi^{'}\rangle$ are the correlations in the non-equilibrium steady state. 
 %The second transformation $S_{2}$ diagonalizes the right boundary. Finally, the composition  $S_{2} S_{1}$ allows to simplify the commutators of the MPA algebra. 
\subsection{Correlations in the non-equilibrium steady state}\label{subsection-exact}
In this section we write a formula for the stationary non-equilibrium steady state correlations between $m$- points of the chain.
\begin{theorem}[Correlations in the non-equilibrium steady state]\label{thm-correlations}
%Consider $m\in \{1,\ldots,L\}$ points with coordinates given by $(x_{1},\ldots,x_{m})$, such that $x_{k}<x_{k+1}$ $\forall k\in \{1,\ldots,m-1\}$
%and consider $m$ species $a_{k}\in \{1,\ldots,N-1\}$. 
Let $m\in \{1,\ldots,L\}$ and consider $m$ sites $1 \le x_{1} < x_2 < \ldots < x_m \le L $ and
$m$ species $a_{k}\in\{1,\ldots,N-1\}$  with $k=1,2,\ldots, m$. Then the $m$-point correlations with respect to the non-equilibrium steady state measure are given by 
\begin{equation}\label{exactCorrelations}
\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]\\=\sum_{t_{1},\ldots,t_{m}=0}^{1}\left(\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\right)\mathcal{P}_{x_{1},\ldots,x_{m}}(t_{1},\ldots,t_{m})
\end{equation}
where
\begin{equation}\label{abs-probabilities}
	\mathcal{P}_{x_{1},\ldots,x_{m}}(t_{1},\ldots,t_{m})=\sum_{c_{1}=t_{1}}^{1}\ldots\sum_{c_{m}=t_{m}}^{1}f(c_{1},\ldots,c_{m})\prod_{j=1}^{m}(-1)^{c_{j}-t_{j}}g_{j}(x_{j},c_{j},\ldots,c_{m})
\end{equation}
with
\begin{equation}\label{powerCoeffNOspec}
	f(c_{1},\ldots,c_{m})=\frac{(L+1-\sum_{a=1}^{m}c_{a})!}{(L+1)!}
\end{equation}
and
\begin{equation}\label{powerCoeffSpecies}
	g_{j}(x_{j},c_{j},\ldots,c_{m})=\left(L+2-x_{j}-\sum_{k=j}^{m}c_{k}\right)^{c_{j}}\dt
\end{equation}
\end{theorem}

\subsubsection{Examples}
We give examples  of correlations for $m=1,2,3$ applying formula \eqref{exactCorrelations}.
\paragraph{One-point correlations}
We consider the average with respect to $\mu$ of the occupation variable of the species $a_{1}\in \{1,\ldots,N-1\}$ at coordinate $x_{1}\in \{1,\ldots,L\}$. %This reads
%\begin{equation}
%\mathbb{E}_{\mu}\left[\mathbbm{1}_{\{\tau_{x_{1}}=a_{1}\}}\right]=\sum_{t_{1}=0}^{1}\alpha_{a_{1}}^{t_{1}}\beta_{a_{1}}^{1-t_{1}}\sum_{c_{1}=t_{1}}^{1}f(L,c_{1})g_{a_{1}}(L,x_{1},c_{1})\dt
%\end{equation}
Using \eqref{abs-probabilities} we obtain the absorption probabilities 
\begin{equation}
	\mathcal{P}_{x_1}(0)=1-\frac{(L+1-x_{1})}{(L+1)}\qquad \mathcal{P}_{x_1}(1)=\frac{(L+1-x_{1})}{(L+1)}
\end{equation} 
where $\mathcal{P}_{x_1}(1)$ is the probability that a single random walk started at $x_1$ is absorbed in $\{0\}$. Then, using \eqref{exactCorrelations}, we have
\begin{equation}\label{one-pts-corr}
		\langle\rho_{a_{1}}^{x_{1}}\rangle= \mathbb{E}\left[\mathbbm{1}_{\{\tau_{x_{1}}=a_{1}\}}\right]=\frac{(L+1-x_{1})}{(L+1)}\alpha_{a_{1}}+\frac{x_{1}}{(L+1)}\beta_{a_{1}}\dt
\end{equation}
\paragraph{Two-point correlations}
We consider the average with respect to $\mu$ of the occupation variable of the species  $a_{1},a_{2}\in \{1,\ldots,N-1\}$ at coordinates $x_{1},x_{2}\in \{1,\ldots,L\}$ such that $x_{1}< x_{2}$. 
%\begin{equation}
%	\begin{split}
%		 \mathbb{E}_{\mu}\left[\mathbbm{1}_{\{\tau_{x_{1}}=a_{1}\}}\mathbbm{1}_{\{\tau_{x_{2}}=a_{2}\}}\right]= \sum_{t_{1}=0}^{1}\sum_{t_{2}=0}^{1}\left(\prod_{k=1}^{2}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\right)\sum_{c_{1}=t_{1}}^{1}\sum_{c_{2}=t_{2}}^{1}f(L,c_{1},c_{2})\prod_{j=1}^{2}g_{k}(L,x_{j},c_{j},c_{2})\dt
%	\end{split}
%\end{equation}
Using \eqref{abs-probabilities} we obtain the absorption probabilities
\begin{equation}
	\begin{split}
	&\mathcal{P}_{x_{1},x_{2}}(0,0)=\frac{x_{1}(x_{2}-1)}{L(L+1)}\\
	&\mathcal{P}_{x_{1},x_{2}}(0,1)=\frac{x_{1}(L+1-x_{2})}{L(L+1)}
\end{split}\quad \begin{split}
	&\mathcal{P}_{x_{1},x_{2}}(1,0)=\frac{x_{2}(L+1-x_{1})}{L(L+1)}\\
	&\mathcal{P}_{x_{1},x_{2}}(1,1)=\frac{(L-x_{1})(L+1-x_{2})}{L(L+1)}
	\end{split}\dt
\end{equation}
%where $\bm{\xi}=\delta_{a_{1}}^{x_{1}}+\delta_{a_{2}}^{x_{2}}$. 
%Then, by using \eqref{exactCorrelations} we have 
%\begin{equation}
%\begin{split}
%	\mathbb{E}_{\mu}\left[\mathbbm{1}_{\{\tau_{x_{1}}=a_{1}\}}\mathbbm{1}_{\{\tau_{x_{2}}=a_{2}\}}\right]=&\frac{x_{1}(x_{2}-1)}{L(L+1)}\beta_{a_{1}}\beta_{a_{2}}+\frac{x_{1}(L+1-x_{2})}{L(L+1)}\beta_{a_{1}}\alpha_{a_{2}}\\
%	+&\frac{x_{2}(L+1-x_{1})}{L(L+1)}\alpha_{a_{1}}\beta_{a_{2}}+\frac{(L-x_{1})(L+1-x_{2})}{L(L+1)}\alpha_{a_{1}}\alpha_{a_{2}}
%\end{split}
%\end{equation}
%\begin{equation}
%	\begin{split}
%	 \mathbb{E}_{\mu}\left[\mathbbm{1}_{\{\tau_{x_{1}}=a_{1}\}}\mathbbm{1}_{\{\tau_{x_{2}}=a_{2}\}}\right]&=\beta_{a_{1}}\beta_{a_{2}}+\frac{(L+1-y_{1})}{L+1}(\alpha_{a_{1}}-\beta_{a_{1}})\beta_{a_{2}}+\frac{(L+1-x_{2})}{L+1}(\alpha_{a_{2}}-\beta_{a_{2}})\beta_{a_{1}}\\&+\frac{(L-x_{1})(L+1-x_{2})}{L(L+1)}(\alpha_{a_{1}}-\beta_{a_{1}})(\alpha_{a_{2}}-\beta_{a_{2}})\dt
%	\end{split}
%\end{equation}
Therefore, we compute the second cumulant, i.e. the two-point connected correlation
\begin{equation}\label{two-pts-corr}
	\begin{split}
		 \mathbb{E}\left[\left(\mathbbm{1}_{\{\tau_{x_{1}}=a_{1}\}}-\langle\rho_{a_{1}}^{x_{1}}\rangle \right)\left(\mathbbm{1}_{\{\tau_{x_{2}}=a_{2}\}}-\langle\rho_{a_{2}}^{x_{2}}\rangle \right)\right]=-\frac{x_{1}(L-x_{2}+1)}{L(L+1)^{2}}\left(\alpha_{a_{1}}-\beta_{a_{1}}\right)\left(\alpha_{a_{2}}-\beta_{a_{2}}\right)\dt
	\end{split}
\end{equation}
\paragraph{Three-point correlations}
We consider the average with respect to $\mu$ of the occupation variable of the species  $a_{1},a_{2},a_{3}\in \{1,\ldots,N-1\}$ at coordinates $x_{1},x_{2},x_{3}\in \{1,\ldots,L\}$ such that $x_{1}<x_{2}<x_{3}$. Using \eqref{abs-probabilities} we compute the absorption probabilities 
\begin{equation}
	\begin{split}
	&\mathcal{P}_{x_{1},x_{2},x_{3}}(0,0,0)=\frac{x_{1}(x_{2}-1)(x_{3}-2)}{L(L^{2}-1)}\\
 &\mathcal{P}_{x_{1},x_{2},x_{3}}(0,1,1)=\frac{x_1 (L-x_2) (L-x_3+1)}{L \left(L^2-1\right)}\\
	&\mathcal{P}_{x_{1},x_{2},x_{3}}(0,1,0)=\frac{x_1 (L(1-x_3)+x_2 (x_3-2)+1)}{L(L^2-1)}\\
	&\mathcal{P}_{x_{1},x_{2},x_{3}}(0,0,1)=\frac{x_1 (x_2-1) (L-x_3+1)}{L \left(L^2-1\right)}
	\end{split}\quad \begin{split}
	&\mathcal{P}_{x_{1},x_{2},x_{3}}(1,1,0)=\frac{(L-1) x_2 (x_3-1)-x_1 (x_2-1)
		(x_3-2)}{L \left(L^2-1\right)}\\
	&\mathcal{P}_{x_{1},x_{2},x_{3}}(1,0,1)=\frac{(L-x_3+1) (x_2 (L-x_1-1)+x_1)}{L
		\left(L^2-1\right)}\\
	&\mathcal{P}_{x_{1},x_{2},x_{3}}(1,0,0)=\frac{ x_{2} (x_3-1)(L-1)-x_1 (x_2-1)
	(x_3-2)}{L \left(L^2-1\right)}\\
	&\mathcal{P}_{x_{1},x_{2},x_{3}}(1,1,1)=\frac{(L-x_{1}-1)(L-x_{2})(L-x_{3}+1)}{L(L^{2}-1)}
\end{split}
\end{equation}

Therefore, we compute the third cumulant, i.e. the three-point connected correlation as 
\begin{align}
	&\mathbb{E}\left[\left(\mathbbm{1}_{\{Y_{x_{1}}=a_{1}\}}-\langle\rho_{a_{1}}^{x_{1}}\rangle \right)\left(\mathbbm{1}_{\{\tau_{x_{2}}=a_{2}\}}-\langle\rho_{a_{2}}^{x_{2}}\rangle \right)\left(\mathbbm{1}_{\{\tau_{x_{3}}=a_{3}\}}-\langle\rho_{a_{3}}^{x_{3}}\rangle \right)\right]\nonumber\\
	=&-\frac{2x_{1}(L+1-2x_{2})(L+1-x_{3})}{(L+1)^{3}(L-1)L}(\alpha_{a_{1}}-\beta_{a_{1}})(\alpha_{a_{2}}-\beta_{a_{2}})(\alpha_{a_{3}}-\beta_{a_{3}})\dt
\end{align}
\begin{remark} The first and second cumulants, computed in \eqref{one-pts-corr} and \eqref{two-pts-corr} respectively, match with those found in Section 4.3 of \cite{vanicat2017exact}.
	\end{remark}

\subsection{Proof of Theorem~\ref{thm-correlations}}
The proof of formula \eqref{exactCorrelations} is split into the following steps:
\begin{itemize}
	\item In Section~\ref{subsectionSTransf}, we introduce two similarity transformations $S_1$ and  $S_2$. 
	 The first turns the Hamiltonian $H$ into $H^{'}= S_{1} H S_{1}^{-1} $ with boundary terms in a triangular form;
	 the second  turns the Hamiltonian $H'$ into $H^{''} = S_{2}H' S_{2}^{-1}$ having the left boundary in a triangular form and the right boundary is diagonal.
	Associated to $H^{'}$ and $H^{''}$ there are two ground states denoted by $|\Psi^{'}\rangle$ and $|\Psi^{''}\rangle$ respectively. 
	\item In Section~\ref{subsectionSSHsec} we apply the MPA to $H^{''}$. Here the commutation relations defining the matrix algebra are simpler. The explicit expression for $|\Psi^{''}\rangle$ is determined (see \eqref{ResulsBasis}).
	\item In subsection~\ref{subsectionSSdual}, we invert the similarity transformation $S_2$ to recover  the ground state $|\Psi^{'}\rangle$ from the explicit expression of $|\Psi^{''}\rangle$  (see \eqref{ABS}).
	\item In Section~\ref{correlation-section} we show, using duality, that the correlations are in turn the components of the vector $|\Psi^{'}\rangle$. By exploiting a binomial formula we rewrite the correlations in terms
	as a polynomial in the left and right boundary densities with coefficients given by the absorption probabilities, as claimed in Corollary~\ref{Corolollary-ABS_Corr-abstract}.
	For completeness, we also show that Corollary~\ref{Corollary-NESS-ABS} is in turn corresponding to the transformation $|\Psi\rangle = S_{1}^{-1} |\Psi'\rangle$.
	\end{itemize}
\subsubsection{The similarity transformations}\label{subsectionSTransf}
Consider the matrices
\begin{align}
	\mathcal{S}_{1}:=\exp{\left(\sum_{a=1}^{N-1}e_{Na}\right)}\label{S1Gen}\co\\
	\mathcal{S}_{2}:=\exp{\left(-\sum_{a=1}^{N-1}\beta_{a}e_{aN}\right)}\dt\label{S2Gen}
\end{align}
We introduce
\begin{equation}\label{transformationV}
 \mathcal{S}_{1}^{x}:=\exp{\left(\sum_{a=1}^{N-1}e_{Na}^{x}\right)}\co %=\mathbbm{1}+\sum_{a=2}^{N}e_{1a}
\end{equation}
that is a copy of $\mathcal{S}_{1}$ (see \eqref{S1Gen}) acting at site $x$. 
Observe that $\mathcal{S}_{1}^{x}$ is the transposed of the bulk part of the duality matrix $d_{x}$  at site $x\in \{1,\ldots,L\}$ (see \eqref{bulkElementDualityMatrix}). By taking the product over the chain we define
\begin{equation}\label{S1-Whole}
    S_{1}=\prod_{x=1}^{L}\mathcal{S}_{1}^{x}
\end{equation}
that is related with the bulk duality matrix by 
\begin{equation}
S_{1}=\prod_{x=1}^{L}d_{x}^{T}\dt
\end{equation}
 As the transformation $S_{1}$ is invertible,  we can introduce 
\begin{equation}\label{hatHamiltonian}
   H^{'}=H^{'}_{\text{left}}+H_{\text{bulk}}+H^{'}_{\text{right}}
\end{equation}
that is related to $H$ by
\begin{equation}\label{similarV}
   H^{'}=S_{1}HS_{1}^{-1}\dt
\end{equation}
The bulk part of the Hamiltonian is left unchanged (because it is written using the coproduct of the second Casimir, see equation \eqref{hamiltonianCasimir}) while the boundaries are given by 
\begin{equation}
	H^{'}_{\text{left}}=\sum_{a=1}^{N-1}\left(\alpha_{a}e_{aN}^{1}-e_{aa}^{1}\right)\qquad H^{'}_{\text{right}}=\sum_{a=1}^{N-1}\left(\beta_{a}e_{aN}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
Indeed, using \eqref{HadTransfBoundary}, \eqref{ratesConditions} and \eqref{transpostionPropertyFund}, we obtain for the left boundary
\begin{equation}
	\begin{split}
		H_{\text{left}}^{'}&=S_{1}H_{\text{left}}S_{1}^{-1}=\exp{\left(\sum_{c=1}^{N-1}e_{Nc}^{1}\right)}\sum_{a,b=1}^{N}\alpha_{a}\left(e_{ab}^{1}-e_{bb}^{1}\right)\exp{\left(-\sum_{c=1}^{N-1}e_{Nc}^{1}\right)}
		\\&=
		\left\{\exp{\left(-\sum_{c=1}^{N-1}e_{cN}^{1}\right)}\sum_{a,b=1}^{N}\alpha_{a}\left(e_{ba}^{1}-e_{bb}^{1}\right)\exp{\left(\sum_{c=1}^{N-1}e_{cN}^{1}\right)}\right\}^{T}
		\\&=
		\sum_{a=1}^{N-1}\left(\alpha_{a}e_{Na}^{1}-e_{aa}^{1}\right)^{T}
		\\&=
		\sum_{a=1}^{N-1}\left(\alpha_{a}e_{aN}^{1}-e_{aa}^{1}\right)\dt
	\end{split}
\end{equation}
Similarly, for the right boundary we have 
\begin{equation}
H_{\text{right}}^{'}=S_{1}H_{\text{right}}S_{1}^{-1}=
% \exp{\left(\sum_{c=1}^{N-1}e_{1N}^{L}\right)}\sum_{a,b=1}^{N}\alpha_{a}\left(e_{ab}^{L}-e_{bb}^{L}\right)\exp{\left(-\sum_{c=1}^{N-1}e_{Nc}^{L}\right)}=
\sum_{a=1}^{N-1}\left(\beta_{a}e_{aN}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
\begin{remark}
	The boundary Hamiltonians $H_{\text{left}}^{'}$ and $H_{\text{right}}^{'}$ relates to the transposed of  $\widetilde{H}_{\text{left}}$ and $\widetilde{H}_{\text{right}}$, i.e. to the transposed of the boundary part of the dual Hamiltonian  defined in \eqref{siteDualHamiltonian}. They are obtained replacing the  extra-site bosonic creation operators $(\mathbf{a}^{+})_{a}^{0}$  (resp. $(\mathbf{a}^{+})_{a}^{L+1}$) with the corresponding reservoir parameters $\alpha_a$ (resp. $\beta_a$).
\end{remark}
We have the following correspondence between eigenvectors with zero eigenvalues
\begin{equation}\label{S1-Inverse}
	|\Psi\rangle =S_{1}^{-1}|\Psi^{'}\rangle
\end{equation}
where $|\Psi^{'}\rangle$ satisfies $H^{'}|\Psi^{'}\rangle=0$.

We introduce
 \begin{equation}
 	\mathcal{S}_{2}^{x}:=\exp{\left(-\sum_{a=1}^{N-1}\beta_{a}e_{aN}^{x}\right)}\co%=\mathbbm{1}-\sum_{a=2}^{N}\beta_{a}e_{a1}
 \end{equation}
which is a copy of $\mathcal{S}_{2}$ (see \eqref{S2Gen}) acting at site $x$. By taking the product over the chain we obtain 
\begin{equation}\label{S2-Whole}
	S_{2}=\prod_{x=1}^{L}\mathcal{S}_{2}^{x}\dt
\end{equation}
We define 
\begin{equation}\label{HSecond}
	H^{''}=H_{\text{left}}^{''}+H_{\text{bulk}}+H_{\text{right}}^{''}\co
\end{equation}
that is related to $H'$ by
\begin{equation}
	H^{''}=S_{2}H^{'}S_{2}^{-1}\dt
\end{equation}
The boundary Hamiltonians of $H^{''}$ reads
\begin{align}\label{tri-diag-boundaryHamiltonians}
	H_{\text{left}}^{''}=e_{NN}^{1}-\mathbbm{1}+\sum_{a=1}^{N-1}(\alpha_{a}-\beta_{a})e_{aN}^{1}\qquad H_{\text{right}}^{''}=e_{NN}^{L}-\mathbbm{1}\dt
\end{align}
The left one is lower triangular and depends on the differences between the boundary parameters, while the right one is diagonal. Equations \eqref{tri-diag-boundaryHamiltonians} are proved using the Hadamard formula \eqref{HadamardFormula} with $B=\sum_{c=1}^{N-1}\beta_{c}e_{cN}$, $A_{N}=e_{aN},$ and $A_{a}=e_{aa}$.

Therefore, using \eqref{HT-BA} we obtain for the left boundary 
\begin{equation}
	\begin{split}
		H_{\text{left}}^{''}&=S_{2}H_{\text{left}}^{'}S_{2}^{-1}
% 		=\exp{\left(-\sum_{c=1}^{N-1}\beta_{c}e_{cN}^{1}\right)}\sum_{a=1}^{N-1}\left(\alpha_{a}e_{aN}^{1}-e_{aa}^{1}\right)\exp{\left(-\sum_{c=1}^{N-1}\beta_{c}e_{cN}^{1}\right)}
		=
		\sum_{a=1}^{N-1}\left(\alpha_{a}e_{aN}^{1}-e_{aa}^{1}-\beta_{a}e_{aN}^{1}\right)
		=
		e_{NN}^{1}-\mathbbm{1}+\sum_{a=1}^{N-1}(\alpha_{a}-\beta_{a})e_{aN}^{1}\dt
	\end{split}
\end{equation}
Similarly, using again \eqref{HT-BA} we have for the right boundary 
\begin{equation}
	\begin{split}
			H_{\text{right}}^{''}=S_{2}H_{\text{right}}^{'}S_{2}^{-1}
			=
			\sum_{a=1}^{N-1}\left(\beta_{a}e_{aN}^{L}-e_{aa}^{L}-\beta_{a}e_{aN}^{L}\right)=
			e_{NN}^{L}-\mathbbm{1}\dt
	\end{split}
\end{equation}
The relation between eigenvectors with vanishing eigenvalue is 
\begin{equation}\label{S2-Inverse}
	|\Psi^{'}\rangle = S_{2}^{-1}|\Psi^{''}\rangle
\end{equation}
and $|\Psi^{''}\rangle$ solves
\begin{equation}\label{steadyS-SECOND-def}
	H^{''}|\Psi^{''}\rangle=0\dt
\end{equation}
It is convenient to introduce also the transformation
\begin{equation}\label{similarity}
	S=S_{2}S_{1}=\prod_{x=1}^{L}\exp{\left(-\sum_{a=1}^{N-1}\beta_{a}e_{aN}^{x}\right)}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\dt
\end{equation}
This matrix $S$ connects $H$ with $H^{''}$ by 
\begin{equation}
H=S^{-1}H^{''}S
\end{equation}
and, the relation between eigenvectors with vanishing eigenvalue is
\begin{equation}
	|\Psi\rangle=S^{-1}|\Psi^{''}\rangle\dt
\end{equation}

\subsubsection{Closed formula for $|\Psi^{''}\rangle$}\label{subsectionSSHsec}

The ground state of the Hamiltonian $H^{''}$ is given by 
\begin{equation}\label{ResulsBasis}
	|\Psi^{''} \rangle= \sum_{\bm{\tau}\in \Omega^{'}}\frac{1}{\left(1+L\right)!}\prod_{x=1}^{L} (\alpha_{\tau_{x}}-\beta_{\tau_{x}})^{(1-\delta_{\tau_{x},N})}\left(1+\sum_{j=x}^{L}\delta_{\tau_{j},N}\right) |\bm{\tau}\rangle
\end{equation}

where 
\begin{equation}
    \delta_{\tau_{x},N}:=\begin{cases}
        N\quad \text{if}\quad \tau_{x}=N\\
        0\quad \text{otherwise}
    \end{cases}\dt
\end{equation}

\paragraph{Proof of formula \eqref{ResulsBasis}}  
We consider the vector with elements given by the operators of the MPA $(X_{1},\ldots,X_{N})$ and we act on it with the similarity transformation $S=S_{2}S_{1}$ obtaining new operators $(\Xt_{1},\ldots,\Xt_{N})$ that will satisfy simpler commutation relations. 
For the sake of notation we denote by
\begin{equation}
    \begin{pmatrix}
		X_{a}\\ 
		X_{N}
	\end{pmatrix}=\begin{pmatrix}
	    X_{1}\\
     \vdots\\
     X_{N}
	\end{pmatrix}\dt
\end{equation}
We define the transformed matrix product operators
\begin{equation}\label{Xtildes2b}
	\begin{pmatrix}
		\Xt_{a}\\ 
		\Xt_{N}
	\end{pmatrix} =S\begin{pmatrix}
		X_{a}\\X_{N}
	\end{pmatrix}=\begin{pmatrix} 
		X_{a}-\beta_{a}(X_{1}+\ldots+X_{N})\\
				X_{1}+\ldots +X_{N}\\
	\end{pmatrix}\dt
\end{equation}
We can also reverse the transformation by $S^{-1}$ and get back: 
\begin{equation}\label{Xes}
	\begin{pmatrix}
		X_{a}\\
		X_{N} 
	\end{pmatrix} =S^{-1}\begin{pmatrix}
		\widetilde{X}_{a}\\
		\widetilde{X}_{N}
	\end{pmatrix}=\begin{pmatrix}
		\Xt_{a}+\beta_{a}\Xt_{N}\\ 
		\beta_N\Xt_{N}-(\Xt_{1}+\ldots+\Xt_{N-1})
	\end{pmatrix}\dt
\end{equation}
Summing over $b\in \{1,\ldots,N\}$ in commutation relations \eqref{bulk} we have
\begin{equation} 
	\left[X_{a},\Xt_{N}\right]=(\alpha_{a}-\beta_{a})\Xt_{N}
\end{equation}
with $\widetilde{X}_{N}=\widetilde{X}_{1}+\ldots+\widetilde{X}_{N}$. 
Therefore, we obtain the commutation relations for $\Xt_{a}$ 
\begin{equation}\label{commutationsBula}
	\left[\Xt_{a},\Xt_{N}\right]=(\alpha_{a}-\beta_{a})\Xt_{N}\co
\end{equation}
cf.~\eqref{commutationsBula2}.
Moreover using \eqref{Xtildes2b} and \eqref{Xes}, the action of $\Xt_{a}$ on the boundary vectors are given by 
\begin{equation}\label{commLEFT}
	\langle W|\left((\alpha_{a}-\beta_{a})\Xt_{N}-\Xt_{a}\right)=(\alpha_{a}-\beta_{a})\langle W|
	\co
\end{equation}
\begin{equation}\label{commRIGHT}
	\Xt_{a} |V\rangle= (\alpha_{a}-\beta_{a})|V\rangle
	\co
\end{equation} 
cf.~\eqref{commLEFT2}.
Using the transformed operators $(\widetilde{X}_{a})_{a\in\{1,\ldots,N\}}$, the vector $|\Psi^{''}\rangle$ of the matrix product ansatz is written as
\begin{equation}\label{toDetermineTilde}
	|\Psi^{''}\rangle = \frac{1}{Z_{L}}\sum_{\bm{\tau}\in \Omega^{'}}\langle W|\prod_{x=1}^{L}\widetilde{X}_{\tau_{x}}
	|V \rangle \,|\bm{\tau}\rangle
\end{equation}
where the normalization is given by 
\begin{align}
		Z_{L}&=\langle W|(X_{1}+\ldots+X_{N})^{L}|V\rangle=\langle W|\widetilde{X}_{N}^{L}|V\rangle\dt
\end{align}

To determine the eigenvector \eqref{toDetermineTilde}, we compute the coefficient $\langle W|\prod_{x=1}^{L}\widetilde{X}_{\tau_{x}}
|V \rangle$ and the normalization $Z_{L}$. Using \eqref{commutationsBula} we have
\begin{equation}\label{UsefulRelation}
	\widetilde{X}_{a}\widetilde{X}_{N}=\widetilde{X}_{N}\left((\alpha_{a}-\beta_{a})+\widetilde{X}_{a}\right)\dt
\end{equation}
Fix $ a\in \{1,\ldots,N-1\}$ and arbitrary $ \ell,n\in \mathbb{N}$. By applying the above formula \eqref{UsefulRelation} we have
\begin{align}\label{usefulRelaton-many}
	\widetilde{X}_{a}^{n}\widetilde{X}_{N}^{\ell}&=
	\widetilde{X}_{N}^{\ell}\left(\widetilde{X}_{a}+\ell(\alpha_{a}-\beta_{a})\right)^{n}\dt
\end{align}

By taking the product over the sites $x$ from $1$ to $L$ we obtain 
\begin{equation}
	\prod_{x=1}^{L}\widetilde{X}_{\tau_{x}}=\widetilde{X}_{N}^{\sum_{x=1}^{L}\delta_{\tau_{x},N}}\prod_{x=1}^{L}\left(\widetilde{X}_{\tau_{x}}+(\alpha_{\tau_{x}}-\beta_{\tau_{x}})\sum_{j=x}^{L}\delta_{\tau_{x},N}\right)^{1-\delta_{\tau_{x},N}}\dt
\end{equation}
Multiplying by the boundary vectors and by using \eqref{commRIGHT} we have 
\begin{align}
	\langle W|\prod_{x=1}^{L}\widetilde{X}_{\tau_{x}}
	|V \rangle=
	\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-\delta_{\tau_{x},N}}\left(1+\sum_{j=x}^{L}\delta_{\tau_{j},N}\right)^{1-\delta_{\tau_{x},N}}\langle W|\widetilde{X}_{N}^{\sum_{x=1}^{L}\delta_{\tau_{x},N}}|V\rangle \dt
\end{align}
We now compute $\langle W|\widetilde{X}_{N}^{\sum_{i=1}^{L}\delta_{\tau_{x},N}}|V\rangle$. For all $n\in \mathbb{N}$ we have that  
\begin{align}
	\langle W|\widetilde{X}_{N}^{n}|V\rangle&=\langle W|\widetilde{X}_{N}\widetilde{X}_{N}^{n-1}|V\rangle=\langle W|\widetilde{X}_{N}^{n-1}|V\rangle +\langle W|\frac{1}{(\alpha_{a}-\beta_{a})}\widetilde{X}_{a}\widetilde{X}_{N}^{n-1}|V\rangle\nonumber
	\\&=
	\langle W|\widetilde{X}_{N}^{n-1}|V\rangle+\frac{1}{(\alpha_{a}-\beta_{a})}\langle W|\widetilde{X}_{N}^{n-1}\left(\widetilde{X}_{a}+(\alpha_{a}-\beta_{a})(n-1)\right)|V\rangle\nonumber
	\\&=
	\langle W|\widetilde{X}_{N}^{n-1}|V\rangle+\left(n+1-1\right)\langle W|\widetilde{X}_{N}^{n-1}|V\rangle\nonumber
	\\&=
	\left(2+n-1\right)\langle W|\widetilde{X}_{N}^{n-1}|V\rangle\nonumber
	\\&=
	\frac{(1+n)!}{n!}\langle W|\widetilde{X}_{N}^{n-1}|V\rangle
\end{align}
where in the second equality we used \eqref{usefulRelaton-many}. 
This leads to the recursion relation
\begin{align}
		\langle W|\widetilde{X}_{N}^{n}|V\rangle=\frac{(1+n)!}{n!}\langle W|\widetilde{X}_{N}^{n-1}|V\rangle
\end{align}
with $	\langle W|\widetilde{X}_{N}^{0}|V\rangle=1$. 
Therefore, we obtain 
\begin{align}
	\langle W|\widetilde{X}_{N}^{n}|V\rangle&=%\frac{\Gamma(2+n_{N})}{\Gamma(2+n_{N}-1)}\langle W|\widetilde{X}_{N}^{n_{N}-1}|V\rangle=\langle W|\widetilde{X}_{N}^{n_{N}}|V\rangle=\frac{\Gamma(2+n_{N})}{\Gamma(2+n_{N}-1)}\frac{\Gamma(2+n_{N}-1)}{\Gamma(2+n_{N}-2)}\langle W|\widetilde{X}_{N}^{n_{N}-2}|V\rangle\\&=
	\frac{(1+n)!}{n!}\frac{n!}{(n-1)!}\ldots \frac{3!}{2!}\frac{2!}{1!}\langle W|\widetilde{X}_{N}^{0}|V\rangle=
	(1+n)!\dt
\end{align}
By using this result we have that
\begin{equation}
	\langle W|\prod_{x=1}^{L}\widetilde{X}_{\tau_{x}}
	|V \rangle=(1+\sum_{x=1}^{N}\delta_{\tau_{x},N})!\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-\delta_{\tau_{x},N}}\left(1+\sum_{j=x}^{L}\delta_{\tau_{j},N}\right)^{1-\delta_{\tau_{x},N}}\dt
\end{equation}
The normalization constant is then given by
\begin{align}
	Z_{L}&=\langle W|(X_{1}+\ldots+X_{N})^{L}|V\rangle=\langle W|\widetilde{X}_{N}^{L}|V|\rangle=
	(L+1)!\dt
\end{align}
Therefore, we write
\begin{equation}\label{resulEsteady}
	|\Psi^{''}\rangle= \sum_{\bm{\tau}\in \Omega^{'}}\frac{(1+\sum_{x=1}^{L}\delta_{\tau_{x},N})!}{(L+1)!}\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-\delta_{\tau_{x},N}}\left(1+\sum_{j=x}^{L}\delta_{\tau_{j},N}\right)^{1-\delta_{\tau_{x},N}}|\bm{\tau}\rangle\dt
\end{equation}
We observe that, for all $x$, we have
\begin{equation}
	\left(1+\sum_{j=x}^{L}\delta_{\tau_{j},N}\right)^{1-\delta_{\tau_{x},N}}=\frac{(1+\sum_{j=x+1}^{L}\delta_{\tau_{j},N})!}{(\sum_{j=x}^{L}\delta_{\tau_{j},N})!}\dt
\end{equation}

It follows that 
\begin{align}\label{remove-exponent}
&\frac{(1+\sum_{x=1}^{L}\delta_{\tau_{x},N})!}{(L+1)!}\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-\delta_{\tau_{x},N}}\left(1+\sum_{j=x}^{L}\delta_{\tau_{j},N}\right)^{1-\delta_{\tau_{x},N}}\nonumber\\
=&\frac{(1+\sum_{x=1}^{L}\delta_{\tau_{x},N})!}{(L+1)!}\prod_{x=1}^{L}\frac{(1+\sum_{j=x+1}^{L}\delta_{\tau_{j},N})!}{(\sum_{j=x}^{L}\delta_{\tau_{j},N})!}\nonumber\\
=&\frac{1}{(L+1)!}\prod_{x=1}^{L}\frac{(1+\sum_{j=x}^{L}\delta_{\tau_{j},x})!}{(\sum_{j=x}^{L}\delta_{\tau_{j},x})!}\nonumber\\
=&\frac{1}{(L+1)!}\prod_{x=1}^{L}\left(1+\sum_{j=x}^{L}\delta_{\tau_{j},N}\right)\dt
\end{align}
Therefore, we obtain \eqref{ResulsBasis}.
\begin{flushright}
	$\square$
\end{flushright} 
\subsubsection{Closed formula for $|\Psi^{'}\rangle$}\label{subsectionSSdual}
By knowing the ground state  of the Hamiltonian $H^{''}$ we use \eqref{S2-Inverse} to retrieve the ground state of the Hamiltonian $H^{'}$. The result is the following 
\begin{equation}\label{ABS-vect}
    |\Psi^{'}\rangle =\sum_{\bm{\tau}\in \Omega^{'}}\Psi^{'}(\bm{\tau})|\bm{\tau}\rangle 
\end{equation}
where  
\begin{equation}\label{ABS}
		\begin{split}
			\Psi^{'}(\bm{\tau})=\frac{1}{(L+1)!}\sum_{c_{1}=\delta_{\tau_{1},N}}^{1}\ldots\sum_{c_{L}=\delta_{\tau_{L},N}}^{1}\prod_{x=1}^{L}\left(1+\sum_{j=x}^{L}c_{j}\right)(\alpha_{\tau_{x}}-\beta_{\tau_{x}})^{1-c_{x}}\beta_{\tau_{x}}^{c_{x}-\delta_{\tau_{x},N}}\dt
		\end{split}
	\end{equation} 
\paragraph{Proof of formula \eqref{ABS-vect}} We have that $|\Psi^{'}\rangle=S_{2}^{-1}|\Psi^{''}\rangle$, thus we  show how the transformation $S_{2}^{-1}$ acts on the vector $|\Psi^{''}\rangle$. Using the exponential series we have that  
\begin{equation}
   \mathcal{S}_{2}^{-1}=\mathbbm{1}+\sum_{a=1}^{N-1}\beta_{a}e_{aN}\co
\end{equation}
which action on the occupation variable $|\sigma\rangle$ of a site is 
\begin{equation}
  \mathcal{S}_{2}^{-1} |\sigma\rangle= |\sigma\rangle+\delta_{N,\sigma}\sum_{a=1}^{N-1}\beta_a|a\rangle\dt
\end{equation} 
Therefore, taking the tensor product over the chain we obtain 
\begin{equation}
 S_{2}^{-1} |\bm{\sigma}\rangle=\left( |\sigma_1\rangle+\delta_{N,\sigma_1}\sum_{a_1=1}^{N-1}\beta_{a_1}|a_1\rangle\right)\otimes\ldots\otimes \left( |\sigma_L\rangle+\delta_{N,\sigma_L}\sum_{a_L=1}^{N-1}\beta_{a_L}|a_L\rangle\right)\dt
\end{equation} 
By projecting over a vector $\langle \bm{\tau}|$ we have 
\begin{equation}\label{goodTP-S2}
  \langle \bm{\tau}|S_{2}^{-1} |\bm{\sigma}\rangle=\prod_{x=1}^L\left[ \delta_{\tau_x,\sigma_x}+\delta_{N,\sigma_x}\beta_{\tau_x} (1-\delta_{\tau_x,N})\right]\dt
\end{equation} 
From this it follows that
\begin{align}\label{ABS_intermediate}
		\Psi^{'}(\bm{\tau})=&\langle\bm{\tau}|\Psi^{'}\rangle=\langle \bm{\tau}|S_{2}^{-1}|\Psi^{''}\rangle = \sum_{\bm{\sigma}\in \Omega'}  \langle \bm{\tau}|S_{2}^{-1} |\bm{\sigma}\rangle\langle\bm{\sigma}|\Psi^{''}\rangle \nonumber\\=&\sum_{\bm{\sigma}\in \Omega^{'}}\frac{1}{(L+1)!}\prod_{x=1}^L(\alpha_{\sigma_{x}}-\beta_{\sigma_{x}})^{1-\delta_{\sigma_{x},N}}\left(1+\sum_{j=x}^{L}\delta_{\sigma_{j},N}\right)\left[ \delta_{\tau_x,\sigma_x}+\beta_{\tau_x} \delta_{N,\sigma_x} (1-\delta_{\tau_x,N})\right]\dt
	\end{align}
	where we used the expression \eqref{ResulsBasis} for $\Psi^{''}(\bm{\sigma})$.
We observe that, for any fixed $x\in \{1,\ldots,L\}$ and for any fixed $\bm{\tau},\bm{\sigma}\in\Omega^{'}$, we have 
\begin{align}
	\left[\delta_{\tau_{x},\sigma_{x}}+\delta_{\sigma_{x},N}(1-\delta_{\tau_{x},N})\beta_{\tau_{x}}\right]\left(\alpha_{\sigma_{x}}-\beta_{\sigma_{x}}\right)^{1-\delta_{\sigma_{x},N}}=	\left[\delta_{\tau_{x},\sigma_{x}}+\delta_{\sigma_{x},N}(1-\delta_{\tau_{x},N})\beta_{\tau_{x}}\right]\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-\delta_{\sigma_{x},N}}\dt
\end{align}
Indeed, on one hand if $\sigma_{x}\neq N$ only the term $\delta_{\tau_{x},\sigma_{x}}$ does survive in the square brackets of the above equation, that is non-vanishing only if $\sigma_{x}=\tau_{x}$. This implies that both the left-hand-side and the right-hand-side are either $(\alpha_{\tau_{x}}-\beta_{\tau_{x}})$ ro $0$. On the other hand, if $\sigma_{x}=N$ we have $(\alpha_{\sigma_{x}}-\beta_{\sigma_{x}})^{1-\delta_{\sigma_{x},N}}=(\alpha_{\tau_{x}}-\beta_{\tau_{x}})^{1-\delta_{\sigma_{x},N}}=1$, thus both sides of the equality are either $\beta_{\tau_{x}}$ or $0$.  \\
Moreover, we have that 
\begin{align}
	\delta_{\sigma_{x},\tau_{x}}+\delta_{\sigma_{x},N}(1-\delta_{\tau_{x},N})\beta_{\tau_{x}}=&\beta_{\tau_{x}}^{\delta_{\sigma_{x},N}(1-\delta_{\tau_{x},N})}\left(\delta_{\tau_{x},\sigma_{x}}+\delta_{\sigma_{x},N}(1-\delta_{\tau_{x},N})\right)\nonumber\\
	=&\beta_{\tau_{x}}^{\delta_{\sigma_{x},N}(1-\delta_{\tau_{x},N})}\left(\delta_{\tau_{x},N}\delta_{\sigma_{x},N}+(1-\delta_{\tau_{x},N})(1-\delta_{\sigma_{x},N})+(1-\delta_{\tau_{x},N})\delta_{\sigma_{x},N}\right)\nonumber\\
	=&\beta_{\tau_{x}}^{\delta_{\sigma_{x},N}(1-\delta_{\tau_{x},N})}\left(\delta_{\sigma_{x},N}+(1-\delta_{\sigma_{x},N})(1-\delta_{\tau_{x},N})\right)
\end{align}
Indeed, considering the first and last equalities we have that: if $\tau_{x}=\sigma_{x}$, both sides are $1$; if $\tau_{x}\neq N$ and $\sigma_{x}=N$ both sides are $\beta_{\tau_{x}}$; while, in all other cases, both sides are $0$.

As a consequence we can write \eqref{ABS_intermediate} as 
\begin{align}
\Psi^{'}(\bm{\tau})%=&\frac{1}{(L+1)!}\sum_{\sigma_{1}=1}^{N}\ldots\sum_{\sigma_{L}=1}^{N}\prod_{x=1}^{L}\left(\alpha_{\sigma_{x}}-\beta_{\sigma_{x}}\right)^{1-c_{x}}\left(1+\sum_{j=x}^{L}c_{x}\right)\left[\delta_{\tau_{x},\sigma_{x}}c_{x}+\delta_{\tau_{x},\sigma_{x}}(1-c_{x})+\beta_{\tau_{x}}c_{x}(1-\delta_{\tau_{x},N})\right]\nonumber\\
=&\frac{1}{(L+1)!}\sum_{\bm{\sigma}\in \Omega^{'}}^{N}\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-\delta_{\sigma_{x},N}}\left(1+\sum_{j=x}^{L}\delta_{\sigma_{j},N}\right)\nonumber\\&\left[\beta_{\tau_{x}}^{\delta_{\sigma_{x},N}(1-\delta_{\tau_{x},N})}\left(\delta_{\sigma_{x},N}+(1-\delta_{\sigma_{x},N})(1-\delta_{\tau_{x},N})\right)\right]\dt
%=&\frac{1}{(L+1)!} \sum_{c_{1}=0}^{1}\ldots\sum_{c_{L}=0}^{1}\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-c_{x}}\left(1+\sum_{j=x}^{L}c_{x}\right)\left[\delta_{\tau_{x},N}c_{x}+(1-\delta_{\tau_{x},N})(1-c_{x})+\beta_{\tau_{x}}c_{x}(1-\delta_{\tau_{x},N})\right]\nonumber\\
%=&\frac{1}{(L+1)!}\sum_{c_{1}=\delta_{\tau_{1},N}}^{1}\ldots\sum_{c_{L}=\delta_{\tau_{L},N}}^{1}\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-c_{x}}\left(1+\sum_{j=x}^{L}c_{x}\right)\beta_{\tau_{x}}^{c_{x}(1-\delta_{\tau_{x},N})}
\end{align}
We observe that the argument of the summation above does not distinguish the colours of the $\sigma$'s but only whether, at each site $x$, $\sigma_{x}$ is occupied or empty. Therefore, we can replace the summation over $\sigma_{1},\ldots,\sigma_{L}$ with a summation over the indices  $c_{1},\ldots,c_{L}\in \{0,1\}$. Thus, we obtain 
\begin{align}
\Psi^{'}(\bm{\tau})	=&\frac{1}{(L+1)!}\sum_{c_{1}=0}^{1}\ldots\sum_{c_{L}=0}^{1}\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-c_{x}}\left(1+\sum_{j=x}^{L}c_{j}\right)\left[\beta_{\tau_{x}}^{c_{x}}\left(c_{x}+(1-c_{x})(1-\delta_{\tau_{x},N})\right)\right]\dt
\end{align}
Moreover, because of the term $(1-\delta_{\tau_{x},N})(1-c_{x})$ we can make the indices $c_{x}$ vary in the set $\{\delta_{\tau_{x},N},1\}$. Indeed, when $c_{x}=0$, this term is non-vanishing only if $\tau_{x}\neq N$. Therefore, we obtain 
\begin{equation}
	\Psi^{'}(\bm{\tau})=\frac{1}{(L+1)!}\sum_{c_{1}=\delta_{\tau_{1},N}}^{1}\ldots\sum_{c_{L}=\delta_{\tau_{L},N}}^{1}\prod_{x=1}^{L}\left(\alpha_{\tau_{x}}-\beta_{\tau_{x}}\right)^{1-c_{x}}\left(1+\sum_{j=x}^{L}c_{j}\right)\beta_{\tau_{x}}^{c_{x}(1-\delta_{\tau_{x},N})}\dt
\end{equation}

Observing now that $c_{x}(1-\delta_{\tau_{x},N})$ can be replaced by  $c_{x}-\delta_{\tau_{x},N}$, we finally have
\begin{equation}\label{elementsABS}
	\begin{split}
		\Psi^{'}(\bm{\tau})=\frac{1}{(L+1)!}\sum_{c_{1}=\delta_{\tau_{1},N}}^{1}\ldots\sum_{c_{L}=\delta_{\tau_{L},N}}^{1}\prod_{x=1}^{L}(\alpha_{\tau_{x}}-\beta_{\tau_{x}})^{1-c_{x}}\left(1+\sum_{j=x}^{L}c_{j}\right)\beta_{\tau_{x}}^{c_{x}-\delta_{\tau_{x},N}}\dt
	\end{split}
\end{equation} 
\begin{flushright}
    $\square$
\end{flushright}
\begin{remark} By changing the summation indices from $c_x\in \{\delta_{\tau_{x},N},1\}$ to $c_{x}^{'}\in \{0,1-\delta_{\tau_{x},N}\}$ and by  the fact that occupation of each site is bounded by $1$ we rewrite \eqref{ABS} as 
\begin{equation}\label{ABS_1}
	\Psi^{'}(\bm{\tau})=\sum_{c^{'}_{1}=0}^{1-\delta_{\tau_{1},N}}\ldots\sum_{c^{'}_{L}=0}^{1-\delta_{\tau_{L},N}}\frac{(L+1-\sum_{x=1}^{L}c^{'}_{x})!}{(L+1)!}\prod_{x=1}^{L}\left((\alpha_{\tau_{x}}-\beta_{\tau_{x}})\left(2+L-x-\sum_{j=x}^{L}c^{'}_{j}\right)\right)^{c^{'}_{x}}\beta_{\tau_{x}}^{(1-c^{'}_{x}-\delta_{\tau_{x},N})}
\end{equation}
where we used the fact that 
\begin{equation}
	\left(1+\sum_{j=x}^{L}c_{j}\right) =\left(2+L-x-\sum_{j=x}^{L}c^{'}_{j}\right)\dt
\end{equation}
Equation \eqref{ABS_1} will be useful in the following. 
\end{remark}
\subsubsection{Closed formula for correlations}\label{correlation-section}
We consider $m\in \{1,\ldots,L\}$ sites with coordinates $x_{1},\ldots,x_{m}\in \{1,\ldots,L\}$ such that $x_{k}<x_{k+1}$ for all $k\in \{1,\ldots,m-1\}$. We fix $a_{k}\in\{1,\ldots,N-1\}$ and we would like to compute
\begin{equation}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=\sum_{\bm{\tau}\in \Omega^{'}}\left(\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right)\langle\bm{\tau}|\Psi\rangle\dt%=\sum_{\bm{\tau}\in \Omega^{'}}\left(\prod_{i=1}^{m}\mathbbm{1}_{\{\tau_{y_{i}}=s_{i}\}}\right)\Psi(\bm{\tau})
\end{equation}
We fix the dual configuration $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$, corresponding to a vector $|\bm{\xi}\rangle$, and we have 
\begin{equation}
	\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}=D(\bm{\tau},\bm{\xi})\dt
\end{equation}
Thus it follows that 
\begin{align}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=&\sum_{\bm{\tau}\in \Omega^{'}}D(\bm{\tau},\bm{\xi})\langle\bm{\tau}|\Psi\rangle
	=\sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|D|\bm{\xi}\rangle\langle\bm{\tau}|\Psi\rangle\dt
\end{align}
Denoting by $|\hat{\bm{\xi}}\rangle=|\xi^{1},\ldots,\xi^{L}\rangle$, i.e. the vector constructed from $|\bm{\xi}\rangle$ by removing the components at sites $\{0\}$ and $\{L+1\}$ and recalling that  
\begin{equation}
	S_{1}^{T}=\prod_{x=1}^{L}d_{x}
\end{equation} 
we have
\begin{equation}
	\langle \bm{\tau}|S_{1}^{T}|\hat{\bm{\xi}}\rangle=\langle \bm{\tau}|D|\bm{\xi}\rangle \dt
\end{equation}
Therefore, it follows that, using the resolution of the identity
\begin{align}
\label{corr-psi1}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=& \sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|D|\bm{\xi}\rangle\langle\bm{\tau}|\Psi\rangle
	=\sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|S_{1}^{T}|\hat{\bm{\xi}}\rangle\langle\bm{\tau}|\Psi\rangle\nonumber\\
	=&\sum_{\bm{\tau}\in \Omega^{'}}\langle \hat{\bm{\xi}}|S_{1}|\bm{\tau}\rangle\langle\bm{\tau}|\Psi\rangle
	=\langle \hat{\bm{\xi}}|S_{1}|\Psi\rangle
	=\langle \hat{\bm{\xi}}|\Psi^{'}\rangle\dt
\end{align}
Using $\Psi^{'}(\bm{\tau})$ in \eqref{ABS_1} we have 
\begin{align}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=&\sum_{c_{1}=0}^{1}\ldots\sum_{c_{m}=0}^{1}\frac{(L+1-\sum_{a=1}^{m}c_{a})!}{(L+1)!}\prod_{k=1}^{m}(\alpha_{a_{k}}-\beta_{a_{k}})^{c_{k}}\beta_{a_{k}}^{1-c_{k}}\left(2+L-x_{k}-\sum_{j=k}^{m}c_{j}\right)^{c_{k}}\nonumber\\
	=&\sum_{c_{1},\ldots,c_{m}=0}^{1}
	f(c_{1},\ldots,c_{m})\prod_{k=1}^{m}(\alpha_{a_{k}}-\beta_{a_{k}})^{c_{k}}\beta_{a_{k}}^{1-c_{k}}g_{k}(x_{k},c_{k},\ldots,c_{m})\co
\end{align}
where in the last equality we have used the definition of \eqref{powerCoeffNOspec} and \eqref{powerCoeffSpecies}.
As a last step, using the binomial theorem, we rewrite the previous expression as a polynomial in the boundary densities: 
\begin{align}
	&\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]\nonumber\\=&
	\sum_{c_{1},\ldots,c_{m}=0}^{1}
	f(c_{1},\ldots,c_{m})\prod_{k=1}^{m}\left(\sum_{t_{k}=0}^{c_{k}}(-1)^{c_{k}-t_{k}}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}g_{k}(x_{k},c_{k},\ldots,c_{m})\right)\nonumber\\
	=&\sum_{c_{1},\ldots,c_{m}=0}^{1}
	f(c_{1},\ldots,c_{m})\sum_{t_{1}=0}^{c_{1}}\ldots\sum_{t_{m}=0}^{c_{m}}\left(\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\right)\left(\prod_{j=1}^{m}(-1)^{c_{j}-t_{j}}g_{j}(x_{j},c_{j},\ldots,c_{m})\right)\nonumber\\
%	=&\sum_{t_{1},\ldots,t_{m}=0}^{1}
%	\sum_{c_{1}=t_{1}}^{1}\ldots\sum_{c_{m}=t_{m}}^{1}f(c_{1},\ldots,c_{m})\left(\prod_{k=1}^{m}\alpha_{s_{k}}^{t_{k}}\beta_{s_{k}}^{1-t_{k}}\right)\left(\prod_{j=1}^{m}(-1)^{c_{j}-t_{j}}g_{j}(x_{j},c_{j},\ldots,c_{m})\right)\nonumber\\
	=&\sum_{t_{1},\ldots,t_{m}=0}^{1}\left(\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\right)\sum_{c_{1}=t_{1}}^{1}\ldots\sum_{c_{m}=t_{m}}^{1}f(c_{1},\ldots,c_{m})\left(\prod_{k=1}^{m}(-1)^{c_{k}-t_{k}}g_{k}(x_{k},c_{k},\ldots,c_{m})\right)\dt
\end{align}
From third to fourth line we have changed the summation order. Therefore, comparing the above formula and \eqref{absProbabilitiesIntegrable}, we can read off the absorption probabilities given in \eqref{abs-probabilities}.
\begin{flushright}
	$\square$
\end{flushright}
\subsubsection{Interpretation of $\Psi^{'}(\bm{\tau})$ as marginals of the steady state distribution}\label{subsection-interpretationPSIp}
%We argue similarly as in the Section~\ref{correlation-section}. 
We still miss the interpretation of $\Psi^{'}(\bm{\tau})$ when  $\bm{\tau})= (N,\ldots,N)$. We show here that for this particular configuration
$\Psi^{'}(N,\ldots,N) =1$. Indeed, for any $\bm{\tau}\in \Omega^{'}$ we have that 
\begin{align}
	\Psi^{'}(\bm{\tau})=&\langle\bm{\tau}|\Psi^{'}\rangle=\sum_{\bm{\sigma}\in \Omega}\langle \bm{\tau}|S_{1}|\bm{\sigma}\rangle\langle \bm{\sigma}|\Psi\rangle=\sum_{\sigma\in \Omega}\langle \bm{\sigma}|S_{1}^{T}|\bm{\tau}\rangle\langle \bm{\sigma}|\Psi\rangle\dt
\end{align}
We introduce the dual configuration $\bm{\xi}(\bm{\tau})=(N,\tau_{1},\ldots,\tau_{L},N)$ where two extra-sites occupied by the species $N$ have been attached to the vector $\bm{\tau}$, transforming it from a vector of dimension $L$ to one of dimension $L+2$. More explicitly, this dual configuration is given by
\begin{equation}
	\begin{cases}
	\xi_{a}^{x}(\bm{\tau})= \delta_{\tau_{x},a}\qquad \forall a\in \{1,\ldots N\}\quad \forall x\in \{1,\ldots,L\}\\
	\xi_{N}^{0}(\bm{\tau})=\xi_{N}^{L+1}(\bm{\tau})=1
\end{cases}\dt
\end{equation} Now we consider the elements $\langle\bm{\sigma}|D|\xi(\bm{\tau})\rangle$ of the duality matrix $D$ and we have
\begin{align}\label{useful-InInterpretation}
	\Psi^{'}(\bm{\tau})=&\sum_{\sigma\in \Omega}\langle \bm{\sigma}|D|\bm{\xi}(\bm{\tau})\rangle\langle \bm{\sigma}|\Psi\rangle\dt
\end{align}
We observe that for any $x\in \{1,\ldots,L\}$
\begin{equation}
	\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\sigma_{x},a}\geq\xi_{a}^{x}(\bm{\tau})\}}=\left(\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\right)^{(1-\delta_{\tau_{x},N})}%=\begin{cases}
	%	\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\qquad &\text{if}\quad \tau_{x}\neq N\\
	%	1\qquad  &\text{if}\quad \tau_{x}= N
%	\end{cases}\dt
\end{equation}
Indeed, we have that 
\begin{equation}
	\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\sigma_{x},a}\geq\xi_{a}^{x}(\bm{\tau})\}}=\begin{cases}
			\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\qquad &\text{if}\quad \tau_{x}\neq N\\
			1\qquad  &\text{if}\quad \tau_{x}= N
			\end{cases}\dt
\end{equation}
 Therefore using the duality function \eqref{ElementsDualityMatrixChain_1} and equation~\ref{useful-InInterpretation}, we obtain 
 	\begin{equation}\label{interpretation-PsiP}
 	\Psi^{'}(\bm{\tau})=\mathbb{E}\left[\prod_{x=1}^{L}\left(\mathbbm{1}_{\{Y_{x}=\tau_{x}\}} \right)^{(1-\delta_{\tau_{x},N})}\right]\dt
 \end{equation} 
 This last equation completes and generalizes \eqref{corr-psi1} by showing that $\Psi^{'}(N,\ldots,N) =1$.

 Equivalently, we can interpret the components of $\Psi'(\bm{\tau})$ as follows.
 Let $\ell$ be the number of occupied sites in the configuration $\bm{\tau}$, i.e.  $\ell=\sum_{x=1}^{L}(1-\delta_{\tau_{x},N})$. We denote by $x_{1},\ldots,x_{\ell}$, such that $x_{k}<x_{k+1}$ for all $k\in \{1,\ldots,\ell-1\}$, the coordinates where a particle of any (non-empty) type is present. Finally, we call 
 \begin{equation}
 	\Omega_{\ell}^{'}(\bm{\tau}):=\left\{\bm{\sigma}\in \Omega^{'}\;:\; \sigma_{x_{1}}=\tau_{x_{1}},\ldots,\sigma_{x_{\ell}}=\tau_{x_{\ell}}\right\}\dt
 \end{equation} 
Observe that, if $\ell=0$ then $\Omega_{\ell}^{'}(\bm{\tau})$ does coincide with the whole state space $\Omega^{'}$, while, if $\ell=L$ then $\Omega_{\ell}^{'}(\bm{\tau})$ reduces to the configuration $\bm{\tau}$. Denoting the marginal on $\ell$ sites of the steady state distribution by 
\begin{equation}
	\mu_{x_{1},\ldots,x_{\ell}}(\bm{\tau})=\sum_{\bm{\sigma}\in \Omega_{\ell}^{'}(\bm{\tau})}\mu(\bm{\sigma})
\end{equation}
we have that 
\begin{align}
	\Psi^{'}(\bm{\tau})=&\mathbb{E}\left[\prod_{x=1}^{L}\left(\mathbbm{1}_{\{Y_{x}=\tau_{x}\}} \right)^{(1-\delta_{\tau_{x},N})}\right]=\mathbb{E}\left[\prod_{k=1}^{\ell}\mathbbm{1}_{\{Y_{x_{k}}=\tau_{x_{k}}\}}\right]=\sum_{\bm{\sigma}\in \Omega_{\ell}^{'}(\bm{\tau})}\mu(\bm{\sigma})=\mu_{x_{1},\ldots,x_{\ell}}(\bm{\tau})\dt
\end{align} 
It follows that $\Psi^{'}(\bm{\tau})$ is the marginal on $\ell$ sites $x_{1},\ldots,x_{\ell}$ of the non-equilibrium steady state distribution. In particular, when the configuration $\bm{\tau}$ has $\ell$ particles, $\Psi^{'}(\bm{\tau})$ does coincide with the $\ell$-point correlation function. 
%As a consequence, all the marginals of $\mu$ can be found using the absorption probabilities by equation \eqref{exactCorrelations}.

\subsubsection{Steady state}\label{appB}
By the knowledge of the closed formula for correlations \eqref{exactCorrelations}, we find the probability mass-function in the non-equilibrium steady state by using Corollary~\ref{Corollary-NESS-ABS}. Indeed we show here that 
$|\Psi\rangle=S_{1}^{-1}|\Psi^{'}\rangle$ is nothing but Corollary~\ref{Corollary-NESS-ABS}.
On a single site we have that 
\begin{equation}\label{S1-L1}
	\mathcal{S}_{1}^{-1}=\mathbbm{1}+(-1)\sum_{a=1}^{N-1}e_{Na}
\end{equation}
thus,
\begin{equation}
	\mathcal{S}_{1}^{-1}|\sigma\rangle=|\sigma\rangle+(-1)(1-\delta_{\sigma,N})|N\rangle
\end{equation}
By taking the tensor product over the chain we obtain 
\begin{equation}
	S_{1}^{-1}|\bm{\sigma}\rangle=\left(|\sigma_{1}\rangle+(-1)(1-\delta_{\sigma_{1},N})|N\rangle\right)\otimes \ldots \otimes \left(|\sigma_{L}\rangle+(-1)(1-\delta_{\sigma_{L},N})|N\rangle\right)
\end{equation}
therefore, projecting on $\langle \bm{\tau}|$ we have
\begin{equation}\label{Tensor-S1}
	\langle \bm{\tau}|	S_{1}^{-1}|\bm{\sigma}\rangle=\prod_{x=1}^{L}\left(\delta_{\tau_{x},\sigma_{x}}+(-1)\delta_{\tau_{x},N}\sum_{a_{x}=1}^{N-1}\delta_{\sigma_{x},a_{x}}\right)\dt
\end{equation}

		Considering 
		\begin{equation}
			|\Psi^{'}\rangle=\sum_{\bm{\sigma}\in \Omega^{'}}\Psi^{'}(\bm{\sigma})|\bm{\sigma}\rangle\co
		\end{equation} we obtain  
		\begin{align}\label{implicit-PSI-PSIp}
			\langle \bm{\tau}|S_{1}^{-1}|\Psi^{'}\rangle=&\sum_{\bm{\sigma}\in \Omega^{'}}\Psi^{'}(\bm{\sigma})\prod_{x=1}^{L}\left(\delta_{\tau_{x},\sigma_{x}}+(-1)\delta_{\tau_{x},N}\sum_{a_{x}=1}^{N-1}\delta_{\sigma_x,a_{x}}\right)\nonumber\\
			=&\sum_{b=0}^{L}(-1)^{b}\sum_{1\leq q_{1}<\ldots<q_{b}\leq L}\sum_{\bm{\sigma}\in \Omega^{'}}\Psi^{'}(\bm{\sigma})\left(\prod_{x=1\,:\,x\notin\{q_{1},\ldots,q_{b}\}}^{L}\delta_{\tau_{x},\sigma_x}\right)\left(\prod_{x\in \{q_{1},\ldots,q_{b}\}}\delta_{\tau_{x},N}\sum_{a_{x}=1}^{N-1}\delta_{\sigma_{x},a_{x}}\right)\nonumber\\
			=&\sum_{b=0}^{L}(-1)^{b}\sum_{1\leq q_{1}<\ldots<q_{b}\leq L}^{L}\left(\prod_{k=1}^{b}\delta_{\tau_{q_{k}},N}\right)\sum_{a_{q_{1}},\ldots a_{q_{b}}=1}^{N-1}\Psi^{'}(\tau_{1},\ldots,\underbrace{a_{q_{1}}}_{q_{1}},\ldots,\underbrace{a_{q_{b}}}_{q_{b}},\ldots,\tau_{L})
		\end{align}
		where the notation
		\begin{equation}\label{notation-PSI}
			\Psi^{'}(\tau_{1},\ldots,\underbrace{a_{q_{1}}}_{q_{1}},\ldots,\underbrace{a_{q_{b}}}_{q_{b}},\ldots,\tau_{L})
		\end{equation}
		means that, for all $b\in \{0,\ldots,L\}$ and for all $1\leq q_{1}<\ldots<q_{b}\leq L$,
		\begin{equation}
			\begin{cases}
				\tau_{x}=a_{x}\qquad &\text{if}\quad x\in\{q_{1},\ldots,q_{b}\}\\
				\tau_{x}=\tau_{x}\qquad &\text{if}\quad \text{otherwise}
			\end{cases}\dt
		\end{equation}
		In the second equality we have rewritten the product by introducing the summation over $b$, the coordinates $q_{1},\ldots,q_{b}$ and the occupations variables $a_{q_{1}},\ldots,a_{q_{k}}$. In the last equality we have summed over all possible $\sigma\in \Omega^{'}$. Finally, using equation \eqref{interpretation-PsiP}, the result follows. 
\begin{remark}[Equilibrium case]
	Using equation \eqref{implicit-mass-function} with correlation computed in \eqref{exactCorrelations} in the equilibrium case ($\beta_{a}=\alpha_{a}$), we obtain 
	\begin{equation}
		\Psi(\bm{\tau})=\prod_{a=1}^{N}\alpha_{a}^{\sum_{x=1}^{L}\delta_{\tau_{x},a}}
	\end{equation}
As it has to be, this is the probability of the configuration $\bm{\tau}$ under the reversible measure \eqref{reversibleMeasure}, with site marginals distributed as $\text{Multinomial }(1,\alpha_{1},\ldots,\alpha_{N})$. 
\end{remark}
\subsection{The single-species case (SSEP)}
The computations made for the multispecies stirring process can be specialized to the single-species case $N=2$. In this situation we retrieve the SSEP, that was already solved with the quantum inverse scattering method in \cite{frassek2020eigenstates}. We report the Hamiltonians and the similarity transformations of SEEP: 
\begin{equation}
	H=H_{\text{left}}+\sum_{x=1}^{L-1}H_{x,x+1}+H_{\text{right}}
\end{equation}
where the bulk Hamiltonian densities are 
\begin{equation}
	H_{x,x+1}=\sum_{a,b=1}^{2}\left(e_{ab}\otimes e_{ba}-e_{bb}\otimes e_{aa}\right)\dt
\end{equation}
Here $e_{ab}$ are the basis elements of the fundamental representation of $gl(2)$. The boundary Hamiltonians are
\begin{equation}
	H_{\text{left}}=\begin{pmatrix}
		1-\alpha_{1}&\alpha_{1}\\
		\alpha_{2}&1-\alpha_{2}
	\end{pmatrix},\qquad H_{\text{right}}=\begin{pmatrix}
	1-\beta_{1}&\beta_{1}\\
	\beta_{2}&1-\beta_{2}
\end{pmatrix}\dt
\end{equation}
In this case, the similarity transformations read
\begin{equation}
	\mathcal{S}_{1}=\begin{pmatrix}
		1&0\\
		1&1
	\end{pmatrix},\qquad \mathcal{S}_{2}=\begin{pmatrix}
	1&-\beta_{1}\\
	0&1
\end{pmatrix}\dt
\end{equation}
Therefore, we have that $H^{'}$ has the following boundaries
\begin{equation}
	H^{'}_{\text{left}}=\begin{pmatrix}
		-1&\alpha_{1}\\
		0&0
	\end{pmatrix},\qquad H^{'}_{\text{right}}=\begin{pmatrix}
	-1&\beta_{1}\\
	0&0
\end{pmatrix}\co
\end{equation}
while the boundary Hamiltonians of $H^{''}$ are
\begin{equation}
	H^{''}_{\text{left}}=\begin{pmatrix}
		-1&\alpha_{1}-\beta_{1}\\
		0&0
	\end{pmatrix},\qquad H^{''}_{\text{right}}=\begin{pmatrix}
	-1&0\\
	0&0
\end{pmatrix}\dt
\end{equation}
 We only report the main results, since the computations are just a specific case of the ones done for the multispecies stirring process. 
\paragraph{Correlations of SSEP.}
We retrieve the closed formula for correlations in the non-equilibrium steady state found in \cite[(4.26)]{frassek2020eigenstates}:  
\begin{equation}\label{correlations-SSEP}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=1\}}\right]=\sum_{q=0}^{m}(\alpha-\beta)^{q}\beta^{m-q}\sum_{1\leq \ell_{1}<\ldots<\ell_{q}\leq m}\prod_{k=1}^{q}\frac{\left(L+1-x_{\ell_{k}}-q+k\right)}{(2+L-k)}\dt
\end{equation} 
\textbf{Proof of formula \eqref{correlations-SSEP}:} Starting from \eqref{exactCorrelations} we exploit indistinguishability of particles and we introduce $q=\sum_{k=1}^{m}c_{x_{k}}$. There are many values of $c$'s that give the same $q$, therefore we define $q$ %labels denoted by $\ell_{1},\ldots,\ell_{q}\in \{1,\ldots,m\}$ such that  
$1\leq \ell_{1}<\ell_{2}<\ldots<\ell_{q}\leq m$. To each of these label we associate the coordinates $x_{\ell_{1}},\ldots,x_{\ell_{q}}\in\{x_{1},\dots,x_{m}\}$ that are such $c_{x_{\ell_{k}}}=1$ for all $k\in \{1,\ldots,q\}$ and $c_{x}=0$ for $x\notin \{x_{\ell_{1}},\ldots,x_{\ell_{q}}\}$. Therefore, for any fixed $c_{x_{1}},\ldots,c_{x_{m}}$, we find $q$ and $\ell_{1},\ldots,\ell_{q}$ from which we fix $x_{\ell_{1}},\ldots,x_{\ell_{q}}$. Then, we obtain  
	\begin{equation}
		\prod_{k=1}^{m}\left(2+L-x_{k}-\sum_{j=k}^{m}c_{j}\right)^{c_{k}}=\prod_{k=1}^{q}\left(1+L-x_{\ell_{k}}-q+k\right)\dt
	\end{equation}
Indeed, in the product on the left hand side only $q$ terms at coordinates $x_{\ell_{1}},\ldots x_{\ell_{q}}$ do survive. Moreover, any fixed $x_{\ell_{k}}$ has $q-k$ occupied sites at its right.
Furthermore, using the properties of the factorials, we have that 
\begin{equation}\label{Gamma-product}
	\frac{(L+1-\sum_{k=1}^{m}c_{k})!}{(L+2)!}=\prod_{k=1}^{q}\frac{1}{(2+L-k)}\dt
\end{equation}
Varying all possible $c$'s in \eqref{exactCorrelations} is equivalent of varying all possible $q$ and all possible $\ell_{1},\ldots,\ell_{q}$, therefore we obtain 
\begin{equation}
	\sum_{c_{1}\ldots,c_{m}=0}^{m}\frac{(L+1-\sum_{k=1}^{m}c_{k})!}{(L+1)!}\prod_{k=1}^{m}\left(2+L-x_{k}-\sum_{j=k}^{m}\right)=\sum_{q=0}^{m}\sum_{1\leq \ell_{1}<\ldots<\ell_{q}\leq L}\prod_{k=1}^{q}\frac{(1+L-x_{\ell_{k}}-q+k)}{(2+L-k)}
\end{equation}
and \eqref{correlations-SSEP} follows. 
\begin{flushright}
	$\square$
\end{flushright}
\paragraph{The vector $|\Psi^{''}\rangle$ for SSEP.} The ground state of $H^{''}$ is 
\begin{equation}\label{psiII-SSEP}
	|\Psi^{''}\rangle=\sum_{q=0}^{L}\sum_{1\leq \ell_{1}<\ldots<\ell_{q}\leq L}(\alpha_{1}-\beta_{1})^{q}\prod_{k=1}^{q}\frac{\left(1+L-\ell_{k}-q+k\right)}{(2+L-k)}|\bm{\ell},q\rangle
\end{equation}
where $\bm{\ell}=(\ell_{1},\ldots,\ell_{q})$. \\ \\
\textbf{Proof of formula \eqref{psiII-SSEP}:} We specialize equation \eqref{resulEsteady} to the colour-blind case. We introduce $q=\sum_{x=1}^{L}\delta_{\tau_{x},1}$. For each $q\in \{1,\ldots,L\}$ we call $1\leq\ell_{1}<\ldots,\ell_{q}\leq L$ the coordinates where particles of type $1$ are present. Thus, arguing as in the proof of \eqref{correlations-SSEP} and using \eqref{Gamma-product} we have that 
\begin{align}
	&\frac{\left(1+\sum_{x=1}^{L}\delta_{\tau_{x},2}\right)!}{\left(1+L\right)!}\prod_{x=1}^{L}\left[(\alpha_{\tau_{x}}-\beta_{\tau_{x}})\left(1+\sum_{j=x}^{L}\delta_{\tau_{j},2}\right)\right]^{1-\delta_{\tau_{x},2}}\nonumber\\%=\lambda_{1}^{q}\prod_{x=1}^{L}\left(2+L-y_{x}-\sum_{j=x}^{L}\delta_{\tau_{j},1}\right)^{\delta_{\tau_{x},1}}
	=&\frac{\left(1+L-\sum_{x=1}^{L}(1-\delta_{\tau_{x},2})\right)!}{\left(1+L\right)!}\prod_{x=1}^{L}\left[(\alpha_{\tau_{x}}-\beta_{\tau_{x}})\left(2+L-x-\sum_{j=x}^{L}(1-\delta_{\tau_{j},2})\right)\right]^{1-\delta_{\tau_{x},2}}\nonumber\\
	=&(\alpha_{1}-\beta_{1})^{q}\prod_{k=1}^{q}\frac{\left(1+L-\ell_{k}-q+k\right)}{(2+L-k)}\dt
\end{align}
By varying $q$ and $\ell_{1},\ldots,\ell_{q}$, we have a one-to-one mapping with $\bm{\tau}$, therefore we obtain \eqref{psiII-SSEP}.
\begin{flushright}
	$\square$
\end{flushright}
\paragraph{The vector $|\Psi^{'}\rangle$ for SSEP.} The ground state of $H^{'}$ reads
\begin{equation}
	|\Psi^{'}\rangle=\sum_{\bm{\tau}\in\Omega^{'}}\Psi^{'}(\bm{\tau})|\bm{\tau}\rangle
\end{equation}
where 
\begin{equation}\label{PsiI-N2}
	\Psi^{'}(\bm{\tau})=\sum_{q=0}^{|\bm{\tau}|}(\alpha-\beta)^{q}\beta^{L-q}\sum_{1\leq \ell_{1}<\ldots<\ell_{q}\leq |\bm{\tau}|}\prod_{k=1}^{q}\frac{\left(L+1-x_{\ell_{k}}-q+k\right)}{(2+L-k)}\dt
\end{equation}
and where $|\bm{\tau}|=\sum_{x=1}^{L}\delta_{\tau_{x},1}$ . \\ \\
 \textbf{Proof of formula \eqref{PsiI-N2}:}  For any configuration $\bm{\tau}\in \Omega$ we denote by  $|\bm{\tau}|=\sum_{x=1}^{L}\delta_{\tau_{x},2}$ the number of occupied sites. Then, considering equation \eqref{ABS_1} we have the result by adapting the proof of \eqref{correlations-SSEP} in the following way. We select $q\in\{0,\ldots,|\bm{\tau}|\}$ particles among the occupied ones. We call $\ell_{1},\ldots\ell_{q}\in \{1,\ldots,|\tau|\}$ such that $\ell_{1}<\ell_{2}<\ldots<\ell_{q}$. These $(\ell_{k})_{k\in {1,\ldots,q}}$ are labels for the $q$ particles previously selected. For each $\ell_{k}$ we introduce $x_{\ell_{k}}$, i.e. the coordinate where the particle with label $\ell_{k}$ is placed. Thus, we have that  
\begin{equation}
	\prod_{x=1}^{L}\left(L+2-x-\sum_{j=x}^{L}c_{j}\right)^{c_{x}}=\prod_{k=1}^{q}\left(L+2-x_{\ell_{k}}-\sum_{j=x_{\ell_{k}}}^{L}c_{j}\right)=\prod_{k=1}^{q}\left(1+L-x_{\ell_{k}}-q+k\right)\dt
\end{equation}
It follows that
\begin{equation}
	\sum_{c_{1}=0}^{1}\ldots\sum_{c_{L}=0}^{1}\mathbbm{1}_{\{c_{1}+\cdots+c_{L}=q\}}	\prod_{x=1}^{L}\left(L+2-x-\sum_{j=x}^{L}c_{j}\right)^{c_{x}}=\sum_{1\leq \ell_{1}<\ldots<\ell_{q}\leq |\bm{\tau}|}\prod_{k=1}^{q}\left(1+L-x_{\ell_{k}}-q+k\right)\dt
\end{equation} 
Therefore, using \eqref{Gamma-product}  one finds $|\Psi^{'}\rangle$ for the SSEP.
\begin{flushright}
	$\square$
\end{flushright}
\paragraph{The vector $|\Psi\rangle$ for SSEP.}
In the colour-blind case we have that 
\begin{equation}\label{NESS-SSEP}
	|\Psi\rangle=\sum_{\bm{\tau}\in \Omega^{'}}\Psi(\bm{\tau})|\bm{\tau}\rangle
\end{equation}
where
\begin{equation}
	\Psi(\bm{\tau})=\sum_{b=0}^{L}(-1)^{b}\sum_{1\leq q_{1}<\ldots<q_{b}\leq L}\left(\prod_{k=1}^{q}\delta_{\tau_{q_{k}},2}\right)\sum_{s=0}^{|\bm{\tau}|+b}(\alpha-\beta)^{s}\beta^{L-s}\sum_{1\leq \ell_{1}<\ldots<\ell_{s}\leq |\bm{\tau}|}\prod_{k=1}^{s}\frac{\left(L+1-x_{\ell_{k}}-s+k\right)}{(2+L-k)}\dt
\end{equation}

\textbf{Proof of \eqref{NESS-SSEP}:} the proof follows from \eqref{implicit-PSI-PSIp} and from the coefficients $\Psi^{'}(\bm{\tau})$ computed in \eqref{PsiI-N2}. 
\begin{flushright}
	$\square$
\end{flushright}

\section{Extensions beyond the integrable chain}
\label{sec5}
In Sections~\ref{sectionIntegrabiliy} we have shown that, on the geometry of a chain and with hard-core interaction the model is exactly solvable using integrability.  
The aim of this section is to study the process when integrability is lost. We fix $\nu\geq 1$ and, for the sake of simplicity, we keep the geometry of a chain. We provide the following extensions: 
\begin{enumerate}
	\item \textit{Correlations} (Section~\ref{Subsection-ss-nonI}): exploiting duality, the steady state correlations can be expressed in terms of absorption probabilities of the dual particles. These probability are hard to find when $\nu>1$. We give 
	in Appendix~\ref{appendix-1pts-general} the expression for the one-point correlation.
	\item \textit{Instantaneous thermalization} (Section~\ref{Subsection-Thermalized}): starting from the multispecies stirring dynamics, we define the thermalized process, where particle are redistributed on each bond with a measure that is the reversible one conditioned to the conservation of the number of particles on this bond. Also for this thermalized process we prove duality. 	
	\item \textit{Reaction-diffusion process} (Section~\ref{Subsection-RD}): a further dynamics is added. Particle can now change type. The model is carefully constructed with the goal of having the average densities evolving as a closed system of $N-1$ coupled difference-differential equation, providing a generalization of the model studied in \cite{casini2022uphill}. Duality for the boundary driven process is also proved.
\end{enumerate}
In the rest of this section we consider a chain of length $L$, where two reservoirs are attached to sites $1$ and $L$. Each site can host $\nu\geq 1$ particles, whose type is chosen among the $N$ available ones. The graph of the dual process is still a chain, but with two extra-sites $0$ (connected to $1$) and $L+1$ (connected to $L$). The processes considered in this section will be denoted with same name $\left(\bm{n}(t)\right)_{t\geq0}$, their generator will be specified in each subsection. 
We use the notation of Section~\ref{sectionDuality} specialized to the one dimensional chain with $\nu >1$.
E.g. the state space is denoted by 
\begin{equation}\label{stateSpace-Chain}
	\Omega_{L}=\bigotimes_{x=1}^{L} \Omega_{x}\,.
\end{equation}
 The dual process will always be denoted by $(\bm{\xi}(t))_{t\geq 0}$ with the generator specified case by case. The dual configuration space is the one defined in \eqref{dualStateSpace} and adapted to the chain of size $L$ and with two extra-site $\{0,L+1\}$ attached to $\{1,L\}$ respectively. 
 We denote the space of the dual process by $\widetilde{\Omega}_{L}$, i.e.
\begin{equation}\label{dualSpace-chain}
	\widetilde{\Omega_{L}}=\mathbb{N}_{0}^{N}\otimes \Omega_{L}\otimes \mathbb{N}_{0}^{N}\dt
\end{equation}
In the following of this section we will denote by 
\begin{equation}
	\rho_{a}^{\text{left}}=\frac{\alpha_{a}}{|\alpha|},\qquad \qquad \rho_{a}^{\text{right}}=\frac{\beta_{a}}{|\beta|}\,.
\end{equation}
\subsection{Duality for a boundary driven chain}\label{Subsection-ss-nonI}
 As already pointed out the dual process eventually voids the chain by piling up particles of species $\{1,\ldots,N-1\}$ in the extra sites $0$ and $L+1$ and replacing them by types $N$ (the empty). This property allows to characterize the steady state non-equilibrium distribution (that is unique) by the probability that dual particles are absorbed in the extra-sites, called \textit{absorption probabilities}. This is a generalization for the non-integrable chain of the Corollary~\ref{Corolollary-ABS_Corr-abstract}. The process $\left(\bm{n}(t)\right)_{t\geq0}$ has generator
\begin{equation}
	\mathcal{L}=\mathcal{L}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}+\mathcal{L}_{\text{right}}\dt
\end{equation} 
where $\mathcal{L}_{x,x+1}$ is the edge generator defined in \eqref{edgeGenerator} acting on the bond $(x,x+1)$ and $\mathcal{L}_{\text{left}}$ and $\mathcal{L}_{\text{right}}$ are the generators \eqref{siteGenerator} acting on sites $\{1\}$ and $\{L\}$ with parameters $(\alpha_{a})_{a\in \{1,\ldots,N\}}$ and $(\beta_{a})_{a\in \{1,\ldots,N\}}$ respectively. 
The process is dual to $(\bm{\xi}(t))_{t\geq0}$, defined on the dual state space $\widetilde{\Omega}_{L}$, has generator
\begin{equation}
	\widetilde{\mathcal{L}}=\widetilde{\mathcal{L}}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}+\widetilde{\mathcal{L}}_{\text{right}}
\end{equation} 
where $\mathcal{L}_{x,x+1}$ is given by \eqref{Generator} and where the boundary generators $\widetilde{\mathcal{L}}_{\text{left}}$ and $\widetilde{\mathcal{L}}_{\text{right}}$ are given in \eqref{siteDualGenerator}: they absorb particles from sites $\{1\}$ and $\{L\}$ and put them at extra-sites $\{0\}$ and $\{L+1\}$ respectively. The duality function reads is \eqref{dualityElements} specified on the chain. 
We call $\mu_{\text{NESS}}$ the non-equilibrium stationary steady state of the the process $(\bm{n}(t))_{t\geq 0}$, we denote by $\mathbb{E}_{\mu_{\text{NESS}}}[\cdot]$ the expectation with respect to this measure and and we call $\bm{\mathcal{Y}} = (\mathcal{Y}^{1},\ldots,\mathcal{Y}^{L})$ the random vector with law $\mu_{\text{NESS}}$. For every $m\in\{1,\ldots,L\}$ we consider  $x_{1},\ldots,x_{m}\in\{1,\ldots,L\}$ and $a_{1},\ldots,a_{m}\in \{1,\ldots,N-1\}$. We introduce the dual configuration  $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$, meaning that 
\begin{equation}
	\begin{cases}
		\xi_{a}^{x}=\sum_{y=1}^{L}\sum_{b=1}^{N-1}\mathbbm{1}_{\{y=x_{k}\}}\mathbbm{1}_{\{b=a_{k}\}}\qquad &\text{if}\quad x=x_{k}\;\text{and} \; a=a_{k}\\
		\xi_{a}^{x}=0\qquad &\text{otherwise}
	\end{cases}\dt
\end{equation}
The duality function is \eqref{dualityElements}, adapted to the chain. Therefore, we have that
\begin{equation}
	D(\bm{n},\bm{\xi})= \frac{\prod_{k=1}^{m}n_{a_{k}}^{x_{k}}}{C(\nu,m)}\dt
\end{equation}
where $C(\nu,m)$ is a constant that depends on $m$ and on the maximal occupancy $\nu$. 
Therefore, up to this constant, $\mathbb{E}_{\mu_{\text{NESS}}}\left[D(\bm{n},\bm{\xi})\right]$ gives the $m-$points correlations at non-equilibrium steady state. The following equalities hold using ergodicity and duality:
\begin{equation}\label{ExptationSS}
	\begin{split}
		\frac{\mathbb{E}_{\mu_{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\mathcal{Y}_{a_{k}}^{x_{k}}=n_{a_{k}}^{x_{k}}\}}\right]}{C(\nu,m)}&=\lim_{t\to\infty}\mathbb{E}_{\bm{n}}\left[D(\bm{n}(t),\bm{\xi})\right]=\lim_{t\to\infty}\mathbb{E}_{\bm{\xi}}\left[D(\bm{n},\bm{\xi}(t))\right]
		\\&=
		\sum_{t_{1}=0}^{|\xi_{1}|}\ldots\sum_{t_{N}=0}^{|\xi_{N}|}\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{left}}\right)^{t_{a}}\left(\rho_{a}^{\text{right}}\right)^{|\xi_{a}|-t_{a}}\mathcal{P}_{x_{1},\ldots,x_{m}}(t_{1},\ldots,t_{N})
	\end{split}
\end{equation}
here we recall that, generalizing Corollary~\ref{Corolollary-ABS_Corr-abstract}, %that $\rho_{\alpha}^{L}=\frac{\alpha_{a}}{|\alpha|}$, $\rho_{\alpha}^{R}=\frac{\beta_{a}}{|\beta|}$ are the boundary densities and 
\begin{equation}\label{Pass} 
	\mathcal{P}_{x_{1},\ldots,x_{m}}(t_{1},\ldots,t_{N})	=\mathbb{P}\left(\bm{\xi}(\infty)=\sum_{a=1}^{N-1}\left(t_{a}\delta_{a}^{0}+(|\xi_{a}|-t_{a})\delta_{a}^{L+1}\right)| \bm{\xi}(0)=\bm{\xi}\right)
\end{equation}
are the absorption probabilities for the non-integrable chain, i.e. the probabilities that $t_{a}$ particles of species $a$ are absorbed at $\{0\}$. Here we have denoted $|\xi_{a}|=\sum_{x=1}^{L}\xi_{a}^{x}$.
 These correlations can be computed once the absorption probabilities are known. For example, we report in Appendix~\ref{appendix-1pts-general} the solution for the simplest case of one point correlations that is exactly solvable, since the single dual particle behaves as an independent random walker. The absorption probabilities for higher point correlations fulfill  difference equations.

\subsection{Instantaneous thermalization}\label{Subsection-Thermalized}
In this subsection we aim to extend to the multispecies stirring process the concept of \textit{instantaneous thermalization limit} (see \cite{carinci2013duality,KMP}). The thermalized process gives rise, for any pair nearest neighbour sites of the chain, to a redistribution of the total number of particles. The thermalized model that we present in this paper is obtained from the boundary driven model with generator \eqref{Generator} as follows: for each bond $(x,x+1)$ the total number of particles $\epsilon=n^{x}+n^{x+1}$ is redistributed according to the reversible measure, conditioned on the conservation of $\epsilon$. 
\subsubsection{The thermalization measure}
We introduce the reversible measure \eqref{reversibleMeasure} conditioned on the conservation of the particles on that bond
\begin{align}
	\nu_{TH}(r|\epsilon):&=\mathbb{P}\left(n^{x}=r|n^{x}+n^{x+1}=\epsilon\right)\nonumber
	\\&
	=\mathbb{P}\left((n_{1}^{x},\ldots,n_{N}^{x})=
	(r_{1},\ldots,r_{N})\;|\;(n_{1}^{x}+n_{1}^{x+1},\ldots,n_{N}^{x}+n_{N}^{x+1})=(\epsilon_{1},\ldots,\epsilon_{N})\right)\dt
\end{align}
Here $r=(r_{1},\ldots,r_{N})$ and  $\epsilon=(\epsilon_{1},\ldots,\epsilon_{N})$ are such that $\forall a\in\{1,\ldots,N\}$ we have $r_{a}\in \{0,\ldots,\nu\}$ and $\epsilon_{a}\in \{0,\ldots,2\nu\}$, under the constraints 
\begin{equation}
	\epsilon_{N}=2\nu-\sum_{a=1}^{N-1}\epsilon_{a},\qquad r_{N}=\nu-\sum_{a=1}^{N-1}r_{a}\dt
\end{equation}
 Using the conditional probability and the fact that the reversible measure \eqref{reversibleMeasure} is the product over the sites of multinomial distributions we have
\begin{equation}
	\begin{split}
\nu_{TH}(r|\epsilon)&=\frac{\mathbb{P}\left(\left\{n^{x}=r\right\}\,\cap\,\left\{n^{x}+n^{x+1}=\epsilon\right\}\right)}{\mathbb{P}\left(n^{x}+n^{x+1}=\epsilon\right)}=\frac{\mathbb{P}\left(n^{x}=r\right)\,\mathbb{P}\left(n^{x+1}=\epsilon-r\right)}{\mathbb{P}\left(n^{x}+n^{x+1}=\epsilon\right)}\dt
	\end{split}
\end{equation}
%Moreover, it holds that 
%\begin{equation}
%\text{Multinomial}(\nu,\rho_{1},\ldots,\rho_{N})+\text{Multinomial}(\nu,\rho_{1},\ldots,\rho_{N})=\text{Multinomial}(2\nu,\rho_{1},\ldots,\rho_{N})
%\end{equation}
Therefore the thermalization measure reads
\begin{equation}
	\begin{split}
	\nu_{TH}(r|\epsilon)=\frac{\binom{\nu}{r_{1},\ldots,r_{N}}\binom{\nu}{(\epsilon_{1}-r_{1}),\ldots,(\epsilon_{N}-r_{N})}}{\binom{2\nu}{\epsilon_{1}\ldots,\epsilon_{N}}}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}
	\end{split}
\end{equation}
where we denote the multinomial coefficient
\begin{equation}
	\binom{\nu}{n_{1},\ldots,n_{N}}=\frac{\nu!}{n_{1}!\cdots n_{N}!}\dt
\end{equation}
%provided that $n_{1}+\ldots+n_{N}=\nu$ and $n_{1},\ldots,n_{N}\in \mathbb{N}_{0}$. 
\subsubsection{Thermalized multispecies stirring process}
The thermalized multispecies stirring process is defined on the configuration space $\Omega$ introduced in \eqref{stateSpace-Chain} and its generator reads
\begin{equation}\label{thermalizedGenerator}
	\mathcal{L}^{TH}=\mathcal{L}_{\text{left}}^{TH}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{TH}+\mathcal{L}_{\text{right}}^{TH}
\end{equation}
where 
\begin{equation}\label{bulkThermalized}
	\begin{split}
	\mathcal{L}_{x,x+1}^{TH}f(\bm{n})=&\sum_{r_{1}=0}^{n_{1}^{x}+n_{1}^{x+1}}\ldots\sum_{r_{N}=0}^{n_{N}^{x}+n_{N}^{x+1}}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}
	\\&
	\left\{f(n^{1},\ldots,n^{x-1},r,n^{x}+n^{x+1}-r,n^{x+2},\ldots,n^{L})-f(\bm{n})\right\}\nu_{TH}(\bm{r}|\bm{\epsilon})
	\end{split}
\end{equation}
while
\begin{equation}\label{boundaryThermalized}
	\begin{split}
		\mathcal{L}_{\text{left}}^{TH}f(\bm{n})=&\sum_{r_{1}=0}^{\nu}\ldots\sum_{r_{N}=0}^{\nu}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}\left\{f(r,n^{2},\ldots,n^{L})-f(\bm{n})\right\}\\&
		\binom{\nu}{r_{1},\ldots, r_{N}}\prod_{a=1}^{N-1}\left(\frac{\alpha_{a}}{\alpha_{N}}\right)^{r_{a}}\left(\frac{\alpha_{N}}{|\alpha|}\right)^{\nu}
		\end{split}
	\end{equation}
and $\mathcal{L}_{\text{right}}^{TH}$ acting similarly on the site $L$ and with parameters $\beta$'s. 
%\begin{equation}
%	\begin{split}
%		\mathcal{L}_{\text{right}}^{TH}f(\bm{n})=&\sum_{r_{1}=0}^{\nu}\ldots\sum_{r_{N}=0}^{\nu}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}\left\{f(n^{1},\ldots,n^{L-1},r)-f(\bm{n})\right\}\\&
%		\binom{\nu}{r_{1},\ldots, r_{N}}\prod_{a=1}^{N-1}\left(\frac{\beta_{k}}{\beta_{N}}\right)^{r_{k}}\left(\frac{\beta_{N}}{|\beta|}\right)^{\nu}
%	\end{split}
%\end{equation}
In the thermalized bulk generator particles on a bond $x$ and $x+1$ are redistribute according to the measure $\nu_{TH}(r|\epsilon)$, while in the thermalized boundary generators the configuration at site $1$ ($L$) is replaced by $r$, according to a rate depending on $r$ itself and on the reservoirs parameters. 
\subsubsection{The dual thermalized process}\label{subsection-dualTHM}
Here we state the duality result for the thermalized multispecies stirring process with boundary driving.
\begin{proposition}[Duality for the thermalized multispecies stirring process]\label{proposition-dualityTHM}
The thermalized multispecies stirring process $(\bm{n}(t))_{t\geq 0}$ on the state space $\Omega_{L}$, with generator $\mathcal{L}^{TH}$ defined in \eqref{thermalizedGenerator} is dual to the process $(\bm{\xi}(t))_{t\geq 0}$ on the enlarged configuration space $\widetilde{\Omega}_{L}$, with dual generator
\begin{equation}
	\widetilde{\mathcal{L}}^{TH}=\widetilde{\mathcal{L}}_{\text{left}}^{TH}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{TH}+\widetilde{\mathcal{L}}_{\text{right}}^{TH}
\end{equation}
where $\mathcal{L}^{TH}_{x,x+1}$ is \eqref{bulkThermalized} and $\widetilde{\mathcal{L}}_{\text{left}}^{TH}$ and $\widetilde{\mathcal{L}}_{\text{right}}^{TH}$ are absorbing with rate $1$:
\begin{align}
	\widetilde{\mathcal{L}}_{\text{left}}^{TH}f(\bm{\xi})&=\left\{f(\xi^{0}+\xi^{1},0,\xi^{2},\ldots,\xi^{L})-f(\bm{\xi})\right\}\nonumber\\ \widetilde{\mathcal{L}}_{\text{right}}^{TH}f(\bm{\xi})&=\left\{f(\xi^{0},\ldots,\xi^{L-1},0,\xi^{L}+\xi^{L+1})-f(\bm{\xi})\right\}\dt
\end{align}
The duality function is the same of the multispecies stirring process, i.e. 
\begin{equation}
	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{left}}\right)^{\xi_{a}^{0}}\prod_{x=1}^{L}\frac{(\nu-\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{right}}\right)^{\xi_{a}^{L+1}}\dt
\end{equation}
(cf. \eqref{dualityElements}).
\end{proposition}
This dual process is absorbing since the dual boundary generator $\mathcal{L}_{\text{left}}^{TH}$ ($\mathcal{L}_{\text{left}}^{TH}$) remove all dual particles at site $1$ ($L$) and put these particles in the extra-site $0$ ($L+1$) with rate 1.
The proof of Proposition~\ref{proposition-dualityTHM} is standard. For the sake of completeness we report it in Appendix~\ref{appendix-dualityThermalized}.
\subsection{Multispecies stirring process with reaction} \label{Subsection-RD}
We define the average density occupation of species $a$ at site $x$ as
\begin{equation}
	\rho_{a}(x,t)=\mathbb{E}_{\mu_{0}}\left[n_{a}^{x}(t)\right]
\end{equation}
where $\mathbb{E}_{\mu_{0}}$ denotes the expectation with respect to the law of the Markov process $(\bm{n}(t))_{t\geq 0}$ initialized with arbitrary initial measure $\mu_{0}$. Calling
$\mathcal{L}^{rd}$ the generator of the process, these average occupations evolves according to the evolution equation
\begin{equation}\label{EE-density}
	\frac{d \rho_{a}(x,t)}{dt}=\mathbb{E}_{\mu_{0}}\left[\mathcal{L}^{rd}n_{a}^{x}(t)\right]\dt
\end{equation} 
In \cite{casini2022uphill} it has been defined a process whose evolution equation for the average occupation $\rho_{a}(x,t)$ of species $a\in \{1,2\}$ is given by a difference-differential equation with two discrete Laplacians (one for each species, multiplied by a proper diffusivity constant) and a linear reaction term (mutation), i.e.
\begin{equation}\label{2SpeciesRD}
	\frac{d}{dt} \rho_{a}(x,t)=\sigma_{11}\Delta_{L}\rho_{a}(x,t)+\sigma_{12}\Delta_{L}\rho_{\overline{a}}(x,t)+\Upsilon\left(\rho_{\overline{a}}(x,t)-\rho_{a}(x,t)\right)
\end{equation} 
for $ x\in\{1,\ldots,L\}\quad a\in\{1,2\}$  and where $\overline{a}=1$ if $a=2$ and $\overline{a}=1$ if $a=2$. Here $\Delta_{L}\rho(x,t)=\rho_{a}(x+1,t)+\rho_{a}(x-1,t)-2\rho_{a}(x,t)$.
This equation is endowed with boundary conditions $\rho_{a}^{\text{left}},\rho_{\overline{a}}^{\text{left}}$ on the left and $\rho_{a}^{\text{right}},\rho_{\overline{a}}^{\text{right}}$ on the right. 
The generator of this model consists in a sum of three Markov generators: a stirring, a stirring-mutation and a pure mutation one (see section 4 of \cite{casini2022uphill}). For the lattice $\mathbb{Z}$, self-duality has been showed and the duality function is exactly \eqref{elementsABS} in case of hard-core exclusion (see section 5 of \cite{casini2022uphill}). In view of what have been proved in the previous sections of this work, some questions arise naturally:
\begin{enumerate}
		\item Is it possible to extend the results of \cite{casini2022uphill} to a case with arbitrary number of species $N\in \mathbb{N}$ and arbitrary maximal site occupation $\nu\in \mathbb{N}$?
			\item Is it possible to extend the duality result to the boundary-driven case in order to have an absorbing dual process, using the same ${gl}(N)$ Lie algebraic structure?
\end{enumerate}
Therefore, we aim to find a process whose average occupations $\rho_{a}(x,t)$ for any species $a\in\{1,\ldots,N-1\}$ evolve as a system of difference-differential equation given by 
\begin{equation}\label{NSpeciesRD}
	\frac{d}{dt} \rho_{a}(x,t)=\nu\sigma_{11}\Delta_{L}\rho_{a}(x,t)+\nu\sigma_{12}\sum_{b=1,b\neq a}^{N-1}\Delta_{L}\rho_{b}(x,t)+\Upsilon\sum_{b=1,b\neq a}^{N-1}\left(\rho_{b}(x,t)-\rho_{a}(x,t)\right)\dt
\end{equation} 
 We assume that the boundary conditions of the above difference differential system are given and denoted by $\rho_{a}^{\text{left}}$ and $\rho_{a}^{\text{right}}$ for all $a\in\{1,\ldots,N-1\}$ at left and right respectively, further assuming that $\sum_{a=1}^{N}\rho_{a}^{\text{left}}=\sum_{a=1}^{N}\rho_{a}^{\text{right}}=\nu$.
 We construct the process on a chain in order to obtain \eqref{NSpeciesRD} for the evolution of the average occupation variable of type $a$. Once we have found such a process, we construct the Lie algebraic description and we use it to prove absorbing duality.
\subsubsection{The reaction-diffusion process on a chain}
We consider the geometry of a chain of length $L$, where we assume that a reservoir with parameters $(\alpha_{a})_{a\in\{1,\ldots,N\}}$ is attached at site $1$ and a reservoir with parameters $(\beta_{a})_{a\in\{1,\ldots,N\}}$ is attached at site $L$. The state space is $\Omega$ given in \eqref{stateSpace-Chain} and the generator reads
\begin{equation}\label{RDGenerator}
	\mathcal{L}^{rd}=\mathcal{L}_{\text{right}}^{rd}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{rd}+\mathcal{L}_{\text{right}}^{rd}
\end{equation}
where
\begin{equation}\label{edgeGeneratorRD}
	\mathcal{L}_{x,x+1}^{rd}=\nu \sigma_{11}\mathcal{L}_{x,x+1}+\nu \sigma_{12}\sum_{c=1}^{N-2}\mathcal{L}_{x,x+1}^{c}+(\Upsilon-2\nu\sigma_{12})\mathcal{L}_{x,x+1}^{m}
\end{equation}
where $\mathcal{L}_{x,x+1}$ is the edge-generator \eqref{edgeGenerator} of the multispecies stirring process and, for any function $f:\Omega\to \mathbb{R}$ and for any $c\in \{1,\ldots,N-2\}$ we introduce the \textit{stirring-mutation} generator
\begin{equation}
	\mathcal{L}_{x,x+1}^{c}f(\bm{n})=\sum_{a,b=1}^{N}n_{a}^{x}n_{b}^{x+1}\left(f(\bm{n}-\bm{\delta}_{a}^{x}+\bm{\delta}_{h_{c}(b)}^{x}+\bm{\delta}_{h_{c}(a)}^{x+1}-\bm{\delta}_{b}^{x+1})-f(\bm{n})\right)
\end{equation}
with the mapping 
\begin{equation}
	\begin{split}
		h_{c}(\cdot):\{1,\ldots,N\}&\to\{1,\ldots,N\}\\
		k&\to h_{c}(k)
	\end{split}
\end{equation}
defined as
\begin{equation}
	h_{c}(k)=\begin{cases}
		k+c \quad &\text{if}\quad k+c< N \\
		k+c-N+1\quad &\text{if}\quad k+c\geq N\quad \text{and}\quad k\neq N\\
		N\quad &\text{if}\quad k=N
	\end{cases}\dt
\end{equation}
Observe that $h_{c}(\cdot)$ is surjective and injective, and thus invertible.

Moreover, we define the \textit{pure-mutation} generator
\begin{equation}
	\mathcal{L}_{x,x+1}^{m}f(\bm{n})=\sum_{a,b=1}^{N-1}n_{a}^{x}\left(f(\bm{n}-\bm{\delta}_{a}^{x}+\bm{\delta}_{b}^{x})-f(\bm{n})\right)\dt
\end{equation}
The left boundary generator $\mathcal{L}_{\text{left}}^{rd}$ is given by \eqref{siteGenerator} at site $1$,
%reads
%\begin{equation}
%	\mathcal{L}_{\text{left}}^{rd}f(\bm{n})=\sum_{a,b=1}^{N}\alpha_{a}^{x}n_{b}^{x}\left(f(\bm{n}-\bm{\delta}_{b}^{x}+\bm{\delta}_{a}^{x})-f(\bm{n})\right)
%\end{equation}
where the $\alpha$'s parameters are defined in function of the given boundary condition of the difference-differential system:
\begin{align}\label{boundaryParamRD}
		\alpha_{a}&=\sigma_{11}\rho_{a}^{\text{left}}+\sigma_{12}\sum_{b=1\,:\,b\neq a}^{N-1}\rho_{b}^{\text{left}}\qquad \text{if} \quad a\in \{1,\ldots,N-1\}\nonumber\\
		\alpha_{N}&=\nu\left(\sigma_{11}+(N-2)\sigma_{12}\right)-\sum_{a=1}^{N-1}\alpha_{a}\dt
\end{align}
As a consequence, from the above conditions we can define the boundary values in terms of the reservoir parameters as
\begin{equation}
	\rho_{a}^{\text{left}}=\frac{\alpha_{a}\sigma_{11}-\sigma_{12}\sum_{b=1\,:\,b\neq a}^{N-1}\alpha_{b}}{(\sigma_{11}-(N-2)\sigma_{12})(\sigma_{11}+(N-2)\sigma_{12})}\qquad \rho_{N}^{\text{left}}=\frac{\alpha_{N}}{(\sigma_{11}+(N-2)\sigma_{12})}
\end{equation}
for every $a\in\{1,\ldots,N\}$. Moreover, with this choice we have that $\sum_{a=1}^{N}\alpha_{a}=\nu(\sigma_{11}+(N-1)\sigma_{12})$ and $\sum_{a=1}^{N}\rho_{a}^{\text{left}}=\nu$. 
The boundary generator $\mathcal{L}_{\text{right}}^{rd}$ is defined analogously, but the $\alpha$'s are replaced by $\beta$'s and the boundary values $\rho_{a}^{\text{right}}$.
%\begin{remark}: we stress that, up to a specific choice of the $\alpha$'s ($\beta$'s), the left (right) boundary generator $\mathcal{L}_{\text{left}}^{rd}$ ($\mathcal{L}_{\text{right}}^{rd}$) has the form of the one of the multispecies stirring process \eqref{siteGenerator}. 
%\end{remark}
\paragraph{Action on the occupation variable}
We only report the final result. The details are written in Appendix~\ref{appendix-RD}. For arbitrary site $x\in \{2,\ldots,L-1\}$, the action on the occupation variable $f(\bm{n})=n_{a}^{x}$ of particle of type $a\in \{1,\ldots,N-1\}$ at site $x$ reads
\begin{equation}\label{DifferenceEquation}
	\mathcal{L}^{rd}n_{a}^{x}=\nu\sigma_{11}\Delta_{L}n_{a}^{x}+\nu\sigma_{12}\sum_{b=1\,:\,b\neq a}^{N-1}\Delta_{L}n_{b}^{x}+\Upsilon\sum_{b=1\,:\,b\neq a}^{N-1}(n_{b}^{x}-n_{a}^{x})
\end{equation}
where $\Delta_{L}n_{a}^{x}= n_{a}^{x+1}+n_{a}^{x-1}-2n_{a}^{x}$. For $x=1$ and for all $a\in \{1,\ldots,N-1\}$ we have 
\begin{equation}\label{actionBDLine}
	\mathcal{L}^{rd}n_{a}^{1}=\sigma_{11}\nu \left(\rho_{a}^{\text{left}}-2n_{a}^{1}+n_{a}^{2}\right)+\sigma_{12}\nu\sum_{b=1\,:\,b\neq a}^{N-1}\left(\rho_{b}^{\text{left}}-2n_{b}^{1}+n_{b}^{2}\right)+\Upsilon\sum_{b=1\,:\,b\neq a}^{N-1}\left(n_{b}^{1}-n_{a}^{1}\right)\dt
\end{equation}
On the occupation variable $n_{a}^{L}$ the action is similar. Therefore, using the evolution equation \eqref{EE-density} for the average densities and  the action of $\mathcal{L}^{rd}$ expressed by \eqref{DifferenceEquation} and \eqref{actionBDLine} we obtain the desired system of $N-1$ difference-differential equations \eqref{NSpeciesRD}. For these reasons, this model is a generalization to $N$ species and to maximal occupation $\nu$ of the one studied in \cite{casini2022uphill}. 
%This answers the first question we addressed to at the beginning of this section. 
\subsubsection{Reversible measure for the reaction-diffusion process (equilibrium)} 
The process described by the generator \eqref{RDGenerator} is reversible with respect to the homogeneous product measure \begin{equation}\label{reversibleMeasureRD}
	\Lambda_{\text{rev}}=\bigotimes_{x=1}^{L}\Lambda_{\text{rev}}^{x}
\end{equation}
when for any $a\in\{1,\ldots,N\}$
\begin{equation}\label{reversibilityConditionRD}
	\alpha_{a}=\beta_{a}\qquad\text{and}\qquad \alpha_{a}=\alpha_{b}=\alpha \quad \forall a,b\in \{1,\ldots,N-1\}\dt
\end{equation}
This measure has marginals $\Lambda_{\text{rev}}^{x}$ given by 
\begin{equation}
	\Lambda^{x}_{\text{rev}}\sim \text{Multinomial}\left(\twoj,p,\ldots,p,p_{N}\right)
\end{equation}
where 
\begin{equation}
	p=\frac{\alpha}{\alpha_{N}+(N-1)\alpha}\qquad \text{and}\qquad p_{N}=\frac{\alpha_{N}}{\alpha_{N}+(N-1)\alpha}
\end{equation}
Namely,
\begin{equation}
	\Lambda_{\text{rev}}^{x}(n^{x})=\frac{\nu!}{\prod_{a=1}^{N}n_{a}^{x}!}p^{\sum_{a=1}^{N-1}n_{a}^{x}}p^{n_{N}^{x}}
\end{equation}
This can be proved by imposing the detailed balance conditions for the bond $(x,x+1)$ and for the boundaries $\{1,L\}$.\\

\subsubsection{Duality for the reaction-diffusion process}
In this section we formulate duality for the multispecies reaction diffusion process. This answer to the second question addressed at the beginning of Section~\ref{Subsection-RD}.  The dual process is defined on the enlarged state space $\widetilde{\Omega}_{L}$.  
\begin{proposition}[Duality for the reaction-diffusion process]\label{propositin-duality-RD}
The reaction-diffusion multispecies stirring process $(\bm{n}(t))_{t\geq 0}$, on the state space $\Omega_{L}$, with generator $\mathcal{L}^{rd}$ defined in \eqref{RDGenerator} is dual to the process $(\bm{\xi}(t))_{t\geq 0}$ on the state space $\widetilde{\Omega}_{L}$ with dual generator
\begin{equation}\label{DualGeneratorRD}
	\widetilde{\mathcal{L}}_{\text{left}}^{rd}=\widetilde{\mathcal{L}}^{rd}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{rd}+\widetilde{\mathcal{L}}_{\text{right}}^{rd}
\end{equation}
where 
$\mathcal{L}_{x,x+1}^{rd}$ is defined in \eqref{edgeGeneratorRD} and, $\widetilde{\mathcal{L}}_{\text{left}}^{rd}$ and $\widetilde{\mathcal{L}}_{\text{right}}^{rd}$ are the absorbing dual generators $\widetilde{\mathcal{L}}_{x}$ defined in \eqref{siteDualGenerator} and acting at sites $1$ and $L$, with the specific choice of the reservoirs parameters \eqref{boundaryParamRD}. The duality function
is the same of the stirring process, i.e.
\begin{equation}
	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{left}}\right)^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\right)\prod_{a=1}^{N-1}\left(\rho_{a}^{\text{right}}\right)^{\xi_{a}^{L+1}}
\end{equation}
(cf. \eqref{dualityElements}).
\end{proposition} 
%\begin{equation}
%	\widetilde{\Omega}=\mathbb{N}_{0}^{N}\times  \bigtimes_{x\in V} \Omega_{x}\times \mathbb{N}_{0}^{N}
%\end{equation}
%We associate an ``extra site'', denoted by $0$ and $L+1$ attached to $1$ and $L$ respectively,
%where dual particles will accumulate in the course of time. 
%and we denote its configurations by $\bm{\xi} \in \widetilde\Omega$.
%\begin{equation}
%	\bm{\xi}=\left(\xi_{1}^{0},\ldots,\xi_{N}^{0}\right)\otimes\left(\xi_{1}^{x},\ldots,\xi_{N}^{x}\right)^{\otimes L}\otimes \left(\xi_{1}^{L+1},\ldots,\xi_{N}^{L+1}\right)
%\end{equation}
%where the component $\xi_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
%and the component $\xi_{a}^{0},\xi_{a}^{L+1}$  give the number at the extra sites. \\
%where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
%\begin{equation}
%	\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}
%\end{equation}

\paragraph{Proof of Proposition~\ref{propositin-duality-RD}:}
we only prove duality for the bulk generator, since the boundary generators are the same of the multispecies stirring process. In the spirit of Section~\ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the ${gl}(N)$ Lie algebra as 
\begin{equation}
	H^{rd}=H_{\text{left}}+\sum_{x=1}^{L-1}\left(\sigma_{11}\mathcal{H}_{x,x+1}+\sigma_{12}\sum_{c=1}^{N-2}\mathcal{H}_{x,x+1}^{c}+(\Upsilon-2\sigma_{12})\mathcal{H}_{x,x+1}^{m}\right)+H_{\text{right}}\dt
\end{equation}
Here, $\mathcal{H}_{x,x+1}$ is  \eqref{edgeHamiltonian}  and $H_{\text{left}},H_{\text{right}}$ is \eqref{siteHamiltonian} (with the choice of the boundary parameters done in \eqref{boundaryParamRD}). We, then have that 
\begin{equation}
	\mathcal{H}_{x,x+1}^{c}=\sum_{a,b=0}^{N}\left(E_{h_{c}(a)b}\otimes E_{h_{c}(b)a}-E_{aa}\otimes E_{bb}\right)
\end{equation}
and 
\begin{equation}
	\mathcal{H}_{x,x+1}^{m}=\sum_{a,b=1}^{N}\left(E_{ba}\otimes \mathbbm{1}-E_{aa}\otimes \mathbbm{1}\right)\dt
\end{equation}
We prove that, for arbitrary $x\in\{1,\ldots,L\}$,
\begin{align}
	&(\mathcal{H}_{x,x+1}^{c})^{T}D=D\mathcal{H}_{x,x+1}^{c}\qquad \forall c\in \{1,\ldots,N-2\}\label{cDualityRelation}\\
	&(\mathcal{H}_{x,x+1}^{m})^{T}D=D\mathcal{H}_{x,x+1}^{m}\label{mDualityRelation}\dt
\end{align}
Since the matrix $R_{x}$ is the same as in Section~\ref{sectionDuality}, we shall show that $\mathcal{H}_{x,x+1}^{c}$ and $\mathcal{H}_{x,x+1}^{m}$ commute with $\exp{(E^{x})}\exp{(E^{x+1})}$, where $E$ is defined in \eqref{EquationEx}. Using \eqref{eq:comgl}, the bilinearity and associativity of the Kronecker product and the bilinearity of the brackets we obtain 
\begin{align*}
	&\left[\left(\sum_{a=1}^{N-1}E_{aN}^{x}\otimes\sum_{a=1}^{N-1}E_{aN}^{x+1}\right), \sum_{d,\ell=1}^{N}\left(E_{h_{c}(\ell)d}^{x}\otimes E_{h_{c}(d)\ell}^{x+1} -E_{dd}^{x}\otimes E_{\ell\ell}^{x+1}\right)\right]
	\\=&
%	\sum_{d,k=1}^{N-1}\sum_{a,b=1}^{N}\left\{\left((E_{da}^{x}\delta_{h_{c}(b)N}-E_{h_{c}(b)N}^{x}\delta_{ad})-(E_{da}^{x}\delta_{aN}-E_{aN}^{x}\delta_{da})\right)\right. \\ \otimes&\left. \left((E_{kb}^{x+1}\delta_{h_{c}(a)N}-E_{h_{c}N}^{x+1}\delta_{kb})-(E_{kb}^{x+1}\delta_{bN}-E_{bN}^{x+1}\delta_{kb})\right)\right\}
%	\\=&
%	\sum_{d,k=1}^{N-1}\sum_{a,b=1}^{N}\left\{\left(E_{da}^{x}-E_{h_{c}(b)N}^{x}-E_{da}^{x}+E_{aN}^{x}\right)\otimes\left(E_{kb}^{x+1}-E_{h_{c}(a)N}^{x+1}-E_{kb}^{x+1}+E_{bN}^{x+1}\right)\right\}
%	\\=&
	\sum_{a,b=1}^{N-1}\sum_{d,\ell=1}^{N}\left\{\left(E_{dN}^{x}-E_{h_{c}(\ell)N}^{x}\right)\otimes \left(E_{dN}^{x+1}-E_{h_{c}(\ell)N}^{x+1}\right)\right\}=0\dt
\end{align*}
In the last equality we used the fact that map $h_{c}(\cdot)$ is surjective. 
%\paragraph{Duality for $H_{x,x+1}^{m}$} 
Concerning the commutator between $\mathcal{H}_{x,x+1}^{m}$ the proof is similar and gives
\begin{equation}
	\left[\sum_{a=1}^{N-1}E_{aN}^{x}\otimes \sum_{b=1}^{N-1}E_{bN}^{x},\;\sum_{d,\ell=1}^{N}\left(E_{\ell a}^{x}\otimes \mathbbm{1}^{x+1}-E_{dd}^{x}\otimes \mathbbm{1}^{x+1}\right)\right]=0\dt
\end{equation}
\begin{flushright}
	$\square$
\end{flushright}

\appendix
\section*{Appendix}

\section{One point correlations for $\nu\geq 1$}\label{appendix-1pts-general}
We take $m=1$ and we fix a site $x\in\{1,\ldots,L\}$ and a species $a\in \{1,\ldots,N-1\}$. We consider the dual configuration $\bm{\xi}=\delta_{a}^{x}$, then 
\begin{equation}
	\nu D(\bm{n},\delta_{a}^{x})=n_{a}^{x}\dt
\end{equation}
The average occupation reads
\begin{equation}\label{OnePtsCORR}
	\begin{split}
			\mathbb{E}_{\mu_{\text{NESS}}}\left[\mathbbm{1}_{\{\mathcal{Y}_{a}^{x}=n_{a}^{x}\}}\right]&=\nu\lim_{t\to\infty}\mathbb{E}_{\bm{n}}\left[D(\bm{n}(t),\delta_{a}^{x})\right]\\&=\nu\sum_{t_{a}=0}^{1}\left(\rho_{a}^{\text{left}}\right)^{t_{s}}\left(\rho_{s}^{\text{right}}\right)^{1-t_{s}}\mathbb{P}_{\delta_{a}^{x}}\left(\xi_{\infty}=t_{a}\delta_{0}+(1-t_{a})\delta_{L+1}\right)\dt
	\end{split}
\end{equation}
To explicitly find the result, we aim to determine $\mathcal{P}(1,\delta_{a}^{x})=\mathbb{P}_{\delta_{a}^{x}}\left(\xi_{\infty}=t_{a}\delta_{a}^{0}+(1-t_{a})\delta_{a}^{L+1}\right)$. This is the probability of a random walker on the chain $\{1,\ldots,L\}$ with absorbing boundaries, therefore, it fulfils the following equations discrete Laplace equation
\begin{equation}\label{gamblers-ruin}
	\begin{cases}
		\Delta_{L}\mathcal{P}(1,\delta_{a}^{x})=0\\
		\mathcal{P}(1,\delta_{a}^{1})=\frac{|\alpha|}{1+|\alpha|}+\mathcal{P}(1,\delta_{a}^{2})\frac{1}{1+|\alpha|}\\
		\mathcal{P}(1,\delta_{a}^{L})=\mathcal{P}(1,\delta_{a}^{L-1})\frac{1}{1+|\beta|}
	\end{cases}
\end{equation}
where $\Delta_{L}$ is the discrete Laplace operator on the chain of length $L$. The equations \eqref{gamblers-ruin} can be solved and using $\mathcal{P}(1,\delta_{a}^{x})=1-\mathcal{P}(0,\delta_{a}^{x})$ we obtain 
\begin{equation}
	\mathbb{E}_{\mu_{\text{NESS}}}\left[\mathbbm{1}_{\{\mathcal{Y}_{a}^{x}=n_{a}^{x}\}}\right]=\nu\frac{\alpha_{a}(L|\beta|-|\beta|x+1)+\beta_{a}(|\alpha|x+1-|\alpha|)}{|\alpha||\beta|L-|\alpha||\beta|+|\alpha|+|\beta|}\dt
\end{equation}
\begin{remark}: In case $|\alpha|=|\beta|=\nu=1$ (when the chain is integrable) we have 
\begin{equation}
		\mathbb{E}\left[\mathbbm{1}_{\{Y_{x}=a\}}\right]=\frac{\alpha_{a}(L-x+1)+\beta_{a}x}{L+1}
\end{equation}
that is the same result obtained in \eqref{one-pts-corr}.  
\end{remark}
%Already for two-point correlations the computations become more involving and hard to solve.
\section{Proof of duality of thermalized multispecies stirring}\label{appendix-dualityThermalized}
The proof of the duality consists in showing
\begin{align}
	\left(\mathcal{L}_{x,x+1}^{TH}D(\cdot,\bm{\xi})\right)(\bm{n})&=\left(\widetilde{\mathcal{L}}_{x,x+1}^{TH}D(\bm{n},\cdot)\right)(\bm{\xi})\qquad x\in \{1,\ldots,L-1\}\label{bulkDualityTH}\\
	\left(\mathcal{L}_{\text{left}}^{TH}D(\cdot,\bm{\xi})\right)(\bm{n})&=\left(\widetilde{\mathcal{L}}_{\text{left}}^{TH}D(\bm{n},\cdot)\right)(\bm{\xi})\label{boundaryDualityTH}
\end{align}
Duality for the right boundary is similar. \\
Equation \eqref{bulkDualityTH} follows from the proof of the edge duality for the multispecies stirring process. To prove \eqref{boundaryDualityTH} we directly apply the generators on the duality function. For the sake of notation, we introduce 
\begin{equation}
	d(n^{x},\xi^{x}):=\frac{(\nu-\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}
\end{equation}
First, we act with $\mathcal{L}_{\text{left}}^{TH}$ on $D(\bm{n},\bm{\xi})$. By writing explicitly $\rho_{a}^{L}=\frac{\alpha_{a}}{|\alpha|}$, this action gives
\begin{align}\label{ProvaBordoTH}
	\left(\mathcal{L}_{\text{left}}^{TH}D(\cdot,\bm{\xi})\right)(\bm{n})=\nonumber&\sum_{r_{1}=0}^{\nu}\ldots\sum_{r_{N}=0}^{\nu}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}\left\{\left(\prod_{a=1}^{N-1}\left(\frac{\alpha_{a}}{|\alpha|}\right)^{\xi_{a}^{0}}\right)\frac{\left(\nu-\sum_{c=1}^{N-1}\xi_{c}^{1}\right)!}{\nu!}\right.
	\\&\nonumber
	\left.\left(\prod_{b=1}^{N-1}\frac{r_{b}!}{(r_{b}-\xi^{1}_{b})!}\right)\left(\prod_{x=2}^{L}d(n^{x},\xi^{x})\right)\left(\prod_{d=1}^{N-1}(\rho_{d}^{R})^{\xi_{d}^{L+1}}\right)-D(\bm{n},\bm{\xi})\right\}
	\\&
	\frac{\nu!}{r_{1}!\cdots r_{N}!}\prod_{k=1}^{N-1}\left(\frac{\alpha_{k}}{\alpha_{N}}\right)^{r_{k}}\left(\frac{\alpha_{N}}{|\alpha|}\right)^{\nu}
\end{align}
We consider the first addend in the curly bracket of the \eqref{ProvaBordoTH}. We add and remove $\xi_{a}^{1}$ at the exponent of $\left(\frac{\alpha_{a}}{|\alpha|}\right)$ and then we multiplying this first addend by $\left(\frac{\alpha_{N}}{|\alpha|}\right)^{-(r_{1}^{1}+\ldots+r_{N-1})+(r_{1}^{1}+\ldots+r_{N-1})}$. Thus, we have
\begin{align}\label{pezzo1}
	&\sum_{r_{1}=0}^{\nu}\ldots\sum_{r_{N}=0}^{\nu}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}\nonumber\\
	&\left\{\left(\prod_{a=1}^{N-1}\left(\frac{\alpha_{a}}{|\alpha|}\right)^{\xi_{a}^{0}-\xi_{a}^{1}+\xi_{a}^{1}}\right)\frac{\left(\nu-\sum_{c=1}^{N-1}\xi_{c}^{1}\right)!}{\nu!}\left(\prod_{b=1}^{N-1}\frac{r_{b}!}{(r_{b}-\xi^{1}_{b})!}\right)\left(\prod_{x=2}^{L}d(n^{x},\xi^{x})\right)\left(\prod_{d=1}^{N-1}(\rho_{d}^{R})^{\xi_{d}^{L+1}}\right)\right\}\nonumber\\
	&	\frac{\nu!}{r_{1}!\cdots r_{N}!}\left(\prod_{k=1}^{N-1}\left(\frac{\alpha_{k}}{\alpha_{N}}\right)^{r_{k}}\right)\left(\frac{\alpha_{N}}{|\alpha|}\right)^{-(r_{1}^{1}+\ldots+r_{N-1})+(r_{1}^{1}+\ldots+r_{N-1})}\nonumber\\=&
	\left(\prod_{k=1}^{N-1}\left(\frac{\alpha_{k}}{|\alpha|}\right)^{\xi_{k}^{0}+\xi_{k}^{1}}\right)\sum_{r_{1}=0}^{\nu}\ldots\sum_{r_{N}=0}^{\nu}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}	\frac{\left(\nu-\xi_{1}^{1}-\ldots-\xi_{N-1}^{1}\right)!}{(r_{1}-\xi_{1}^{1})!\cdots(r_{N-1}-\xi_{N-1}^{1})!(\nu-r_{1}-\ldots-r_{N-1})!}\nonumber\\&
	\left(\prod_{\ell=1}^{N-1}\left(\frac{\alpha_{\ell}}{|\alpha|}\right)^{r_{\ell}-\xi_{\ell}^{1}}\right)\left(\frac{\alpha_{N}}{|\alpha|}\right)^{\nu-r_{1}-\ldots-r_{N-1}}\left(\prod_{x=2}^{N}d_{x}(n^{x},\xi^{x})\right)\left(\prod_{d=1}^{N-1}(\rho_{d}^{R})^{\xi_{d}^{L+1}}\right)\nonumber
	\\=&
	\left(\prod_{k=1}^{N-1}\left(\rho_{k}^{\text{left}}\right)^{\xi_{k}^{0}+\xi_{k}^{1}}\right)\left(\prod_{x=2}^{L}d(n^{x},\xi^{x})\right)\left(\prod_{d=1}^{N-1}(\rho_{d}^{\text{right}})^{\xi_{d}^{L+1}}\right)
\end{align}
Where we used 
\begin{equation*}
\frac{\left(\nu-\sum_{c=1}^{N-1}\xi_{c}^{1}\right)!}{\nu!}\left(\prod_{b=1}^{N-1}\frac{r_{b}!}{(r_{b}-\xi^{1}_{b})!}\right)\frac{\nu!}{r_{1}!\cdots r_{N}!}=\frac{\left(\nu-\xi_{1}^{1}-\ldots-\xi_{N-1}^{1}\right)!}{(r_{1}-\xi_{1}^{1})!\cdots(r_{N-1}-\xi_{N-1}^{1})!(\nu-r_{1}-\ldots-r_{N-1})!}
\end{equation*}
and 
\begin{align*}
&\left(\prod_{a=1}^{N-1}\left(\frac{\alpha_{a}}{|\alpha|}\right)^{\xi_{a}^{0}-\xi_{a}^{1}+\xi_{a}^{1}}\right)\left(\prod_{k=1}^{N-1}\left(\frac{\alpha_{k}}{\alpha_{N}}\right)^{r_{k}}\right)\left(\prod_{\ell=1}^{N-1}\left(\frac{\alpha_{N}}{|\alpha|}\right)^{\nu-r_{\ell}+r_{\ell}}\right)\\=&	\left(\prod_{k=1}^{N-1}\left(\frac{\alpha_{k}}{|\alpha|}\right)^{\xi_{k}^{0}+\xi_{k}^{1}}\right)\left(\prod_{\ell=1}^{N-1}\left(\frac{\alpha_{\ell}}{|\alpha|}\right)^{r_{\ell}-\xi_{\ell}^{1}}\right)\left(\frac{\alpha_{N}}{|\alpha|}\right)^{\nu-r_{1}-\ldots-r_{N-1}}
\end{align*}
and the multinomial theorem in the last equality. \\ We now consider the second addend in the curly brackets of \eqref{ProvaBordoTH}.
We multiplying it by $\prod_{k=1}^{N-1}\left(\frac{\alpha_{N}}{|\alpha|}\right)^{-r_{k}+r_{k}}$ and we obtain
\begin{equation}\label{pezzo2}
	\begin{split}
		&D(\bm{n},\bm{\xi})\sum_{r_{1}=0}^{\nu}\ldots\sum_{r_{N}=0}^{\nu}\mathbbm{1}_{\{r_{1}+\ldots+r_{N}=\nu\}}\frac{\nu!}{r_{1}!\cdots r_{N}!}\left(\prod_{k=1}^{N-1}\left(\frac{\alpha_{k}}{\alpha_{N}}\right)^{r_{k}}\right)\left(\frac{\alpha_{N}}{|\alpha|}\right)^{\nu}\left(\frac{\alpha_{N}}{|\alpha|}\right)^{-(r_{1}+\ldots+r_{N-1})+(r_{1}+\ldots+r_{N-1})}\\=&D(\bm{n},\bm{\xi})
	\end{split}
\end{equation}
where we used the multinomial theorem. Therefore, replacing \eqref{pezzo1} and \eqref{pezzo2} in \eqref{ProvaBordoTH} we have that
\begin{align}
	\left(\mathcal{L}_{\text{left}}^{TH}D(\cdot,\bm{\xi})\right)(\bm{n})&=\left\{\prod_{a=1}^{N-1}\left(\frac{\alpha_{k}}{|\alpha|}\right)^{\xi_{a}^{0}+\xi_{a}^{1}}\prod_{x=2}^{N}d_{x}(n^{x},\xi^{x})\prod_{d=1}^{N-1}(\rho_{a}^{R})^{\xi_{d}^{L+1}}-D(\bm{n},\bm{\xi})\right\}
	\\&=
	\left(\widetilde{\mathcal{L}}_{\text{left}}^{TH}D(\bm{n},\cdot)\right)(\bm{\xi})
\end{align}
%The proof is similar for $\widetilde{\mathcal{L}}_{\text{right}}^{TH}$. 
\begin{flushright}
	$\square$
\end{flushright}
\section{Action on the occupation variable of $\mathcal{L}^{rd}$}\label{appendix-RD}
\paragraph{Action of $\mathcal{L}_{x,x+1}^{rd}$ on the occupation variable}
For all $x\in \{1,\ldots,L-1\}$ and for all $d\in \{1,\ldots,N-1\}$ we have
\begin{equation}\label{actionGraphRD}
	\begin{split}
		\mathcal{L}_{x,x+1}^{rd}n_{d}^{x}&=\nu \sigma_{11}(n_{d}^{x+1}-n_{d}^{x})+\nu\sigma_{12}\sum_{b=1\,:	,b\neq d}^{N-1}(n_{b}^{x+1}-n_{d}^{x})+(\Upsilon-2\nu\sigma_{12})\sum_{b=1}^{N-1}(n_{b}^{x}-n_{\alpha}^{x})
	\end{split}
\end{equation}
Indeed, 
\begin{enumerate}
	\item \textit{Stirring generator}:

\begin{equation}
	\mathcal{L}_{x,x+1}n_{d}^{x}=\sum_{a,b=1}^{N}n_{a}^{x}n_{b}^{x+1}\left((n_{d}^{x}-\bm{\delta}_{a}^{x}+\bm{\delta}_{b}^{x}+\bm{\delta}_{a}^{x+1}-\bm{\delta}_{b}^{x+1})-n_{d}^{x}\right)
\end{equation}
In the brackets of the right-hand-side of the above equation we have: a contribution $-n_{d}^{x}n_{b}^{x+1}$, when $a=d$; a contribution $n_{a}^{x}n_{d}^{x+1}$, when $b=d$. Thus we obtain 
\begin{align*}
	\mathcal{L}_{x,x+1}n_{d}^{x}=&\sum_{a=1}^{N}n_{a}^{x}n_{d}^{x+1}-\sum_{b=1}^{N}n_{d}^{x}n_{b}^{x+1}\\
	=&\sum_{a=1}^{N-1}n_{a}^{x}n_{d}^{x+1}+(\nu-\sum_{i=1}^{N-1}n_{i}^{x})n_{d}^{x+1}-\sum_{b=1}^{N-1}n_{d}^{x}n_{b}^{x+1}-n_{d}^{x}(\nu-\sum_{i=1}^{N-1}n_{i}^{x+1})\\
	=&\nu\left(n_{d}^{x+1}-n_{d}^{x}\right)
\end{align*} 
where we have used the fact that $n_{N}^{x}=\nu-\sum_{i=1}^{N-1}n_{i}^{x}$.
\item \textit{Stirring-mutation generator}: for every $c\in \{1,\ldots,N-2\}$ we have 
\begin{equation}
	\begin{split}
		\mathcal{L}_{x,x+1}^{c}n_{d}^{x}&=\sum_{a,b=1}^{N}n_{a}^{x}n_{b}^{x+1}\left((n_{d}^{x}-\bm{\delta}_{a}^{x}+\bm{\delta}_{h_{c}(b)}^{x}+\bm{\delta}_{h_{c}(a)}^{x+1}-\bm{\delta}_{b}^{x+1})-n_{d}^{x}\right)
	\end{split}
\end{equation}
In the brackets of the right-hand-side of the above equation we have: a contribution $-n_{d}^{x}n_{\ell}^{x+1}$, when $a=d$ (here we call $\ell=h_{c}(b)$); a contribution $n_{a}^{x}n_{q}^{x+1}$, when $h_{c}(b)=d$. Thus we obtain 
\begin{align}
		\mathcal{L}_{x,x+1}^{c}n_{d}^{x}&=\sum_{a=1}^{N}n_{a}^{x}n_{b}^{x+1}-\sum_{\ell=1}^{N}n_{d}^{x}n_{\ell}^{x+1}
		\\&=
		\sum_{k=1}^{N-1}n_{k}^{x}n_{b}^{x+1}+\left(\nu-\sum_{i=1}^{N-1}n_{i}^{x}\right)n_{b}^{x+1}-\sum_{\ell=1}^{N-1}n_{d}^{x}n_{\ell}^{x+1}-n_{d}^{x}\left(\nu-\sum_{i=1}^{N-1}n_{i}^{x+1}\right)
		\\&=
		\nu(n_{b}^{x+1}-n_{d}^{x})
\end{align}
where  we used the fact that $n_{N}^{x}=\nu-\sum_{i=1}^{N-1}n_{i}^{x}$. Let us observe that, by the definition of the map $h_{c}(\cdot)$ and for the fact that $1\leq c\leq N-2$, we have that in the above equation $b\neq d$ and $b\neq N$.
\item \textit{Pure-mutation generator}: here we have that 
\begin{equation}
	\mathcal{L}_{x,y}^{m}n_{d}^{x}=\sum_{a,b=1}^{N-1}n_{a}^{x}\left((n_{d}^{x}-\bm{\delta}_{a}^{x}+\bm{\delta}_{b}^{x})-n_{d}^{x}\right)=\sum_{b=1}^{N-1}(n_{b}^{x}-n_{d}^{x})
\end{equation}
\end{enumerate}
Then, summing over all indices $c\in\{1,\ldots,N-2\}$, we obtain \eqref{actionGraphRD} and \eqref{DifferenceEquation} follows.
\paragraph{Action of $\mathcal{L}_{\text{left}}^{rd}$ on the occupation variable}
Acting on $n_{d}^{1}$  with $\mathcal{L}_{\text{left}}^{rd}$ we obtain
\begin{equation}\label{actionBoudaryRD}
	\begin{split}
		\mathcal{L}_{\text{left}}^{rd}n_{d}^{1}&=\sum_{a,b=1}^{N}\alpha_{a}n_{b}^{1}\left((n_{d}^{1}-\bm{\delta}_{b}^{1}+\bm{\delta}_{a}^{1})-n_{d}^{1}\right)=\alpha_{d}\sum_{b=1\,:\,b\neq d}^{N}n_{b}^{1}-n_{d}^{1}\sum_{b=1\,:b\neq d}^{N}\alpha_{b}
	\end{split}
\end{equation}
Using equation \eqref{boundaryParamRD}, we obtain 
\begin{align}
			\mathcal{L}_{\text{left}}^{rd}n_{d}^{1}&=
		\alpha_{d}\left(\nu-n_{d}^{1}\right)-n_{d}^{1}\left(\nu(\sigma_{11}+(N-2)\sigma_{12})-\alpha_{d}\right)
		\\&=\alpha_{d}\nu-n_{d}^{1}\nu(\sigma_{11}+(N-2)\sigma_{12})
		\\&=
		\nu\sigma_{11}\rho_{d}^{\text{left}}+\nu\sigma_{12}\left(\sum_{b=1\,:\,b\neq d}^{N-1}\rho_{b}^{\text{left}}\right)-n_{d}^{1}\nu(\sigma_{11}+(N-2)\sigma_{12})
		\\&=
		\nu \sigma_{11} (\rho_{d}^{\text{left}}-n_{d}^{1})+\nu\sigma_{12}\sum_{b=1\,:\,b\neq d}^{N-1}(\rho_{b}^{\text{left}}-n_{d}^{1})
	\end{align}
Therefore, \eqref{actionBDLine} follows.
\bibliography{ref}
\bibliographystyle{unsrt}

\end{document}